### 2019-11-10 paper/submission.tex
Starting in on latex paper generation

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
%\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\begin{document}

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for CVPR Proceedings}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
   The ABSTRACT is to be in fully-justified italicized text, at the top
   of the left-hand column, below the author and affiliation
   information. Use the word ``Abstract'' as the title, in 12-point
   Times, boldface type, centered relative to the column, initially
   capitalized. The abstract is to be in 10-point, single-spaced type.
   Leave two blank lines after the Abstract, then begin the main text.
   Look at previous CVPR abstracts to get a feel for style and length.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Please follow the steps outlined below when submitting your manuscript to
the IEEE Computer Society Press.  This style guide now has several
important modifications (for example, you are no longer warned against the
use of sticky tape to attach your artwork to the paper), so all authors
should read this new version.

%-------------------------------------------------------------------------
\subsection{Language}

All manuscripts must be in English.

\subsection{Dual submission}

Please refer to the author guidelines on the CVPR 2018 web page for a
discussion of the policy on dual submissions.

\subsection{Paper length}
Papers, excluding the references section,
must be no longer than eight pages in length. The references section
will not be included in the page count, and there is no limit on the
length of the references section. For example, a paper of eight pages
with two pages of references would have a total length of 10 pages.
{\bf There will be no extra page charges for CVPR 2018.}

Overlength papers will simply not be reviewed.  This includes papers
where the margins and formatting are deemed to have been significantly
altered from those laid down by this style guide.  Note that this
\LaTeX\ guide already sets figure captions and references in a smaller font.
The reason such papers will not be reviewed is that there is no provision for
supervised revisions of manuscripts.  The reviewing process cannot determine
the suitability of the paper for presentation in eight pages if it is
reviewed in eleven.

%-------------------------------------------------------------------------
\subsection{The ruler}
The \LaTeX\ style defines a printed ruler which should be present in the
version submitted for review.  The ruler is provided in order that
reviewers may comment on particular lines in the paper without
circumlocution.  If you are preparing a document using a non-\LaTeX\
document preparation system, please arrange for an equivalent ruler to
appear on the final output pages.  The presence or absence of the ruler
should not change the appearance of any other content on the page.  The
camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
the \verb'\cvprfinalcopy' command in the document preamble.)  Reviewers:
note that the ruler measurements do not align well with lines in the paper
--- this turns out to be very difficult to do well when the paper contains
many figures and equations, and, when done, looks ugly.  Just use fractional
references (e.g.\ this line is $095.5$), although in most cases one would
expect that the approximate location will be adequate.

\subsection{Mathematics}

Please number all of your sections and displayed equations.  It is
important for readers to be able to refer to any particular equation.  Just
because you didn't refer to it in the text doesn't mean some future reader
might not need to refer to it.  It is cumbersome to have to use
circumlocutions like ``the equation second from the top of page 3 column
1''.  (Note that the ruler will not be present in the final copy, so is not
an alternative to equation numbers).  All authors will benefit from reading
Mermin's description of how to write mathematics:
\url{http://www.pamitc.org/documents/mermin.pdf}.

\subsection{Blind review}

Many authors misunderstand the concept of anonymizing for blind
review.  Blind review does not mean that one must remove
citations to one's own work---in fact it is often impossible to
review a paper unless the previous citations are known and
available.

Blind review means that you do not use the words ``my'' or ``our''
when citing previous work.  That is all.  (But see below for
techreports.)

Saying ``this builds on the work of Lucy Smith [1]'' does not say
that you are Lucy Smith; it says that you are building on her
work.  If you are Smith and Jones, do not say ``as we show in
[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
paper, include reference 7 as you would any other cited work.

An example of a bad paper just asking to be rejected:
\begin{quote}
\begin{center}
    An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of our
   previous paper [1], and show it to be inferior to all
   previously known methods.  Why the previous paper was
   accepted without this analysis is beyond me.

   [1] Removed for blind review
\end{quote}

An example of an acceptable paper:

\begin{quote}
\begin{center}
     An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of the
   paper of Smith \etal [1], and show it to be inferior to
   all previously known methods.  Why the previous paper
   was accepted without this analysis is beyond me.

   [1] Smith, L and Jones, C. ``The frobnicatable foo
   filter, a fundamental contribution to human knowledge''.
   Nature 381(12), 1-213.
\end{quote}

If you are making a submission to another conference at the same time,
which covers similar or overlapping material, you may need to refer to that
submission in order to explain the differences, just as you would if you
had previously published related work.  In such cases, include the
anonymized parallel submission~\cite{Authors14} as additional material and
cite it as
\begin{quote}
[1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324,
Supplied as additional material {\tt fg324.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details can be
found elsewhere, and refer them to a technical report.  For conference
submissions, the paper must stand on its own, and not {\em require} the
reviewer to go to a techreport for further details.  Thus, you may say in
the body of the paper ``further details may be found
in~\cite{Authors14b}''.  Then submit the techreport as additional material.
Again, you may not assume the reviewers will read this material.

Sometimes your paper is about a problem which you tested using a tool which
is widely known to be restricted to a single institution.  For example,
let's say it's 1969, you have solved a key problem on the Apollo lander,
and you believe that the CVPR70 audience would like to hear about your
solution.  The work is a development of your celebrated 1968 paper entitled
``Zero-g frobnication: How being the only people in the world with access to
the Apollo lander source code makes us a wow at parties'', by Zeus \etal.

You can handle this paper like any other.  Don't write ``We show how to
improve our previous work [Anonymous, 1968].  This time we tested the
algorithm on a lunar lander [name of lander removed for blind review]''.
That would be silly, and would immediately identify the authors. Instead
write the following:
\begin{quotation}
\noindent
   We describe a system for zero-g frobnication.  This
   system is new because it handles the following cases:
   A, B.  Previous systems [Zeus et al. 1968] didn't
   handle case B properly.  Ours handles it by including
   a foo term in the bar integral.

   ...

   The proposed system was integrated with the Apollo
   lunar lander, and went all the way to the moon, don't
   you know.  It displayed the following behaviours
   which show how well we solved cases A and B: ...
\end{quotation}
As you can see, the above text follows standard scientific convention,
reads better than the first version, and does not explicitly name you as
the authors.  A reviewer might think it likely that the new paper was
written by Zeus \etal, but cannot make any decision based on that guess.
He or she would have to be sure that no other authors could have been
contracted to solve problem B.
\medskip

\noindent
FAQ\medskip\\
{\bf Q:} Are acknowledgements OK?\\
{\bf A:} No.  Leave them for the final copy.\medskip\\
{\bf Q:} How do I cite my results reported in open challenges?
{\bf A:} To conform with the double blind review policy, you can report results of other challenge participants together with your results in your paper. For your results, however, you should not identify yourself and should not mention your participation in the challenge. Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\

\begin{figure}[t]
\begin{center}
\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
\end{center}
   \caption{Example of caption.  It is set in Roman so that mathematics
   (always set in Roman: $B \sin A = A \sin B$) may be included without an
   ugly clash.}
\label{fig:long}
\label{fig:onecol}
\end{figure}

\subsection{Miscellaneous}

\noindent
Compare the following:\\
\begin{tabular}{ll}
 \verb'$conf_a$' &  $conf_a$ \\
 \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
\end{tabular}\\
See The \TeX book, p165.

The space after \eg, meaning ``for example'', should not be a
sentence-ending space. So \eg is correct, {\em e.g.} is not.  The provided
\verb'\eg' macro takes care of this.

When citing a multi-author paper, you may save space by using ``et alia'',
shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word.)
However, use it only when there are three or more authors.  Thus, the
following is correct: ``
   Frobnication has been trendy lately.
   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''

This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...''
because reference~\cite{Alpher03} has just two authors.  If you use the
\verb'\etal' macro provided, then you need not worry about double periods
when used at the end of a sentence as in Alpher \etal.

For this citation style, keep multiple citations in numerical (not
chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
\cite{Alpher02,Alpher03,Authors14}.

\begin{figure*}
\begin{center}
\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
\end{center}
   \caption{Example of a short caption, which should be centered.}
\label{fig:short}
\end{figure*}

%------------------------------------------------------------------------
\section{Formatting your paper}

All text must be in a two-column format. The total allowable width of the
text area is $6\frac78$ inches (17.5 cm) wide by $8\frac78$ inches (22.54
cm) high. Columns are to be $3\frac14$ inches (8.25 cm) wide, with a
$\frac{5}{16}$ inch (0.8 cm) space between them. The main title (on the
first page) should begin 1.0 inch (2.54 cm) from the top edge of the
page. The second and following pages should begin 1.0 inch (2.54 cm) from
the top edge. On all pages, the bottom margin should be 1-1/8 inches (2.86
cm) from the bottom edge of the page for $8.5 \times 11$-inch paper; for A4
paper, approximately 1-5/8 inches (4.13 cm) from the bottom edge of the
page.

%-------------------------------------------------------------------------
\subsection{Margins and page numbering}

All printed material, including text, illustrations, and charts, must be kept
within a print area 6-7/8 inches (17.5 cm) wide by 8-7/8 inches (22.54 cm)
high.
Page numbers should be in footer with page numbers, centered and .75
inches from the bottom of the page and make it start at the correct page
number rather than the 4321 in the example.  To do this fine the line (around
line 23)
\begin{verbatim}
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\end{verbatim}
where the number 4321 is your assigned starting page.

Make sure the first page is numbered by commenting out the first page being
empty on line 46
\begin{verbatim}
%\thispagestyle{empty}
\end{verbatim}

%-------------------------------------------------------------------------
\subsection{Type-style and fonts}

Wherever Times is specified, Times Roman may also be used. If neither is
available on your word processor, please use the font closest in
appearance to Times to which you have access.

MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge of
the first page. The title should be in Times 14-point, boldface type.
Capitalize the first letter of nouns, pronouns, verbs, adjectives, and
adverbs; do not capitalize articles, coordinate conjunctions, or
prepositions (unless the title begins with such a word). Leave two blank
lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
and printed in Times 12-point, non-boldface type. This information is to
be followed by two blank lines.

The ABSTRACT and MAIN TEXT are to be in a two-column format.

MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use
double-spacing. All paragraphs should be indented 1 pica (approx. 1/6
inch or 0.422 cm). Make sure your text is fully justified---that is,
flush left and flush right. Please do not place any additional blank
lines between paragraphs.

Figure and table captions should be 9-point Roman type as in
Figures~\ref{fig:onecol} and~\ref{fig:short}.  Short captions should be centred.

\noindent Callouts should be 9-point Helvetica, non-boldface type.
Initially capitalize only the first word of section titles and first-,
second-, and third-order headings.

FIRST-ORDER HEADINGS. (For example, {\large \bf 1. Introduction})
should be Times 12-point boldface, initially capitalized, flush left,
with one blank line before, and one blank line after.

SECOND-ORDER HEADINGS. (For example, { \bf 1.1. Database elements})
should be Times 11-point boldface, initially capitalized, flush left,
with one blank line before, and one after. If you require a third-order
heading (we discourage it), use 10-point Times, boldface, initially
capitalized, flush left, preceded by one blank line, followed by a period
and your text on the same line.

%-------------------------------------------------------------------------
\subsection{Footnotes}

Please use footnotes\footnote {This is what a footnote looks like.  It
often distracts the reader from the main flow of the argument.} sparingly.
Indeed, try to avoid footnotes altogether and include necessary peripheral
observations in
the text (within parentheses, if you prefer, as in this sentence).  If you
wish to use a footnote, place it at the bottom of the column on the page on
which it is referenced. Use Times 8-point type, single-spaced.

%-------------------------------------------------------------------------
\subsection{References}

List and number all bibliographical references in 9-point Times,
single-spaced, at the end of your paper. When referenced in the text,
enclose the citation number in square brackets, for
example~\cite{Authors14}.  Where appropriate, include the name(s) of
editors of referenced books.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
Method & Frobnability \\
\hline\hline
Theirs & Frumpy \\
Yours & Frobbly \\
Ours & Makes one's heart Frob\\
\hline
\end{tabular}
\end{center}
\caption{Results.   Ours is better.}
\end{table}

%-------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}

All graphics should be centered.  Please ensure that any point you wish to
make is resolvable in a printed copy of the paper.  Resize fonts in figures
to match the font in the body text, and choose line widths which render
effectively in print.  Many readers (and reviewers), even of an electronic
copy, will choose to print your paper in order to read it.  You cannot
insist that they do otherwise, and therefore must not assume that they can
zoom in to see tiny details on a graphic.

When placing figures in \LaTeX, it's almost always best to use
\verb+\includegraphics+, and to specify the  figure width as a multiple of
the line width as in the example below
{\small\begin{verbatim}
   \usepackage[dvips]{graphicx} ...
   \includegraphics[width=0.8\linewidth]
                   {myfile.eps}
\end{verbatim}
}

%-------------------------------------------------------------------------
\subsection{Color}

Please refer to the author guidelines on the CVPR 2018 web page for a discussion
of the use of color in your document.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit
your finished paper. We MUST have this form before your paper can be
published in the proceedings.

Please direct any questions to the production editor in charge of these
proceedings at the IEEE Computer Society Press: Phone (714) 821-8380, or
Fax (714) 761-1784.

{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}

\end{document}

### 2019-11-10 image.py

@@ -1,30 +1,0 @@
-import os
-import matplotlib.pyplot as plt
-import numpy as np
-
-import nibabel as nib
-from nibabel.testing import data_path
-from nilearn import image
-from nilearn import plotting
-
-
-img_path = "/home/evan/Downloads/ds000221_R1.0.0/sub-010004/ses-02/func/sub-010004_ses-02_task-rest_acq-AP_run-01_bold.nii.gz"
-# img_path = os.path.join(data_path, 'example4d.nii.gz')
-# img = nib.load(img_path)
-# # print(img)
-# print(img.shape)
-
-# plotting.plot_img(example_filename)
-
-# img3d = image.index_img(img_path, 0)
-# print(img3d.shape)
-# plotting.plot_stat_map(img3d)
-# plotting.plot_img(img3d)
-# plt.show()
-
-print("bigboi")
-all_images = list(image.iter_img(img_path))
-for img in all_images[:5]:
-    # img is now an in-memory 3D img
-    plotting.plot_stat_map(img, threshold=3, display_mode="z", cut_coords=1, colorbar=False)
-plotting.show()

### 2019-11-10 src/__main__.py

@@ -1,6 +1,16 @@
 """
 TODO
-1.
+
+Awesome!
+
+Some TODO's to work on in the future
+- More subjects (memory management)
+- Cut timesteps that could either belong to one event or another (boundary dispute) as I'm not 100% confident in event assignment
+
+- Deep learn, VGG
+- Find convolution on non flattened image (3d convolution could get brain regions better?)
+
+- Review HW for good ideas in training, data manipulation, etc
 """
 
 import os

### 2019-11-10 src/data.py

@@ -12,8 +12,10 @@ from bids import BIDSLayout
 DATASET_PATH = os.path.join("data", "ds000157")
 
 # from the paper, time in seconds each scan covers
-SCAN_TIME = 1.6 + 0.023  # time plus echo time
+SCAN_TIME = 1.6  # I believe the 0.023 echo time is contained within total time
 USEFUL_SCANS = 370  # from the description of the t2-weighted scans, supposedly there's 370 unless that's a typo
+
+TVT_SPLITS = (0.7, 0.9)
 
 
 class MLDataset:
@@ -66,12 +68,12 @@ class MLDataset:
         self._flattened = True
 
 
-def get_food_temptation_data() -> BIDSLayout:
+def get_food_temptation_data(dataset_path: str = DATASET_PATH) -> BIDSLayout:
     """
     This project is hardcoded to work with this specific OpenNeuro dataset
     """
-    print("Using dataset path: \"{}\"".format(DATASET_PATH))
-    layout = BIDSLayout(DATASET_PATH)
+    print("Using dataset path: \"{}\"".format(dataset_path))
+    layout = BIDSLayout(dataset_path)
 
     print(layout)
     return layout
@@ -168,6 +170,23 @@ def get_subject_data(layout: BIDSLayout, subject: str):
     flip angle = 72.5◦, FOV= 208 × 119 × 256 mm, SENSE factor AP = 2.4, 30 axial
     3.6 mm slices with 0.4 mm gap, reconstructed voxel size = 4 mm × 4 mm × 4 mm).
     In one functional run 370 scans were made (∼10 min). 
+    
+    During the functional run, stimuli were presented on a screen...
+    
+    B is the T2-weighted, 
+    64 (either SAGITAL/CORONAL) 
+      x 64 (other of SAGITAL/CORONAL) 
+      x 30 (axial/horizontal, accounts for verticallity) 
+      x 370-375 (scans, accounts for TIME)
+      
+    The functional volumes of every subject were realigned to the first
+    volume of the first run, globally normalized to Montreal Neurological Institute space
+    (MNI space) retaining 4 mm × 4 mm × 4 mm voxels, and spatially smoothed with a
+    gaussian kernel of 8 mm full width at half maximum. A statistical parametric map
+    was generated for every subject by fitting a boxcar function to each time series, 
+    convolved with the canonical hemodynamic response function. Data were high-pass
+    filtered with a cutoff of 128 s. Three conditions were modeled: viewing foods, 
+    viewing non-foods and the half-way break.
[…]

### 2019-11-10 src/models/sanity_check.py

@@ -5,12 +5,14 @@ import pandas as pd
 from sklearn.metrics import precision_recall_fscore_support
 from sklearn.svm import SVC
 
+from bids import BIDSLayout
+
 from src import data
 
 
-def test_ml_data(limit: int = 300):
+def test_ml_data(layout: BIDSLayout = None, limit: int = 300):
 
-    ml_dataset = data.get_ml_dataset()
+    ml_dataset = data.get_ml_dataset(layout=layout)
     print("Data gathered")
     print(ml_dataset)
     ml_dataset.normalize()
@@ -28,7 +30,5 @@ def test_ml_data(limit: int = 300):
         columns=['food', 'nonfood'],
         index=['precision', 'recall', 'fscore', 'support']
     )
-
-    # TODO bother with test? meh
 
     return results

### 2019-11-11 paper/submission.tex
Updating paper, and some misc code edits I forget about

@@
 %\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{4321}
+\setcounter{page}{1}
 \begin{document}
 
-%%%%%%%%% TITLE
-\title{\LaTeX\ Author Guidelines for CVPR Proceedings}
+%-------------------------------------------------------------------------
+\title{Deep Learning for 3D Brain Activity Scans}
 
-\author{First Author\\
-Institution1\\
-Institution1 address\\
-{\tt\small firstauthor@i1.org}
-% For a paper whose authors are all at the same institution,
-% omit the following lines up until the closing ``}''.
-% Additional authors and addresses can be added with ``\and'',
-% just like the second author.
-% To save space, use either the email address or home page, not both
-\and
-Second Author\\
-Institution2\\
-First line of institution2 address\\
-{\tt\small secondauthor@i2.org}
+\author{Evan Albright\\
+University of Massachusetts Amherst\\
+140 Governors Dr, Amherst, MA 01002\\
+{\tt\small ealbright@umass.edu}
 }
@@
 
-%%%%%%%%% ABSTRACT
+%-------------------------------------------------------------------------
 \begin{abstract}
-   The ABSTRACT is to be in fully-justified italicized text, at the top
-   of the left-hand column, below the author and affiliation
-   information. Use the word ``Abstract'' as the title, in 12-point
-   Times, boldface type, centered relative to the column, initially
-   capitalized. The abstract is to be in 10-point, single-spaced type.
-   Leave two blank lines after the Abstract, then begin the main text.
-   Look at previous CVPR abstracts to get a feel for style and length.
+   Abstract placeholder
 \end{abstract}
 
-%%%%%%%%% BODY TEXT
-\section{Introduction}
+% ----------------------------------------------------------------------
+% INTRODUCTION
+% ----------------------------------------------------------------------
+\section{Introduction}\label{sec:introduction}
+Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
+for mental health research and disease diagnosis.
+High dimensional, massive datasets, with intertwining patterns of activity and structure in three or four
+dimensions, are a lot to manage by hand.
 
-Please follow the steps outlined below when submitting your manuscript to
-the IEEE Computer Society Press.  This style guide now has several
-important modifications (for example, you are no longer warned against the
-use of sticky tape to attach your artwork to the paper), so all authors
-should read this new version.
+Classification of activity and structures should come quite naturally with deep learning practices.
+Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
+yet often only as a proof of concept.
+Others are working now to try and find universal architectures for most
+neuro-imaging tasks~\cite{henschel2019fastsurfer}, though it seems this is still far from a solved
+application with satisfactory performances across the board.
 
-%-------------------------------------------------------------------------
-\subsection{Language}
+Of more interest to myself, deep learning networks in their proven capacity to segment images based on what entities
+they identify in classification could surely be adopted to highlight what patterns of activity have informed their
+own decision making and perhaps clue researchers in to overlooked details.
 
-All manuscripts must be in English.
+My ambition is to create a highly accurate 3d convolutional network to first classify, then segment brain scans to
+help with my specific application of neuroimaging.
 
-\subsection{Dual submission}
+I may also attempt to explore feature analysis and reduction which has been shown to
+help~\cite{shi2018feature} with deep learning tasks, if only to manage the challenging
+dimensionality of their 3d inputs.
 
-Please refer to the author guidelines on the CVPR 2018 web page for a
-discussion of the policy on dual submissions.
+% ----------------------------------------------------------------------
+% PROBLEM STATEMENT
+% ----------------------------------------------------------------------
+\section{Problem Statement}\label{sec:problem-statement}
+The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
+Netherlands circa 2013 who were presented with various grids of images featuring
+either neutral entities, (e.g.\ office chairs, windows, jackets) or foods~\cite{smeets2013allured}.
+The foods were specifically types of calorically dense, tempting foods, which the individuals viewing would likely
+have formed restraint with given their success in maintaining healthy body mass indices.
+The experiment was meant to highlight brain structures roles in anticipated rewards from consuming delicious foods,
+balanced against the hypothesized activation of restraints.
+In neuro-imaging, this task is known as passive image viewing.
 
-\subsection{Paper length}
-Papers, excluding the references section,
-must be no longer than eight pages in length. The references section
-will not be included in the page count, and there is no limit on the
-length of the references section. For example, a paper of eight pages
-with two pages of references would have a total length of 10 pages.
-{\bf There will be no extra page charges for CVPR 2018.}
+Given the quite important role of food and willpower in our lives, the activation patterns are likely to be powerful
+and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
+This combination gives me high hopes of success in this specific application of learning to classify the viewed
+objects through brain activations of the subjects.
 
-Overlength papers will simply not be reviewed.  This includes papers
-where the margins and formatting are deemed to have been significantly
-altered from those laid down by this style guide.  Note that this
-\LaTeX\ guide already sets figure captions and references in a smaller font.
-The reason such papers will not be reviewed is that there is no provision for
-supervised revisions of manuscripts.  The reviewing process cannot determine
-the suitability of the paper for presentation in eight pages if it is
-reviewed in eleven.
+I hope given the clarity in patterns in the brain scans, that I can develop a neural network to highlight what
+features and locations in the brain scans have been helpful in informing the decisions of classification, in the
+hope that such methods could be adapted to harder tasks in the future.
 
-%-------------------------------------------------------------------------
-\subsection{The ruler}
-The \LaTeX\ style defines a printed ruler which should be present in the
-version submitted for review.  The ruler is provided in order that
-reviewers may comment on particular lines in the paper without
-circumlocution.  If you are preparing a document using a non-\LaTeX\
-document preparation system, please arrange for an equivalent ruler to
-appear on the final output pages.  The presence or absence of the ruler
-should not change the appearance of any other content on the page.  The
-camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
-the \verb'\cvprfinalcopy' command in the document preamble.)  Reviewers:
-note that the ruler measurements do not align well with lines in the paper
---- this turns out to be very difficult to do well when the paper contains
-many figures and equations, and, when done, looks ugly.  Just use fractional
-references (e.g.\ this line is $095.5$), although in most cases one would
-expect that the approximate location will be adequate.
+The dataset can be found from the excellent
+OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}
+It consists of a survey detailing the relationship of each subject with food and diet taken after the experimental
+proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w brain scan,
+and time series of bold scans of about 370-375 scans, which encompasses the approximately 10 minute experiment
+in its entirety.
+The bold scans are quite sizeable given the number of them, however they are a relatively low resolution, consisting
+of 64 sagital by 64 coronal by 30 axial scans.
 
-\subsection{Mathematics}
+You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.
 
-Please number all of your sections and displayed equations.  It is
-important for readers to be able to refer to any particular equation.  Just
-because you didn't refer to it in the text doesn't mean some future reader
-might not need to refer to it.  It is cumbersome to have to use
-circumlocutions like ``the equation second from the top of page 3 column
-1''.  (Note that the ruler will not be present in the final copy, so is not
-an alternative to equation numbers).  All authors will benefit from reading
-Mermin's description of how to write mathematics:
-\url{http://www.pamitc.org/documents/mermin.pdf}.
+You may find this project repository here: \url{https://github.com/e-m-albright/CS682}
+Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
+abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.
 
-\subsection{Blind review}
+% ----------------------------------------------------------------------
+% TECHNICAL APPROACH
+% ----------------------------------------------------------------------
+\section{Technical Approach}\label{sec:technical-approach}
+I intend to explore the efficacy of 3d convolutional networks which have been shown to have a slight performance
+edge on 2d convolutions~\cite{payan2015predicting}.
+Given the nature of the input images being 3d brain scans this makes intuitive sense and could allow for much more
+interesting patterns of brain activity and structuring being recognized.
 
-Many authors misunderstand the concept of anonymizing for blind
-review.  Blind review does not mean that one must remove
-citations to one's own work---in fact it is often impossible to
-review a paper unless the previous citations are known and
-available.
+As for the deeper architecture, I'm hoping to imitate some of the more successful networks like VGG or ResNet,
+however I wouldn't be surprised if they don't adapt well to this particular application, especially given how
+interesting and important it could be to keep the input data in a three dimensional format.
 
-Blind review means that you do not use the words ``my'' or ``our''
-when citing previous work.  That is all.  (But see below for
-techreports.)
+It is my hope that I can also learn from other developments of object detection to see what regions in the
+brain become active for a particular classification, like Faster R-CNN
 
-Saying ``this builds on the work of Lucy Smith [1]'' does not say
-that you are Lucy Smith; it says that you are building on her
-work.  If you are Smith and Jones, do not say ``as we show in
-[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
-paper, include reference 7 as you would any other cited work.
+\subsection{Novelty}\label{subsec:novelty}
 
-An example of a bad paper just asking to be rejected:
-\begin{quote}
-\begin{center}
-    An analysis of the frobnicatable foo filter.
-\end{center}
+I have not seen any application of classification or object segmentation to neuroimaging in 3d
+on brain activity patterns.
+I have seen some works which I hope to leverage with similar aims like that of structural identification
+for tumors, in two dimensions~\cite{akkus2017deep}.
+However at the very least I hope to produce a very successful network on my chosen problem first and foremost rather
+than assume I am the only person to ever think of something as intuitive as bringing neural networks to bear
+on neuro-imaging tasks.
 
-   In this paper we present a performance analysis of our
-   previous paper [1], and show it to be inferior to all
-   previously known methods.  Why the previous paper was
-   accepted without this analysis is beyond me.
+I hope to find a methodology of first classifying correctly the brain scans, improve accuracy as much as possible,
+and adapt the approach into tools for identifying important regions of activity in the
+brain which are being utilized for classification.
 
-   [1] Removed for blind review
-\end{quote}
+I am not chiefly focused on coping with the challenges of memory but it may become too important to overlook.
+Each dataset input is a $64 \times 64 \times 30$ slice of floats, for which we've got
+about $375 scans \times 30 subjects$ total.
+In my work on preliminary results I have already been confronted by the limitations of my current machine and have
+been trying working around the issue until I am ready to attempt to resolve it.
+This may be an unavoidable area to simultaneously try to innovate for the field of
+neuro-imaging analysis with deep learning
 
-An example of an acceptable paper:
+% ----------------------------------------------------------------------
+% INTERMEDIATE RESULTS
+% ----------------------------------------------------------------------
+\section{Intermediate Results}\label{sec:intermediate-results}
+At the date of the milestone, I have shown that my experimental set-up is complete to begin
+iterating towards the novel contributions aimed for by this project.
 
-\begin{quote}
-\begin{center}
-     An analysis of the frobnicatable foo filter.
-\end{center}
+I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
+network for classification on my chosen dataset incorporating my own contributions to the process.
 
-   In this paper we present a performance analysis of the
-   paper of Smith \etal [1], and show it to be inferior to
-   all previously known methods.  Why the previous paper
-   was accepted without this analysis is beyond me.
+I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
+excited nevertheless to try.
 
-   [1] Smith, L and Jones, C. ``The frobnicatable foo
-   filter, a fundamental contribution to human knowledge''.
-   Nature 381(12), 1-213.
-\end{quote}
+\subsection{Data}\label{subsec:data}
+I have become familiar with the chosen dataset, and created a machine learning friendly format
+based off of the data.
 
-If you are making a submission to another conference at the same time,
-which covers similar or overlapping material, you may need to refer to that
-submission in order to explain the differences, just as you would if you
-had previously published related work.  In such cases, include the
-anonymized parallel submission~\cite{Authors14} as additional material and
-cite it as
-\begin{quote}
-[1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324,
-Supplied as additional material {\tt fg324.pdf}.
-\end{quote}
+I have inferred classification labels from the experimental events timings and the estimated time
+of each complete scan in the dataset.
 
-Finally, you may feel you need to tell the reader that more details can be
-found elsewhere, and refer them to a technical report.  For conference
-submissions, the paper must stand on its own, and not {\em require} the
-reviewer to go to a techreport for further details.  Thus, you may say in
-the body of the paper ``further details may be found
-in~\cite{Authors14b}''.  Then submit the techreport as additional material.
-Again, you may not assume the reviewers will read this material.
+\subsection{Sanity Checks}\label{subsec:sanity-checks}
+Using the data, I have implemented a very simple, yet effective 80\% accurate SVM classifier to prove the data
+should be ready for use with my chosen deep learning methodologies.
 
-Sometimes your paper is about a problem which you tested using a tool which
-is widely known to be restricted to a single institution.  For example,
-let's say it's 1969, you have solved a key problem on the Apollo lander,
-and you believe that the CVPR70 audience would like to hear about your
-solution.  The work is a development of your celebrated 1968 paper entitled
-``Zero-g frobnication: How being the only people in the world with access to
-the Apollo lander source code makes us a wow at parties'', by Zeus \etal.
+I'm now confident in the classifications of each of the scans (about 375 per subject, for 30 subjects)
 
-You can handle this paper like any other.  Don't write ``We show how to
-improve our previous work [Anonymous, 1968].  This time we tested the
-algorithm on a lunar lander [name of lander removed for blind review]''.
-That would be silly, and would immediately identify the authors. Instead
-write the following:
-\begin{quotation}
-\noindent
-   We describe a system for zero-g frobnication.  This
-   system is new because it handles the following cases:
-   A, B.  Previous systems [Zeus et al. 1968] didn't
-   handle case B properly.  Ours handles it by including
-   a foo term in the bar integral.
+\subsection{Deep Learning \& Future Works}\label{subsec:deep-learning}
+I have tried feeding the data into very basic neural networks with mixed success, and am in the active process of
+troubleshooting.
+Once I have matched performance with the SVM using an implementation of a reasonable neural network using PyTorch, I
+will begin the process of looking for refinements and innovations alluded to in previous sections, as well as
+transitioning the task to object detection if I can.
 
-   ...
-
-   The proposed system was integrated with the Apollo
-   lunar lander, and went all the way to the moon, don't
-   you know.  It displayed the following behaviours
-   which show how well we solved cases A and B: ...
-\end{quotation}
-As you can see, the above text follows standard scientific convention,
-reads better than the first version, and does not explicitly name you as
-the authors.  A reviewer might think it likely that the new paper was
-written by Zeus \etal, but cannot make any decision based on that guess.
-He or she would have to be sure that no other authors could have been
-contracted to solve problem B.
-\medskip
-
-\noindent
-FAQ\medskip\\
-{\bf Q:} Are acknowledgements OK?\\
-{\bf A:} No.  Leave them for the final copy.\medskip\\
-{\bf Q:} How do I cite my results reported in open challenges?
-{\bf A:} To conform with the double blind review policy, you can report results of other challenge participants together with your results in your paper.
-For your results, however, you should not identify yourself and should not mention your participation in the challenge.
-Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\
-
-\begin{figure}[t]
-\begin{center}
-\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
-   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
-\end{center}
-   \caption{Example of caption.  It is set in Roman so that mathematics
-   (always set in Roman: $B \sin A = A \sin B$) may be included without an
-   ugly clash.}
-\label{fig:long}
-\label{fig:onecol}
-\end{figure}
-
-\subsection{Miscellaneous}
-
-\noindent
-Compare the following:\\
-\begin{tabular}{ll}
- \verb'$conf_a$' &  $conf_a$ \\
- \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
-\end{tabular}\\
-See The \TeX book, p165.
-
-The space after \eg, meaning ``for example'', should not be a
-sentence-ending space. So \eg is correct, {\em e.g.} is not.  The provided
-\verb'\eg' macro takes care of this.
-
-When citing a multi-author paper, you may save space by using ``et alia'',
-shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word.)
-However, use it only when there are three or more authors.  Thus, the
-following is correct: ``
-   Frobnication has been trendy lately.
-   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
-   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''
-
-This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...''
-because reference~\cite{Alpher03} has just two authors.  If you use the
-\verb'\etal' macro provided, then you need not worry about double periods
-when used at the end of a sentence as in Alpher \etal.
-
-For this citation style, keep multiple citations in numerical (not
-chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
-\cite{Alpher02,Alpher03,Authors14}.
-
-\begin{figure*}
-\begin{center}
-\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
-\end{center}
-   \caption{Example of a short caption, which should be centered.}
-\label{fig:short}
-\end{figure*}
-
-%------------------------------------------------------------------------
-\section{Formatting your paper}
-
-All text must be in a two-column format. The total allowable width of the
-text area is $6\frac78$ inches (17.5 cm) wide by $8\frac78$ inches (22.54
-cm) high. Columns are to be $3\frac14$ inches (8.25 cm) wide, with a
-$\frac{5}{16}$ inch (0.8 cm) space between them. The main title (on the
-first page) should begin 1.0 inch (2.54 cm) from the top edge of the
-page. The second and following pages should begin 1.0 inch (2.54 cm) from
-the top edge. On all pages, the bottom margin should be 1-1/8 inches (2.86
-cm) from the bottom edge of the page for $8.5 \times 11$-inch paper; for A4
-paper, approximately 1-5/8 inches (4.13 cm) from the bottom edge of the
-page.
-
-%-------------------------------------------------------------------------
-\subsection{Margins and page numbering}
-
-All printed material, including text, illustrations, and charts, must be kept
-within a print area 6-7/8 inches (17.5 cm) wide by 8-7/8 inches (22.54 cm)
-high.
-Page numbers should be in footer with page numbers, centered and .75
-inches from the bottom of the page and make it start at the correct page
-number rather than the 4321 in the example.  To do this fine the line (around
-line 23)
-\begin{verbatim}
-%\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{4321}
-\end{verbatim}
-where the number 4321 is your assigned starting page.
-
-Make sure the first page is numbered by commenting out the first page being
-empty on line 46
-\begin{verbatim}
-%\thispagestyle{empty}
-\end{verbatim}
-
-%-------------------------------------------------------------------------
-\subsection{Type-style and fonts}
-
-Wherever Times is specified, Times Roman may also be used. If neither is
-available on your word processor, please use the font closest in
-appearance to Times to which you have access.
-
-MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge of
-the first page. The title should be in Times 14-point, boldface type.
-Capitalize the first letter of nouns, pronouns, verbs, adjectives, and
-adverbs; do not capitalize articles, coordinate conjunctions, or
-prepositions (unless the title begins with such a word). Leave two blank
-lines after the title.
-
-AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
-and printed in Times 12-point, non-boldface type. This information is to
-be followed by two blank lines.
-
-The ABSTRACT and MAIN TEXT are to be in a two-column format.
-
-MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use
-double-spacing. All paragraphs should be indented 1 pica (approx. 1/6
-inch or 0.422 cm). Make sure your text is fully justified---that is,
-flush left and flush right. Please do not place any additional blank
-lines between paragraphs.
-
-Figure and table captions should be 9-point Roman type as in
-Figures~\ref{fig:onecol} and~\ref{fig:short}.  Short captions should be centred.
-
-\noindent Callouts should be 9-point Helvetica, non-boldface type.
-Initially capitalize only the first word of section titles and first-,
-second-, and third-order headings.
-
-FIRST-ORDER HEADINGS. (For example, {\large \bf 1. Introduction})
-should be Times 12-point boldface, initially capitalized, flush left,
-with one blank line before, and one blank line after.
-
-SECOND-ORDER HEADINGS. (For example, { \bf 1.1. Database elements})
-should be Times 11-point boldface, initially capitalized, flush left,
-with one blank line before, and one after. If you require a third-order
-heading (we discourage it), use 10-point Times, boldface, initially
-capitalized, flush left, preceded by one blank line, followed by a period
-and your text on the same line.
-
-%-------------------------------------------------------------------------
-\subsection{Footnotes}
-
-Please use footnotes\footnote {This is what a footnote looks like.  It
-often distracts the reader from the main flow of the argument.} sparingly.
-Indeed, try to avoid footnotes altogether and include necessary peripheral
-observations in
-the text (within parentheses, if you prefer, as in this sentence).  If you
-wish to use a footnote, place it at the bottom of the column on the page on
-which it is referenced. Use Times 8-point type, single-spaced.
-
-%-------------------------------------------------------------------------
-\subsection{References}
-
-List and number all bibliographical references in 9-point Times,
-single-spaced, at the end of your paper. When referenced in the text,
-enclose the citation number in square brackets, for
-example~\cite{Authors14}.  Where appropriate, include the name(s) of
-editors of referenced books.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|}
-\hline
-Method & Frobnability \\
-\hline\hline
-Theirs & Frumpy \\
-Yours & Frobbly \\
-Ours & Makes one's heart Frob\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Results.   Ours is better.}
-\end{table}
-
-%-------------------------------------------------------------------------
-\subsection{Illustrations, graphs, and photographs}
-
-All graphics should be centered.  Please ensure that any point you wish to
-make is resolvable in a printed copy of the paper.  Resize fonts in figures
-to match the font in the body text, and choose line widths which render
-effectively in print.  Many readers (and reviewers), even of an electronic
-copy, will choose to print your paper in order to read it.  You cannot
-insist that they do otherwise, and therefore must not assume that they can
-zoom in to see tiny details on a graphic.
-
-When placing figures in \LaTeX, it's almost always best to use
-\verb+\includegraphics+, and to specify the  figure width as a multiple of
-the line width as in the example below
-{\small\begin{verbatim}
-   \usepackage[dvips]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]
-                   {myfile.eps}
-\end{verbatim}
-}
-
-%-------------------------------------------------------------------------
-\subsection{Color}
-
-Please refer to the author guidelines on the CVPR 2018 web page for a discussion
-of the use of color in your document.
-
-%------------------------------------------------------------------------
-\section{Final copy}
-
-You must include your signed IEEE copyright release form when you submit
-your finished paper. We MUST have this form before your paper can be
-published in the proceedings.
-
-Please direct any questions to the production editor in charge of these
-proceedings at the IEEE Computer Society Press: Phone (714) 821-8380, or
-Fax (714) 761-1784.
-
+% ----------------------------------------------------------------------
+% END
+% ----------------------------------------------------------------------
 {\small
 \bibliographystyle{ieee}
-\bibliography{egbib}
+\bibliography{project}
 }

### 2019-11-11 src/data.py

@@ -174,7 +174,7 @@ def get_subject_data(layout: BIDSLayout, subject: str):
     During the functional run, stimuli were presented on a screen...
     
     B is the T2-weighted, 
-    64 (either SAGITAL/CORONAL) 
+      64 (either SAGITAL/CORONAL) 
       x 64 (other of SAGITAL/CORONAL) 
       x 30 (axial/horizontal, accounts for verticallity) 
       x 370-375 (scans, accounts for TIME)

### 2019-11-11 src/models/vgg_check.py

@@ -1,0 +1,80 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+from torch.autograd import Variable
+from torch.optim import lr_scheduler
+from torch.utils.data import DataLoader
+from torch.utils.data import sampler
+
+import torchvision
+from torchvision import datasets, models, transforms
+
+
+dtype = torch.float16
+
+
+if torch.cuda.is_available():
+    device = torch.device('cuda')
+else:
+    device = torch.device('cpu')
+print('using device:', device)
+
+data_dir = '../input/kermany2018/oct2017/OCT2017 '
+TRAIN = 'train'
+VAL = 'val'
+TEST = 'test'
+
+# VGG-16 Takes 224x224 images as input, so we resize all of them
+data_transforms = {
+    TRAIN: transforms.Compose([
+        # Data augmentation is a good practice for the train set
+        # Here, we randomly crop the image to 224x224 and
+        # randomly flip it horizontally.
+        transforms.RandomResizedCrop(224),
+        transforms.RandomHorizontalFlip(),
+        transforms.ToTensor(),
+    ]),
+    VAL: transforms.Compose([
+        transforms.Resize(256),
+        transforms.CenterCrop(224),
+        transforms.ToTensor(),
+    ]),
+    TEST: transforms.Compose([
+        transforms.Resize(256),
+        transforms.CenterCrop(224),
+        transforms.ToTensor(),
+    ])
+}
+
+image_datasets = {
+    x: datasets.ImageFolder(
+        os.path.join(data_dir, x),
+        transform=data_transforms[x]
+    )
+    for x in [TRAIN, VAL, TEST]
+}
+
+dataloaders = {
+    x: torch.utils.data.DataLoader(
+        image_datasets[x], batch_size=8,
+        shuffle=True, num_workers=4
+    )
+    for x in [TRAIN, VAL, TEST]
+}
+
+dataset_sizes = {x: len(image_datasets[x]) for x in [TRAIN, VAL, TEST]}
+
+for x in [TRAIN, VAL, TEST]:
+    print("Loaded {} images under {}".format(dataset_sizes[x], x))
+
+print("Classes: ")
+class_names = image_datasets[TRAIN].classes
+print(image_datasets[TRAIN].classes)
+
+
+
+
+

### 2019-11-28 src/models/sklearn/svm.py
Fixing up dataset loaders for better readiness to preprocess

@@ -5,26 +5,26 @@ import pandas as pd
 from sklearn.metrics import precision_recall_fscore_support
 from sklearn.svm import SVC
 
-from bids import BIDSLayout
-
-from src.data.ml import get_dset
+from src.data.ml import Dataset
 
 
-def test_svm(layout: BIDSLayout = None, limit: int = 300):
+def test_svm(limit: int = 300):
+    dataset = Dataset(limit=5, splits=(0.8, 0.2))
+    print(dataset)
 
-    ml_dataset = get_dset(layout=layout)
-    print("Data gathered")
-    print(ml_dataset)
-    ml_dataset.normalize()
-    ml_dataset.flatten()
-    print(ml_dataset)
+    X_train, y_train = dataset.get_train(flat=True, numpy=True)
+    X_val, y_val = dataset.get_val(flat=True, numpy=True)
 
-    svc = SVC(kernel='linear', C=1.)
+    svc = SVC(
+        kernel='linear',
+        C=1.,
+    )
+
     print("Training SVC model on {} scans".format(limit))
-    svc.fit(ml_dataset.X_train[:limit], ml_dataset.y_train[:limit])
+    svc.fit(X_train[:limit], y_train[:limit])
 
     print("Predicting using SVC model")
-    metrics = precision_recall_fscore_support(ml_dataset.y_val, svc.predict(ml_dataset.X_val))
+    metrics = precision_recall_fscore_support(y_val, svc.predict(X_val))
     results = pd.DataFrame(
         data=metrics,
         columns=['food', 'nonfood'],

### 2019-11-28 src/__main__.py
High scoring MLP, need to figure out proper integration of normalization!

@@ -32,25 +32,20 @@ def run():
 
         from src.data.ml import Dataset
 
-        Dataset()
-
-        # x, y = next(iter(l_train))
-        # print(x[0, 8:14, 20:32, 20:32])
-        # import numpy as np
-        # dimensions = np.prod(X.shape[1:])
+        dataset = Dataset()
 
         # for lr in [1e-1, 1e-2, 1e-3, 1e-4]:
         #     print("\n\nLR: {}".format(lr))
         lr = 1e-2
 
-        # fc_model = fc.model(dimensions)
-        fc_model = fc.Net()
+        fc_model = fc.model(dataset.dimensions)
+        # fc_model = fc.Net()
         fc_optimizer = fc.optimizer(fc_model, learning_rate=lr)
 
         train(
             fc_model,
             fc_optimizer,
-            # dset,
+            dataset,
             epochs=iargs.epochs,
         )
 

### 2019-11-28 src/data/ml.py

@@ -3,6 +3,7 @@ Data from our chosen experimental data in a readily consumed format for machine 
 """
 import numpy as np
 
+from sklearn.preprocessing import normalize
 import torch
 from torch.nn import functional as F
 from torch.utils.data import DataLoader, TensorDataset
@@ -17,12 +18,14 @@ BATCH_SIZE = 64
 class Dataset:
     def __init__(self, limit=3, splits=TVT_SPLITS):
         X, y = self._get_experimental_data(limit=limit)
-        dataset = self._clean(X, y)
+        dataset = self._transform(X, y)
         train, val, test = self._split(dataset, splits)
 
         self._train = train
         self._val = val
         self._test = test
+
+        self._clean()
 
     @staticmethod
     def _get_split(split, flat=False, numpy=False):
@@ -48,6 +51,10 @@ class Dataset:
     def get_test(self, *args, **kwargs):
         return self._get_split(self._test, *args, **kwargs)
     test = property(get_test)
+
+    @property
+    def dimensions(self):
+        return np.prod(list(self.train[0].shape)[1:])
 
     def __repr__(self):
         s_train = list(self.train[0].shape)
@@ -108,14 +115,7 @@ class Dataset:
         return scans.T, labels.T
 
     @staticmethod
-    def _clean(X, y):
-        # Normalize
-        X = (X - X.mean()) / X.std()
-
-        # TODO
-        #   self.mean_image = np.mean(self.X_train, axis=0)
-        #   subtract mean image
-
+    def _transform(X, y):
         # Load in data, add a fake channel dimension to allow interpolation
         data = torch.from_numpy(X.reshape(X.shape[0], 1, *X.shape[1:]))
         labels = torch.from_numpy(y)
@@ -145,6 +145,21 @@ class Dataset:
         return torch.utils.data.random_split(
             dataset,
             [train_size, val_size, test_size])
+
+    def _clean(self):
+        pass
+    #     train =
+    #
+    #     # Normalize
+    #     # TODO normalize per image? or accross everything?
+    #     # X = (X - X.mean()) / X.std()
+    #
+    #     mean_image = np.mean(X_train, axis=0)
+    #     X_train -= mean_image
+    #     X_val -= mean_image
+    #
+    #     X_train = normalize(X_train)
+    #     X_val = normalize(X_val)
 
     def get_loaders(self) -> (DataLoader, DataLoader, DataLoader):
         return [

### 2019-11-28 src/models/sklearn/mlp.py

@@ -8,19 +8,32 @@ from sklearn.neural_network import MLPClassifier
 from src.data.ml import Dataset
 
 
-def test_mlp(limit: int = 300):
-    dataset = Dataset(limit=5, splits=(0.8, 0.2))
+def test_mlp(limit: int = 1000):
+    dataset = Dataset(limit=3, splits=(0.8, 0.2))
     print(dataset)
 
     X_train, y_train = dataset.get_train(flat=True, numpy=True)
     X_val, y_val = dataset.get_val(flat=True, numpy=True)
 
+    # Normalize
+    # TODO normalize per image? or accross everything?
+    # X = (X - X.mean()) / X.std()
+
+    import numpy as np
+    mean_image = np.mean(X_train, axis=0)
+    X_train -= mean_image
+    X_val -= mean_image
+
+    from sklearn.preprocessing import normalize
+    X_train = normalize(X_train)
+    X_val = normalize(X_val)
+
     classifier = MLPClassifier(
         solver='adam',
-        hidden_layer_sizes=(100, 100),
+        hidden_layer_sizes=(200, 200, 200),
         random_state=1,
     )
-    
+
     print("Training MLPClassifier model on {} scans".format(limit))
     classifier.fit(X_train[:limit], y_train[:limit])
 

### 2019-11-28 src/utils/training.py

@@ -5,6 +5,7 @@ import torch
 from torch.nn import functional as F
 
 from src.env import device, dtype
+from src.data.ml import Dataset
 
 
 def accuracy(l, model):
@@ -23,8 +24,10 @@ def accuracy(l, model):
         print("{:d} / {:d} correct (%{:.2f})".format(num_correct, num_samples, 100 * acc))
 
 
-def train(model, optimizer, l_train, l_validate, epochs=1):
+def train(model, optimizer, dataset: Dataset, epochs=1):
     model = model.to(device=device)
+
+    l_train, l_validate, _ = dataset.get_loaders()
 
     for e in range(epochs):
         print("\nEpoch: {}, with {} batches".format(e, len(l_train)))

### 2019-11-29 src/data/ml.py
Fixing dataset to standardize and subtract mean image, seems very helpful, hoping I'm ready for fc network to learn

@@ -19,17 +19,16 @@ class Dataset:
     def __init__(self, limit=3, splits=TVT_SPLITS):
         X, y = self._get_experimental_data(limit=limit)
         dataset = self._transform(X, y)
-        train, val, test = self._split(dataset, splits)
+        splits = self._split(dataset, splits)
+        train, val, test = self._clean(*splits)
 
         self._train = train
         self._val = val
         self._test = test
 
-        self._clean()
-
     @staticmethod
     def _get_split(split, flat=False, numpy=False):
-        data, labels = split.dataset[split.indices]
+        data, labels = split.tensors
 
         if flat:
             data = data.view(data.shape[0], -1)
@@ -120,6 +119,9 @@ class Dataset:
         data = torch.from_numpy(X.reshape(X.shape[0], 1, *X.shape[1:]))
         labels = torch.from_numpy(y)
 
+        # Standardize data, 0 - 1
+        data = (data - data.mean()) / (data.max() - data.min())
+
         # Data is quite large, downsampling is fairly important to survive
         #  mini-batch x channels x [optional depth] x [optional height] x width.
         #  orientation doesn't matter
@@ -146,20 +148,23 @@ class Dataset:
             dataset,
             [train_size, val_size, test_size])
 
-    def _clean(self):
-        pass
-    #     train =
-    #
-    #     # Normalize
-    #     # TODO normalize per image? or accross everything?
-    #     # X = (X - X.mean()) / X.std()
-    #
-    #     mean_image = np.mean(X_train, axis=0)
-    #     X_train -= mean_image
-    #     X_val -= mean_image
-    #
-    #     X_train = normalize(X_train)
-    #     X_val = normalize(X_val)
+    @staticmethod
+    def _clean(train, val, test):
+        X_train, y_train = train.dataset[train.indices]
+        X_val, y_val = val.dataset[val.indices]
+        X_test, y_test = test.dataset[test.indices]
+
+        # Take out the mean training image from all splits
+        mean_image = X_train.mean(axis=0)
+        X_train -= mean_image
+        X_val -= mean_image
+        X_test -= mean_image
+
+        return (
+            TensorDataset(X_train, y_train),
+            TensorDataset(X_val, y_val),
+            TensorDataset(X_test, y_test),
+        )
 
     def get_loaders(self) -> (DataLoader, DataLoader, DataLoader):
         return [

### 2019-11-29 src/models/sklearn/mlp.py

@@ -15,19 +15,6 @@ def test_mlp(limit: int = 1000):
     X_train, y_train = dataset.get_train(flat=True, numpy=True)
     X_val, y_val = dataset.get_val(flat=True, numpy=True)
 
-    # Normalize
-    # TODO normalize per image? or accross everything?
-    # X = (X - X.mean()) / X.std()
-
-    import numpy as np
-    mean_image = np.mean(X_train, axis=0)
-    X_train -= mean_image
-    X_val -= mean_image
-
-    from sklearn.preprocessing import normalize
-    X_train = normalize(X_train)
-    X_val = normalize(X_val)
-
     classifier = MLPClassifier(
         solver='adam',
         hidden_layer_sizes=(200, 200, 200),

### 2019-11-29 src/__main__.py
Torch models still not learning, frustrating, exploring misuse of gradients

@@ -34,20 +34,19 @@ def run():
 
         dataset = Dataset()
 
-        # for lr in [1e-1, 1e-2, 1e-3, 1e-4]:
-        #     print("\n\nLR: {}".format(lr))
-        lr = 1e-2
+        for lr in [1e-1, 1e-2, 1e-3, 1e-4]:
+            print("\n\nLR: {}".format(lr))
 
-        fc_model = fc.model(dataset.dimensions)
-        # fc_model = fc.Net()
-        fc_optimizer = fc.optimizer(fc_model, learning_rate=lr)
+            fc_model = fc.model(dataset.dimensions)
+            # fc_model = fc.Net()
+            fc_optimizer = fc.optimizer(fc_model, learning_rate=lr)
 
-        train(
-            fc_model,
-            fc_optimizer,
-            dataset,
-            epochs=iargs.epochs,
-        )
+            train(
+                fc_model,
+                fc_optimizer,
+                dataset,
+                epochs=iargs.epochs,
+            )
 
 
 if __name__ == "__main__":

### 2019-11-29 src/data/ml.py

@@ -3,8 +3,8 @@ Data from our chosen experimental data in a readily consumed format for machine 
 """
 import numpy as np
 
-from sklearn.preprocessing import normalize
 import torch
+from torch.autograd import Variable
 from torch.nn import functional as F
 from torch.utils.data import DataLoader, TensorDataset
 

### 2019-11-29 src/models/fc.py

@@ -12,23 +12,23 @@ import torch.optim as optim
 from src.defs.layers import flatten, Flatten
 
 
-def model(input_dimension, hidden_layer_size: int = 6000):
+def model(idim, hdim: int = 6000):
     return nn.Sequential(
 
         Flatten(),
 
-        nn.Linear(input_dimension, hidden_layer_size),
+        nn.Linear(idim, hdim),
 
-        # nn.ReLU(),
-        nn.Tanh(),
+        nn.ReLU(),
+        # nn.Tanh(),
 
         # nn.Dropout(0.2),
         #
-        # nn.Linear(hidden_layer_size, hidden_layer_size),
+        # nn.Linear(hdim, hdim),
         #
         # nn.Tanh(),
 
-        nn.Linear(hidden_layer_size, 1),
+        nn.Linear(hdim, 1),
 
         nn.Sigmoid(),
 

### 2019-11-29 src/utils/training.py

@@ -2,6 +2,7 @@
 Neural Network training and evaluation
 """
 import torch
+from torch.autograd import Variable
 from torch.nn import functional as F
 
 from src.env import device, dtype
@@ -34,16 +35,18 @@ def train(model, optimizer, dataset: Dataset, epochs=1):
 
         for t, (x, y) in enumerate(l_train):
             model.train()
-            x = x.to(device=device, dtype=dtype)
-            y = y.to(device=device, dtype=torch.long)
 
-            scores = model.forward(x)
+            x = Variable(x, requires_grad=True)
+
+            x = x.to(device=device, dtype=dtype)
+            y = y.to(device=device, dtype=dtype)  #dtype=torch.long)
+
+            scores = model(x).squeeze()
+            # scores = model.forward(x)
             # print("ALL ONES: ", torch.all(scores.eq(1.0)))
 
-            from torch import nn
-            # loss = nn.BCELoss(scores, y)
-            loss = F.cross_entropy(scores, y)
-            # loss = F.binary_cross_entropy(scores, y)
+            # loss = F.cross_entropy(scores, y)
+            loss = F.binary_cross_entropy(scores, y)
 
             optimizer.zero_grad()
             loss.backward()

### 2019-11-29 src/data/ml.py
Fixing SVM performance for sanity

@@ -16,9 +16,9 @@ BATCH_SIZE = 64
 
 
 class Dataset:
-    def __init__(self, limit=3, splits=TVT_SPLITS):
+    def __init__(self, standardize=True, limit=3, splits=TVT_SPLITS):
         X, y = self._get_experimental_data(limit=limit)
-        dataset = self._transform(X, y)
+        dataset = self._transform(X, y, standardize=standardize)
         splits = self._split(dataset, splits)
         train, val, test = self._clean(*splits)
 
@@ -114,13 +114,14 @@ class Dataset:
         return scans.T, labels.T
 
     @staticmethod
-    def _transform(X, y):
+    def _transform(X, y, standardize=True):
         # Load in data, add a fake channel dimension to allow interpolation
         data = torch.from_numpy(X.reshape(X.shape[0], 1, *X.shape[1:]))
         labels = torch.from_numpy(y)
 
-        # Standardize data, 0 - 1
-        data = (data - data.mean()) / (data.max() - data.min())
+        if standardize:
+            # Standardize data, 0 - 1
+            data = (data - data.mean()) / (data.max() - data.min())
 
         # Data is quite large, downsampling is fairly important to survive
         #  mini-batch x channels x [optional depth] x [optional height] x width.

### 2019-11-29 src/models/sklearn/svm.py

@@ -9,7 +9,7 @@ from src.data.ml import Dataset
 
 
 def test_svm(limit: int = 300):
-    dataset = Dataset(limit=5, splits=(0.8, 0.2))
+    dataset = Dataset(standardize=False, limit=5, splits=(0.8, 0.2))
     print(dataset)
 
     X_train, y_train = dataset.get_train(flat=True, numpy=True)

### 2019-11-29 src/models/fc.py
Argh

@@ -12,26 +12,22 @@ import torch.optim as optim
 from src.defs.layers import flatten, Flatten
 
 
-def model(idim, hdim: int = 6000):
+def model(idim, hdim: int = 300):
     return nn.Sequential(
 
         Flatten(),
 
         nn.Linear(idim, hdim),
-
         nn.ReLU(),
-        # nn.Tanh(),
-
         # nn.Dropout(0.2),
-        #
-        # nn.Linear(hdim, hdim),
-        #
-        # nn.Tanh(),
 
-        nn.Linear(hdim, 1),
+        nn.Linear(hdim, hdim),
+        nn.ReLU(),
 
-        nn.Sigmoid(),
+        nn.Linear(hdim, 2),
 
+        # nn.Sigmoid(),
+        # nn.LogSigmoid(),
     )
 
 

### 2019-11-29 src/utils/training.py

@@ -39,14 +39,16 @@ def train(model, optimizer, dataset: Dataset, epochs=1):
             x = Variable(x, requires_grad=True)
 
             x = x.to(device=device, dtype=dtype)
-            y = y.to(device=device, dtype=dtype)  #dtype=torch.long)
+            y = y.to(device=device, dtype=torch.long)  # cross entropy
+            # y = y.to(device=device, dtype=dtype)  # binary cross entropy
 
-            scores = model(x).squeeze()
+            scores = model(x)#.squeeze()
+            print(scores.shape, y.shape)
             # scores = model.forward(x)
             # print("ALL ONES: ", torch.all(scores.eq(1.0)))
 
-            # loss = F.cross_entropy(scores, y)
-            loss = F.binary_cross_entropy(scores, y)
+            loss = F.cross_entropy(scores, y)
+            # loss = F.binary_cross_entropy(scores, y)
 
             optimizer.zero_grad()
             loss.backward()

### 2019-11-29 src/__main__.py
Finally got proof of learning with pytorch. christ...

@@ -34,16 +34,17 @@ def run():
 
         dataset = Dataset()
 
-        for lr in [1e-1, 1e-2, 1e-3, 1e-4]:
-            print("\n\nLR: {}".format(lr))
+        for lr in [1e-1]:  #, 1e-2, 1e-3, 1e-4]:
+            print("\nLR: {}".format(lr))
 
-            fc_model = fc.model(dataset.dimensions)
-            # fc_model = fc.Net()
+            # fc_model = fc.model(dataset.dimensions)
+            fc_model = fc.Net(dataset.dimensions)
             fc_optimizer = fc.optimizer(fc_model, learning_rate=lr)
 
             train(
                 fc_model,
                 fc_optimizer,
+                fc.criterion,
                 dataset,
                 epochs=iargs.epochs,
             )

### 2019-11-29 src/models/fc.py

@@ -44,32 +44,35 @@ def optimizer(model, learning_rate: float = 1e-2):
     #     lr=learning_rate)
 
 
-# # TODO net or model
-# class Net(nn.Module):
-#
-#     def __init__(self, idim: int, hdim: int = 6000):
-#         super(Net, self).__init__()
-#         self.fc1 = nn.Linear(idim, hdim)
-#         self.fc2 = nn.Linear(hdim, 2)
-#
-#     def forward(self, x):
-#         x = flatten(x)
-#         x = self.fc1(x)
-#         x = F.tanh(x)
-#         x = self.fc2(x)
-#
-#         return F.log_softmax(x)
-#
-#     def predict(self, x):
-#
-#         pred = F.softmax(self.forward(x))
-#         ans = []
-#
-#         # Pick the class with maximum weight
-#         for t in pred:
-#             if t[0] > t[1]:
-#                 ans.append(0)
-#             else:
-#                 ans.append(1)
-#
-#         return torch.tensor(ans)
+criterion = nn.NLLLoss()
+
+
+class Net(nn.Module):
+
+    def __init__(self, idim: int, hdim: int = 300):
+        super(Net, self).__init__()
+        self.fc1 = nn.Linear(idim, hdim)
+        self.fc2 = nn.Linear(hdim, hdim)
+        self.fc3 = nn.Linear(hdim, 2)
+
+    def forward(self, x):
+        x = flatten(x)
+        x = F.relu(self.fc1(x))
+        x = F.relu(self.fc2(x))
+        x = self.fc3(x)
+
+        return F.log_softmax(x)
+
+    def predict(self, x):
+
+        pred = F.softmax(self.forward(x))
+        ans = []
+
+        # Pick the class with maximum weight
+        for t in pred:
+            if t[0] > t[1]:
+                ans.append(0)
+            else:
+                ans.append(1)
+
+        return torch.tensor(ans)

### 2019-11-29 src/utils/training.py

@@ -2,7 +2,6 @@
 Neural Network training and evaluation
 """
 import torch
-from torch.autograd import Variable
 from torch.nn import functional as F
 
 from src.env import device, dtype
@@ -25,7 +24,7 @@ def accuracy(l, model):
         print("{:d} / {:d} correct (%{:.2f})".format(num_correct, num_samples, 100 * acc))
 
 
-def train(model, optimizer, dataset: Dataset, epochs=1):
+def train(model, optimizer, criterion, dataset: Dataset, epochs=1):
     model = model.to(device=device)
 
     l_train, l_validate, _ = dataset.get_loaders()
@@ -36,19 +35,22 @@ def train(model, optimizer, dataset: Dataset, epochs=1):
         for t, (x, y) in enumerate(l_train):
             model.train()
 
-            x = Variable(x, requires_grad=True)
-
+            # x = x.requires_grad_()
+            #
             x = x.to(device=device, dtype=dtype)
             y = y.to(device=device, dtype=torch.long)  # cross entropy
-            # y = y.to(device=device, dtype=dtype)  # binary cross entropy
+            # # y = y.to(device=device, dtype=dtype)  # binary cross entropy
+            #
+            # scores = model(x)#.squeeze()
+            # print(scores.shape, y.shape)
+            # # scores = model.forward(x)
+            # # print("ALL ONES: ", torch.all(scores.eq(1.0)))
+            #
+            # loss = F.cross_entropy(scores, y)
+            # # loss = F.binary_cross_entropy(scores, y)
 
-            scores = model(x)#.squeeze()
-            print(scores.shape, y.shape)
-            # scores = model.forward(x)
-            # print("ALL ONES: ", torch.all(scores.eq(1.0)))
-
-            loss = F.cross_entropy(scores, y)
-            # loss = F.binary_cross_entropy(scores, y)
+            scores = model(x)
+            loss = criterion(scores, y)
 
             optimizer.zero_grad()
             loss.backward()

### 2019-11-29 src/__main__.py
Working on 2d and 3d cnns, stuck seemingly

@@ -1,18 +1,8 @@
 """
-TODO
-
-Awesome!
-
-Some TODO's to work on in the future
-- More subjects (memory management)
-- Cut timesteps that could either belong to one event or another (boundary dispute) as I'm not 100% confident in event assignment
-
-- Deep learn, VGG
-- Find convolution on non flattened image (3d convolution could get brain regions better?)
-
-- Review HW for good ideas in training, data manipulation, etc
-
+Project entry point, primarily pick your desired model framework and launch training/evaluation
 """
+import os
+import pickle
 
 
 def run():
@@ -27,27 +17,84 @@ def run():
         print(mlp.test_mlp())
 
     elif iargs.model == "fc":
+        from src.data.ml import Dataset
         from src.models import fc
+        from src.utils import plotting
         from src.utils.training import train
 
+        dataset = Dataset(dimensions='3d')
+
+        model = fc.Net(dataset.dimensions)
+        optimizer = fc.optimizer(model, learning_rate=1e-1)
+        criterion = fc.criterion()
+
+        losses, accuracies = train(
+            model,
+            optimizer,
+            criterion,
+            dataset,
+            epochs=iargs.epochs,
+            print_frequency=iargs.print_freq,
+        )
+
+        if iargs.plot:
+            plotting.plot_loss(losses)
+            plotting.plot_accuracies(accuracies)
+
+    elif iargs.model in ["conv2d", "2d"]:
         from src.data.ml import Dataset
+        from src.models import conv2d
+        from src.utils import plotting
+        from src.utils.training import train
 
-        dataset = Dataset()
+        dataset = Dataset(dimensions='2d')
 
-        for lr in [1e-1]:  #, 1e-2, 1e-3, 1e-4]:
-            print("\nLR: {}".format(lr))
+        model = conv2d.model()
+        optimizer = conv2d.optimizer(model, learning_rate=1e-2)
+        criterion = conv2d.criterion()
 
-            # fc_model = fc.model(dataset.dimensions)
-            fc_model = fc.Net(dataset.dimensions)
-            fc_optimizer = fc.optimizer(fc_model, learning_rate=lr)
+        losses, accuracies = train(
+            model,
+            optimizer,
+            criterion,
+            dataset,
+            epochs=iargs.epochs,
+            print_frequency=iargs.print_freq,
+        )
 
-            train(
-                fc_model,
-                fc_optimizer,
-                fc.criterion,
[…]

### 2019-11-29 src/args.py

@@ -6,13 +6,23 @@ parser = argparse.ArgumentParser()
 parser.add_argument(
     "-m", "--model",
     nargs="?",
-    choices=["svm", "mlp", "fc"],
-    default="fc",
+    choices=["svm", "mlp", "fc", "2d", "conv2d", "3d", "conv3d"],
+    default="conv3d",
     type=str,
 )
 parser.add_argument(
     "-e", "--epochs",
     default=3,
+    type=int,
+)
+parser.add_argument(
+    "-p", "--plot",
+    action="store_true",
+    default=False,
+)
+parser.add_argument(
+    "--print-freq",
+    default=2,
     type=int,
 )
 parser.add_argument(

### 2019-11-29 src/data/ml.py

@@ -4,7 +4,6 @@ Data from our chosen experimental data in a readily consumed format for machine 
 import numpy as np
 
 import torch
-from torch.autograd import Variable
 from torch.nn import functional as F
 from torch.utils.data import DataLoader, TensorDataset
 
@@ -16,21 +15,22 @@ BATCH_SIZE = 64
 
 
 class Dataset:
-    def __init__(self, standardize=True, limit=3, splits=TVT_SPLITS):
+    def __init__(self, standardize=True, dimensions: str = "3d", scale=1.0, limit=3, splits=TVT_SPLITS):
+        self._d = dimensions
+
         X, y = self._get_experimental_data(limit=limit)
-        dataset = self._transform(X, y, standardize=standardize)
+        dataset = self._transform(X, y, standardize=standardize, scale=scale)
         splits = self._split(dataset, splits)
-        train, val, test = self._clean(*splits)
+        train, val, test = self._clean(*splits, dimensions=dimensions)
 
         self._train = train
         self._val = val
         self._test = test
 
-    @staticmethod
-    def _get_split(split, flat=False, numpy=False):
+    def _get_split(self, split, numpy=False):
         data, labels = split.tensors
 
-        if flat:
+        if self._d == "1d":
             data = data.view(data.shape[0], -1)
 
         if numpy:
@@ -114,7 +114,16 @@ class Dataset:
         return scans.T, labels.T
 
     @staticmethod
-    def _transform(X, y, standardize=True):
+    def _transform(X, y, standardize=True, scale=3./4):
+
+        # Cut down the data size, there's a lot of empty space in the scans
+        # this may be clumsy but really helps keep the size of data down
+        # retain all samples, vertical dimensions, cut the front to back & side to side
+        print("Shaving bulk of empty space out of scans")
+        print("Before shave: ", X.shape)
+        X = X[:, :, 10:54, 12:52]
+        print("After shave: ", X.shape)
+
         # Load in data, add a fake channel dimension to allow interpolation
         data = torch.from_numpy(X.reshape(X.shape[0], 1, *X.shape[1:]))
         labels = torch.from_numpy(y)
@@ -131,7 +140,7 @@ class Dataset:
             np.prod(data.shape[1:])))
         data = F.interpolate(
             data,
-            scale_factor=2./4.,
+            scale_factor=scale,
         )
         data = data.squeeze()
         print("After downsampling: shape={}, features={}".format(
[…]

### 2019-11-29 src/models/conv2d.py

@@ -1,0 +1,138 @@
+"""
+Explore performance of traditional 2d convolutional networks on a flattened view of the brain scans
+"""
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import torch.optim as optim
+import torchvision
+from torchvision.models import resnet
+
+
+def optimizer(model, learning_rate: float = 1e-2):
+    return optim.SGD(
+        model.parameters(),
+        lr=learning_rate,
+        momentum=0.9,
+        # nesterov=True,
+    )
+    # return optim.Adam(
+    #     model.parameters(),
+    #     lr=learning_rate,
+    # )
+
+
+def criterion():
+    return nn.CrossEntropyLoss()
+
+
+def model():
+    return SingleChannelResNet(
+        resnet.BasicBlock,  # block
+        [2, 2, 2, 2],  # layers
+        num_classes=2,
+        zero_init_residual=False,
+        groups=1,
+        width_per_group=64,
+        replace_stride_with_dilation=None,
+        norm_layer=None
+    )
+
+
+class SingleChannelResNet(nn.Module):
+
+    def __init__(self, block, layers, num_classes=1000, zero_init_residual=False,
+                 groups=1, width_per_group=64, replace_stride_with_dilation=None,
+                 norm_layer=None):
+        super(SingleChannelResNet, self).__init__()
+        if norm_layer is None:
+            norm_layer = nn.BatchNorm2d
+        self._norm_layer = norm_layer
+
+        self.inplanes = 64
+        self.dilation = 1
+        if replace_stride_with_dilation is None:
+            # each element in the tuple indicates if we should replace
+            # the 2x2 stride with a dilated convolution instead
+            replace_stride_with_dilation = [False, False, False]
+        if len(replace_stride_with_dilation) != 3:
+            raise ValueError("replace_stride_with_dilation should be None "
+                             "or a 3-element tuple, got {}".format(replace_stride_with_dilation))
+        self.groups = groups
+        self.base_width = width_per_group
+        self.conv1 = nn.Conv2d(1, self.inplanes, kernel_size=7, stride=2, padding=3,
+                               bias=False)
+        self.bn1 = norm_layer(self.inplanes)
+        self.relu = nn.ReLU(inplace=True)
+        self.maxpool = nn.MaxPool2d(kernel_size=3, stride=2, padding=1)
+        self.layer1 = self._make_layer(block, 64, layers[0])
+        self.layer2 = self._make_layer(block, 128, layers[1], stride=2,
[…]

### 2019-11-29 src/models/conv3d.py

@@ -5,389 +5,196 @@ import torch
 import torch.nn as nn
 import torch.nn.functional as F
 import torch.optim as optim
+import torchvision
+from torchvision.models import video
 
 from src.defs import weights
 
 
-#
-# learning_rate = 1e-2
-#
-# factor = 8
-# in_channel = 1
-# channel_1 = 4 * factor
-# channel_2 = 8 * factor
-# channel_3 = 8 * factor
-# channel_4 = 16 * factor
-#
-# # torch.nn.Conv3d(in_channels, out_channels, kernel_size, stride=1, padding=0, dilation=1, groups=1, bias=True, padding_mode='zeros')
-#
-# model = nn.Sequential(
-#     #     nn.Conv3d(in_channel, channel_1, kernel_size=7, stride=2, padding=3),
-#     nn.Conv3d(in_channel, channel_1, kernel_size=7, stride=2, padding=3),
-#     nn.ReLU(),
-#     nn.Conv3d(channel_1, channel_2, kernel_size=5, stride=1, padding=2),
-#     nn.ReLU(),
-#     nn.MaxPool3d(kernel_size=2),
-#     nn.Dropout(p=0.4),
-#
-#     nn.Conv3d(channel_2, channel_3, kernel_size=5, stride=2, padding=2),
-#     nn.ReLU(),
-#     nn.Conv2d(channel_3, channel_4, kernel_size=3, stride=1, padding=2),
-#     nn.ReLU(),
-#     nn.MaxPool2d(kernel_size=2),
-#     nn.Dropout(p=0.4),
-#
-#     Flatten(),
-#     nn.Linear(1152, 128),
-#     nn.ReLU(),
-#     nn.Linear(128, num_classes),
-# )
-#
-#
-# train_part34(model, optimizer, epochs=1)
-#
-#
-#
-# model = resnet10()
-#
-# optimizer = optim.SGD(
-#     model.parameters(),
-#     lr=learning_rate,
-#     momentum=0.9,
-#     nesterov=True,
-# )
-#
-# train_part34(model, optimizer, epochs=1)
-#
-#
-# def three_layer_convnet(x, params):
-#     """
-#     Performs the forward pass of a three-layer convolutional network with the
-#     architecture defined above.
-#
-#     Inputs:
-#     - x: A PyTorch Tensor of shape (N, 3, H, W) giving a minibatch of images
-#     - params: A list of PyTorch Tensors giving the weights and biases for the
-#       network; should contain the following:
[…]

### 2019-11-29 src/models/fc.py

@@ -12,23 +12,23 @@ import torch.optim as optim
 from src.defs.layers import flatten, Flatten
 
 
-def model(idim, hdim: int = 300):
-    return nn.Sequential(
-
-        Flatten(),
-
-        nn.Linear(idim, hdim),
-        nn.ReLU(),
-        # nn.Dropout(0.2),
-
-        nn.Linear(hdim, hdim),
-        nn.ReLU(),
-
-        nn.Linear(hdim, 2),
-
-        # nn.Sigmoid(),
-        # nn.LogSigmoid(),
-    )
+# def model(idim, hdim: int = 300):
+#     return nn.Sequential(
+#
+#         Flatten(),
+#
+#         nn.Linear(idim, hdim),
+#         nn.ReLU(),
+#         # nn.Dropout(0.2),
+#
+#         nn.Linear(hdim, hdim),
+#         nn.ReLU(),
+#
+#         nn.Linear(hdim, 2),
+#
+#         # nn.Sigmoid(),
+#         # nn.LogSigmoid(),
+#     )
 
 
 def optimizer(model, learning_rate: float = 1e-2):
@@ -39,12 +39,9 @@ def optimizer(model, learning_rate: float = 1e-2):
         # nesterov=True,
     )
 
-    # return optim.Adam(
-    #     model.parameters(),
-    #     lr=learning_rate)
 
-
-criterion = nn.NLLLoss()
+def criterion():
+    return nn.NLLLoss()
 
 
 class Net(nn.Module):
@@ -61,14 +58,13 @@ class Net(nn.Module):
         x = F.relu(self.fc2(x))
         x = self.fc3(x)
 
-        return F.log_softmax(x)
+        return F.log_softmax(x)#, dim=x.dim())
 
     def predict(self, x):
 
         pred = F.softmax(self.forward(x))
         ans = []
 
-        # Pick the class with maximum weight
         for t in pred:
             if t[0] > t[1]:
                 ans.append(0)

### 2019-11-29 src/models/resnet3d.py

@@ -1,0 +1,292 @@
+import torch
+import torch.nn as nn
+
+
+class Conv3DSimple(nn.Conv3d):
+    def __init__(self,
+                 in_planes,
+                 out_planes,
+                 midplanes=None,
+                 stride=1,
+                 padding=1):
+
+        super(Conv3DSimple, self).__init__(
+            in_channels=in_planes,
+            out_channels=out_planes,
+            kernel_size=(3, 3, 3),
+            stride=stride,
+            padding=padding,
+            bias=False)
+
+    @staticmethod
+    def get_downsample_stride(stride):
+        return (stride, stride, stride)
+
+
+class Conv2Plus1D(nn.Sequential):
+
+    def __init__(self,
+                 in_planes,
+                 out_planes,
+                 midplanes,
+                 stride=1,
+                 padding=1):
+        super(Conv2Plus1D, self).__init__(
+            nn.Conv3d(in_planes, midplanes, kernel_size=(1, 3, 3),
+                      stride=(1, stride, stride), padding=(0, padding, padding),
+                      bias=False),
+            nn.BatchNorm3d(midplanes),
+            nn.ReLU(inplace=True),
+            nn.Conv3d(midplanes, out_planes, kernel_size=(3, 1, 1),
+                      stride=(stride, 1, 1), padding=(padding, 0, 0),
+                      bias=False))
+
+    @staticmethod
+    def get_downsample_stride(stride):
+        return (stride, stride, stride)
+
+
+class Conv3DNoTemporal(nn.Conv3d):
+
+    def __init__(self,
+                 in_planes,
+                 out_planes,
+                 midplanes=None,
+                 stride=1,
+                 padding=1):
+
+        super(Conv3DNoTemporal, self).__init__(
+            in_channels=in_planes,
+            out_channels=out_planes,
+            kernel_size=(1, 3, 3),
+            stride=(1, stride, stride),
+            padding=(0, padding, padding),
+            bias=False)
+
+    @staticmethod
+    def get_downsample_stride(stride):
+        return (1, stride, stride)
+
+
+class BasicBlock(nn.Module):
+
+    expansion = 1
+
+    def __init__(self, inplanes, planes, conv_builder, stride=1, downsample=None):
+        midplanes = (inplanes * planes * 3 * 3 * 3) // (inplanes * 3 * 3 + 3 * planes)
+
[…]

### 2019-11-29 src/models/sklearn/mlp.py

@@ -9,11 +9,11 @@ from src.data.ml import Dataset
 
 
 def test_mlp(limit: int = 1000):
-    dataset = Dataset(limit=3, splits=(0.8, 0.2))
+    dataset = Dataset(dimensions="1d", limit=3, splits=(0.8, 0.2))
     print(dataset)
 
-    X_train, y_train = dataset.get_train(flat=True, numpy=True)
-    X_val, y_val = dataset.get_val(flat=True, numpy=True)
+    X_train, y_train = dataset.get_train(numpy=True)
+    X_val, y_val = dataset.get_val(numpy=True)
 
     classifier = MLPClassifier(
         solver='adam',

### 2019-11-29 src/models/sklearn/svm.py

@@ -9,11 +9,11 @@ from src.data.ml import Dataset
 
 
 def test_svm(limit: int = 300):
-    dataset = Dataset(standardize=False, limit=5, splits=(0.8, 0.2))
+    dataset = Dataset(standardize=False, dimensions="1d", limit=5, splits=(0.8, 0.2))
     print(dataset)
 
-    X_train, y_train = dataset.get_train(flat=True, numpy=True)
-    X_val, y_val = dataset.get_val(flat=True, numpy=True)
+    X_train, y_train = dataset.get_train(numpy=True)
+    X_val, y_val = dataset.get_val(numpy=True)
 
     svc = SVC(
         kernel='linear',

### 2019-11-29 src/utils/hyper.py

@@ -1,0 +1,36 @@
+"""
+Neural Network training and evaluation
+"""
+from src.data.ml import Dataset
+from src.utils.training import train as vanilla_train
+
+
+LEARNING_RATES = [1e-1]
+
+
+class Hyperparameters:
+
+    def __init__(
+            self,
+            epochs: int = 2,
+            learning_rates: list = LEARNING_RATES,
+    ):
+        self.learning_rates = learning_rates
+        self.epochs = epochs
+
+
+def train(model_fn, optimizer_fn, criterion_fn, dataset: Dataset, h: Hyperparameters):
+
+    for lr in h.learning_rates:
+        print("\nLR: {}".format(lr))
+
+        model = model_fn(dataset.dimensions)
+        optimizer = optimizer_fn(model, learning_rate=lr)
+
+        vanilla_train(
+            model,
+            optimizer,
+            criterion_fn,
+            dataset,
+            epochs=h.epochs,
+        )

### 2019-11-29 src/utils/plotting.py

@@ -1,0 +1,56 @@
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+
+def plot_loss(losses: list):
+    ax = sns.lineplot(range(len(losses)), losses)
+    plt.show()
+
+
+def plot_accuracies(accuracies: list):
+    ax = sns.lineplot(range(len(accuracies)), accuracies)
+    plt.show()
+
+
+# from src.data.ml import Dataset
+# from src.utils.plotting import show_slices
+# d = Dataset()
+# t,l = d.train
+# s = t[0]
+# show_slices(s[15, :, :], s[:, 28, :], s[:, :, 32])
+#
+# avg_scan = s.mean(dim=0)
+# # plt.imshow(  tensor_image.permute(1, 2, 0)  )
+# # plt.imshow(avg_scan.view(1, 64, 64))
+
+"""
+s.shape
+from src.data.ml import Dataset
+from src.utils.plotting import show_slices
+d = Dataset(dimensions="2d",scale=5./6)
+t,l = d.train
+s = t[0]
+show_slices([s[15, :, :], s[:, 20, :], s[:, :, 24]])
+s.shape
+from src.data.ml import Dataset
+from src.utils.plotting import show_slices
+d = Dataset(dimensions="2d")
+t,l = d.train
+s = t[0]
+show_slices([s[:, 16, :], s[:, :, 16]])
+s.shape
+from matplotlib import pyplot as plt
+plt.imshow(s.T, cmap="gray")
+
+# SHOW THE AVERAGE
+plt.imshow(s[0,:,:].T, cmap="gray")
+30 * 44 * 40
+"""
+
+def show_slices(slices):
+    """ Function to display row of image slices """
+    fig, axes = plt.subplots(1, len(slices))
+    for i, slice in enumerate(slices):
+        axes[i].imshow(slice.T, cmap="gray", origin="lower")
+
+    plt.suptitle("Center slices for EPI image")

### 2019-11-29 src/utils/training.py

@@ -1,6 +1,8 @@
 """
 Neural Network training and evaluation
 """
+from sklearn import metrics
+
 import torch
 from torch.nn import functional as F
 
@@ -8,7 +10,7 @@ from src.env import device, dtype
 from src.data.ml import Dataset
 
 
-def accuracy(l, model):
+def evaluate(l, model):
     num_correct = 0
     num_samples = 0
     model.eval()  # set model to evaluation mode
@@ -23,40 +25,66 @@ def accuracy(l, model):
         acc = float(num_correct) / num_samples
         print("{:d} / {:d} correct (%{:.2f})".format(num_correct, num_samples, 100 * acc))
 
+    return acc
+    # roc_auc = metrics.roc_auc_score
+    # roc_auc(y, pred)
 
-def train(model, optimizer, criterion, dataset: Dataset, epochs=1):
+
+def train(
+        model,
+        optimizer,
+        criterion,
+        dataset: Dataset,
+        epochs=1,
+        print_frequency: int = 1):
+
     model = model.to(device=device)
 
     l_train, l_validate, _ = dataset.get_loaders()
+    num_batches = len(l_train)
+
+    print("Training with {} batches".format(num_batches))
+    losses, accuracies = [], []
 
     for e in range(epochs):
-        print("\nEpoch: {}, with {} batches".format(e, len(l_train)))
+        print("\nEpoch: {}".format(e + 1))
 
-        for t, (x, y) in enumerate(l_train):
+        round_loss = 0.0
+        for i, (x, y) in enumerate(l_train):
+
+            # Prepare for training
             model.train()
-
             # x = x.requires_grad_()
-            #
             x = x.to(device=device, dtype=dtype)
-            y = y.to(device=device, dtype=torch.long)  # cross entropy
-            # # y = y.to(device=device, dtype=dtype)  # binary cross entropy
-            #
-            # scores = model(x)#.squeeze()
-            # print(scores.shape, y.shape)
-            # # scores = model.forward(x)
-            # # print("ALL ONES: ", torch.all(scores.eq(1.0)))
-            #
-            # loss = F.cross_entropy(scores, y)
-            # # loss = F.binary_cross_entropy(scores, y)
+            y = y.to(device=device, dtype=torch.long)
 
+            # Forward pass
             scores = model(x)
+            # print(scores, y)
+            # print("ALL ONES: ", torch.all(scores.eq(1.0)))
+
             loss = criterion(scores, y)
 
+            # Save info for plotting
+            round_loss += loss.item()
[…]

### 2019-11-29 src/__main__.py
Fixing up hyperparameters to try to fix conv2d

@@ -44,27 +44,20 @@ def run():
     elif iargs.model in ["conv2d", "2d"]:
         from src.data.ml import Dataset
         from src.models import conv2d
-        from src.utils import plotting
-        from src.utils.training import train
+        from src.utils import hyper, plotting
 
         dataset = Dataset(dimensions='2d')
 
-        model = conv2d.model()
-        optimizer = conv2d.optimizer(model, learning_rate=1e-2)
-        criterion = conv2d.criterion()
-
-        losses, accuracies = train(
-            model,
-            optimizer,
-            criterion,
+        hyper.train(
+            conv2d.model,
+            conv2d.optimizer,
+            conv2d.criterion,
             dataset,
-            epochs=iargs.epochs,
-            print_frequency=iargs.print_freq,
+            hyper.Hyperparameters(
+                epochs=iargs.epochs,
+                learning_rates=[1e-1, 1e-2, 1e-3, 1e-4],
+            )
         )
-
-        if iargs.plot:
-            plotting.plot_loss(losses)
-            plotting.plot_accuracies(accuracies)
 
     elif iargs.model in ["conv3d", "3d"]:
         from src.data.ml import Dataset

### 2019-11-29 src/models/conv2d.py

@@ -26,7 +26,7 @@ def criterion():
     return nn.CrossEntropyLoss()
 
 
-def model():
+def model(*args, **kwargs):
     return SingleChannelResNet(
         resnet.BasicBlock,  # block
         [2, 2, 2, 2],  # layers

### 2019-11-29 src/models/conv3d.py

@@ -11,20 +11,12 @@ from torchvision.models import video
 from src.defs import weights
 
 
-# torch.nn.Conv3d(in_channels, out_channels, kernel_size, stride=1, padding=0, dilation=1, groups=1, bias=True, padding_mode='zeros')
-
-
 def optimizer(model, learning_rate: float = 1e-2):
     return optim.SGD(
         model.parameters(),
         lr=learning_rate,
         momentum=0.9,
-        # nesterov=True,
     )
-    # return optim.Adam(
-    #     model.parameters(),
-    #     lr=learning_rate,
-    # )
 
 
 def criterion():
@@ -43,19 +35,6 @@ class SingleChannelStem(nn.Sequential):
 
 
 def resnet3d():
-    # def __init__(self, block, conv_makers, layers,
-    #              stem, num_classes=400,
-    #              zero_init_residual=False):
-    #     """Generic resnet video generator.
-    #
-    #     Args:
-    #         block (nn.Module): resnet building block
-    #         conv_makers (list(functions)): generator function for each layer
-    #         layers (List[int]): number of blocks per layer
-    #         stem (nn.Module, optional): Resnet stem, if None, defaults to conv-bn-relu. Defaults to None.
-    #         num_classes (int, optional): Dimension of the final FC layer. Defaults to 400.
-    #         zero_init_residual (bool, optional): Zero init bottleneck residual BN. Defaults to False.
-    #     """
     pretrained = False
     progress = True
 
@@ -65,10 +44,6 @@ def resnet3d():
          conv_makers=[video.Conv3DSimple] * 4,
          layers=[2, 2, 2, 2],
          stem=SingleChannelStem)
-    return video.r3d_18(
-        pretrained=False,
-        progress=True,
-        stem=SingleChannelStem)
 
 
 class Net(nn.Module):

### 2019-11-29 src/models/nilearn.py

@@ -1,32 +1,0 @@
-"""
-TODO use nilearn?
-"""
-
-from nilearn.datasets import fetch_haxby
-data_files = fetch_haxby()
-
-# Load behavioral data
-import pandas as pd
-behavioral = pd.read_csv(data_files.session_target[0], sep=" ")
-
-# Restrict to face and house conditions
-conditions = behavioral['labels']
-condition_mask = conditions.isin(['face', 'house'])
-
-# Split data into train and test samples, using the chunks
-condition_mask_train = (condition_mask) & (behavioral['chunks'] <= 6)
-condition_mask_test = (condition_mask) & (behavioral['chunks'] > 6)
-
-# Apply this sample mask to X (fMRI data) and y (behavioral labels)
-# Because the data is in one single large 4D image, we need to use
-# index_img to do the split easily
-from nilearn.image import index_img
-func_filenames = data_files.func[0]
-X_train = index_img(func_filenames, condition_mask_train)
-X_test = index_img(func_filenames, condition_mask_test)
-y_train = conditions[condition_mask_train]
-y_test = conditions[condition_mask_test]
-
-# Compute the mean epi to be used for the background of the plotting
-from nilearn.image import mean_img
-background_img = mean_img(func_filenames)

### 2019-11-29 src/utils/hyper.py

@@ -26,11 +26,12 @@ def train(model_fn, optimizer_fn, criterion_fn, dataset: Dataset, h: Hyperparame
 
         model = model_fn(dataset.dimensions)
         optimizer = optimizer_fn(model, learning_rate=lr)
+        criterion = criterion_fn()
 
         vanilla_train(
             model,
             optimizer,
-            criterion_fn,
+            criterion,
             dataset,
             epochs=h.epochs,
         )

### 2019-11-30 src/__main__.py
Finally found success with convolutional neural networks! 2d learning, not stable and needs refinement, but proof finally!

@@ -17,47 +17,13 @@ def run():
         print(mlp.test_mlp())
 
     elif iargs.model == "fc":
-        from src.data.ml import Dataset
         from src.models import fc
-        from src.utils import plotting
-        from src.utils.training import train
-
-        dataset = Dataset(dimensions='3d')
-
-        model = fc.Net(dataset.dimensions)
-        optimizer = fc.optimizer(model, learning_rate=1e-1)
-        criterion = fc.criterion()
-
-        losses, accuracies = train(
-            model,
-            optimizer,
-            criterion,
-            dataset,
-            epochs=iargs.epochs,
-            print_frequency=iargs.print_freq,
-        )
-
-        if iargs.plot:
-            plotting.plot_loss(losses)
-            plotting.plot_accuracies(accuracies)
+        # epochs recommended above 80 - 100
+        fc.run(iargs)
 
     elif iargs.model in ["conv2d", "2d"]:
-        from src.data.ml import Dataset
         from src.models import conv2d
-        from src.utils import hyper, plotting
-
-        dataset = Dataset(dimensions='2d')
-
-        hyper.train(
-            conv2d.model,
-            conv2d.optimizer,
-            conv2d.criterion,
-            dataset,
-            hyper.Hyperparameters(
-                epochs=iargs.epochs,
-                learning_rates=[1e-1, 1e-2, 1e-3, 1e-4],
-            )
-        )
+        conv2d.run(iargs)
 
     elif iargs.model in ["conv3d", "3d"]:
         from src.data.ml import Dataset

### 2019-11-30 src/data/ml.py

@@ -170,6 +170,11 @@ class Dataset:
         X_val -= mean_image
         X_test -= mean_image
 
+        # # Standardize data, 0 - 1 (mean pushes neg, is it okay? lets readjust
+        # X_train = (X_train - X_train.mean()) / (X_train.max() - X_train.min())
+        # X_val = (X_val - X_val.mean()) / (X_val.max() - X_val.min())
+        # X_test = (X_test - X_test.mean()) / (X_test.max() - X_test.min())
+
         # Add an image channel for non 1d formats
         if dimensions in ["2d", "3d"]:
             print("Adding a channel dimension for convolutional networks")
@@ -177,12 +182,15 @@ class Dataset:
             X_val = X_val.view(X_val.shape[0], 1, *X_val.shape[1:])
             X_test = X_test.view(X_test.shape[0], 1, *X_test.shape[1:])
 
-        # Flatten axially for 2d, creating the average top-down view
+        # Flatten axially for 2d, choosing a manner of getting a horizontal slice
         if dimensions == "2d":
             print("Flattening axially for 2d view")
-            X_train = X_train.mean(dim=2)
-            X_val = X_val.mean(dim=2)
-            X_test = X_test.mean(dim=2)
+            # Taking the mean is worse than a single slice, I think it confuses info way too much
+            # X_train = X_train.mean(dim=2)
+            slice_height = X_train.shape[2] // 2
+            X_train = X_train[:, :, slice_height, :, :]
+            X_val = X_val[:, :, slice_height, :, :]
+            X_test = X_test[:, :, slice_height, :, :]
 
         return (
             TensorDataset(X_train, y_train),

### 2019-11-30 src/models/conv2d.py

@@ -8,13 +8,17 @@ import torch.optim as optim
 import torchvision
 from torchvision.models import resnet
 
+from src.data.ml import Dataset
+from src.models import conv2d
+from src.utils import hyper, plotting
+
 
 def optimizer(model, learning_rate: float = 1e-2):
     return optim.SGD(
         model.parameters(),
         lr=learning_rate,
         momentum=0.9,
-        # nesterov=True,
+        nesterov=True,
     )
     # return optim.Adam(
     #     model.parameters(),
@@ -27,7 +31,7 @@ def criterion():
 
 
 def model(*args, **kwargs):
-    return SingleChannelResNet(
+    model = SingleChannelResNet(
         resnet.BasicBlock,  # block
         [2, 2, 2, 2],  # layers
         num_classes=2,
@@ -37,6 +41,35 @@ def model(*args, **kwargs):
         replace_stride_with_dilation=None,
         norm_layer=None
     )
+
+    # Unfreeze / allow training (might not actually be frozen to start with though)
+    for params in model.parameters():
+        params.requires_grad = True
+
+    return model
+
+
+class Net(nn.Module):
+    def __init__(self, idims):
+        super(Net, self).__init__()
+        self.conv1 = nn.Conv2d(1, 6, 5)
+        self.pool = nn.MaxPool2d(2, 2)
+        self.conv2 = nn.Conv2d(6, 16, 5)
+        # self.fc1 = nn.Linear(16 * 5 * 5, 120)
+        self.fc1 = nn.Linear(57344, 120)
+        self.fc2 = nn.Linear(120, 84)
+        self.fc3 = nn.Linear(84, 2)
+
+    def forward(self, x):
+        x = self.pool(F.relu(self.conv1(x)))
+        x = self.pool(F.relu(self.conv2(x)))
+        print("FOFO", x.shape)
+        # x = x.view(-1, 16 * 5 * 5)
+        x = x.view(-1, 57344)
+        x = F.relu(self.fc1(x))
+        x = F.relu(self.fc2(x))
+        x = self.fc3(x)
+        return x
 
 
 class SingleChannelResNet(nn.Module):
@@ -136,3 +169,17 @@ class SingleChannelResNet(nn.Module):
     # Allow for accessing forward method in a inherited class
     forward = _forward
 
+
+def run(iargs):
+    dataset = Dataset(dimensions='2d')
+
+    hyper.train(
+        model,
+        optimizer,
+        criterion,
[…]

### 2019-11-30 src/models/fc.py

@@ -9,7 +9,10 @@ import torch.nn as nn
 import torch.nn.functional as F
 import torch.optim as optim
 
+from src.data.ml import Dataset
 from src.defs.layers import flatten, Flatten
+from src.utils import plotting
+from src.utils.training import train
 
 
 # def model(idim, hdim: int = 300):
@@ -36,7 +39,7 @@ def optimizer(model, learning_rate: float = 1e-2):
         model.parameters(),
         lr=learning_rate,
         momentum=0.9,
-        # nesterov=True,
+        nesterov=True,
     )
 
 
@@ -58,7 +61,7 @@ class Net(nn.Module):
         x = F.relu(self.fc2(x))
         x = self.fc3(x)
 
-        return F.log_softmax(x)#, dim=x.dim())
+        return F.log_softmax(x)
 
     def predict(self, x):
 
@@ -72,3 +75,22 @@ class Net(nn.Module):
                 ans.append(1)
 
         return torch.tensor(ans)
+
+
+def run(iargs):
+    dataset = Dataset(dimensions='3d')
+
+    model = Net(dataset.dimensions)
+
+    losses, accuracies = train(
+        model,
+        optimizer(model, learning_rate=1e-1),
+        criterion(),
+        dataset,
+        epochs=iargs.epochs,
+        print_frequency=iargs.print_freq,
+    )
+
+    if iargs.plot:
+        plotting.plot_loss(losses)
+        plotting.plot_accuracies(accuracies)

### 2019-11-30 src/models/sklearn/mlp.py

@@ -9,11 +9,14 @@ from src.data.ml import Dataset
 
 
 def test_mlp(limit: int = 1000):
-    dataset = Dataset(dimensions="1d", limit=3, splits=(0.8, 0.2))
+    dataset = Dataset(dimensions="2d", limit=3, splits=(0.8, 0.2))
     print(dataset)
 
     X_train, y_train = dataset.get_train(numpy=True)
     X_val, y_val = dataset.get_val(numpy=True)
+
+    X_train = X_train.reshape(X_train.shape[0], -1)
+    X_val = X_val.reshape(X_val.shape[0], -1)
 
     classifier = MLPClassifier(
         solver='adam',

### 2019-11-30 src/models/sklearn/svm.py

@@ -8,12 +8,17 @@ from sklearn.svm import SVC
 from src.data.ml import Dataset
 
 
-def test_svm(limit: int = 300):
-    dataset = Dataset(standardize=False, dimensions="1d", limit=5, splits=(0.8, 0.2))
+def test_svm(limit: int = 400):
+    # 1d is the standard way, seems a little better than the single slice 2d method
+    # but this is not a bad approach
+    dataset = Dataset(standardize=False, dimensions="2d", limit=5, splits=(0.7, 0.2))
     print(dataset)
 
     X_train, y_train = dataset.get_train(numpy=True)
     X_val, y_val = dataset.get_val(numpy=True)
+
+    X_train = X_train.reshape(X_train.shape[0], -1)
+    X_val = X_val.reshape(X_val.shape[0], -1)
 
     svc = SVC(
         kernel='linear',

### 2019-11-30 src/utils/plotting.py

@@ -1,5 +1,6 @@
-import seaborn as sns
 import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
 
 
 def plot_loss(losses: list):
@@ -45,6 +46,24 @@ plt.imshow(s.T, cmap="gray")
 # SHOW THE AVERAGE
 plt.imshow(s[0,:,:].T, cmap="gray")
 30 * 44 * 40
+
+
+def imshow(inp, title=None):
+    inp = inp.numpy().transpose((1, 2, 0))
+    # plt.figure(figsize=(10, 10))
+    plt.axis('off')
+    plt.imshow(inp)
+    if title is not None:
+        plt.title(title)
+    plt.pause(0.001)
+
+def show_databatch(inputs, classes):
+    out = torchvision.utils.make_grid(inputs)
+    imshow(out, title=[class_names[x] for x in classes])
+
+# Get a batch of training data
+inputs, classes = next(iter(dataloaders[TRAIN]))
+show_databatch(inputs, classes)
 """
 
 def show_slices(slices):
@@ -54,3 +73,49 @@ def show_slices(slices):
         axes[i].imshow(slice.T, cmap="gray", origin="lower")
 
     plt.suptitle("Center slices for EPI image")
+
+
+def show_experimental_scans():
+    from nilearn import plotting
+    from nilearn import image
+    from src.data import experiment
+
+    layout = experiment.get_food_temptation_data()
+
+    # subject data is S_ID, hi-res, exp-data, events
+    subject_data = experiment.get_all_subject_data(layout)
+
+    # take first subject, exp-data
+    scans = subject_data[0][2]
+
+    scans.orthoview()
+
+    show_variance(scans)
+
+    # pick a single scan to show
+    single_scan = scans.slicer[:, :, :, 0]
+
+    plotting.plot_stat_map(
+        single_scan,
+        # threshold=30,
+        # display_mode="z",
+        # cut_coords=1,
+        # colorbar=False,
+    )
+    plotting.plot_img(single_scan)
+    plotting.plot_epi(single_scan)
+    plotting.plot_roi(single_scan)
+    plotting.plot_anat(single_scan)
+
+
+def show_variance(scans):
+    d = scans.get_data()
+    # Where in the image is there actually any information for classification?
+
+    flat_d = d.reshape(-1, d.shape[-1])
[…]

### 2019-11-30 src/utils/training.py

@@ -76,11 +76,15 @@ def train(
             optimizer.step()
 
             # Log performance
-            if i % print_frequency == 0:
+            if False and i % print_frequency == 0:
                 print("[{:>3d}/{:>3d}] Loss: {:.4f}".format(
                     i + 1, num_batches,
                     loss.item(),
                 ))
+
+            # Try to help memory constraints out
+            del x, y, scores
+            torch.cuda.empty_cache()
 
         accuracy = evaluate(l_validate, model)
 

### 2019-11-30 src/__main__.py
Cleaning things up a little before the next push into cnn improvements

@@ -1,8 +1,6 @@
 """
 Project entry point, primarily pick your desired model framework and launch training/evaluation
 """
-import os
-import pickle
 
 
 def run():
@@ -26,34 +24,8 @@ def run():
         conv2d.run(iargs)
 
     elif iargs.model in ["conv3d", "3d"]:
-        from src.data.ml import Dataset
-        from src.models import conv3d, resnet3d
-        from src.utils.training import train
-
-        dset_path = "saved_dataset.pkl"
-        # if os.path.exists(dset_path):
-        #     with open(dset_path, 'rb') as f:
-        #         dataset = pickle.load(f)
-        # else:
-        dataset = Dataset(dimensions='3d', scale=1./2, limit=10)
-        with open(dset_path, 'wb') as f:
-            pickle.dump(dataset, f)
-
-        print(dataset.dimensions)
-
-        # model = conv3d.Net(dataset.dimensions)
-        model = resnet3d.r3d_18()
-        optimizer = conv3d.optimizer(model, learning_rate=1e-2)
-        criterion = conv3d.criterion()
-
-        train(
-            model,
-            optimizer,
-            criterion,
-            dataset,
-            epochs=iargs.epochs,
-            print_frequency=iargs.print_freq,
-        )
+        from src.models import conv3d
+        conv3d.run(iargs)
 
 
 if __name__ == "__main__":

### 2019-11-30 src/data/ml.py

@@ -72,10 +72,8 @@ class Dataset:
         """
         Get brain scans as labeled data, partitioned into Train, Validate, and Test
 
-        :param layout: optional, supply if you already have a layout on hand
-        :param limit: TODO I'm figuring out some memory constraints, the data if mishandled can exceed memory
+        :param limit: The data can quite easily exceed memory available
 
-        TODO how am I going to cope with the memory issue?
         64 * 64 * 30 = Per TimeStep (122,880)
 
         dtype is <f4, for little endian, 4 bit float (Did I see it was float32 at one point?)

### 2019-11-30 src/models/conv3d.py

@@ -1,14 +1,15 @@
 """
-TODO create model with 3d conv layers
+Explore performance of more novel 3d convolutional networks on 3d brain scans
 """
-import torch
+import os
+import pickle
+
 import torch.nn as nn
-import torch.nn.functional as F
 import torch.optim as optim
-import torchvision
 from torchvision.models import video
 
-from src.defs import weights
+from src.data.ml import Dataset
+from src.utils.training import train
 
 
 def optimizer(model, learning_rate: float = 1e-2):
@@ -173,3 +174,321 @@ class C3D(nn.Module):
         # self.apply(weights.random_weight())
         self.apply(weights_init)
 
+
+import torch
+import torch.nn as nn
+
+
+class Conv3DSimple(nn.Conv3d):
+    def __init__(self,
+                 in_planes,
+                 out_planes,
+                 midplanes=None,
+                 stride=1,
+                 padding=1):
+
+        super(Conv3DSimple, self).__init__(
+            in_channels=in_planes,
+            out_channels=out_planes,
+            kernel_size=(3, 3, 3),
+            stride=stride,
+            padding=padding,
+            bias=False)
+
+    @staticmethod
+    def get_downsample_stride(stride):
+        return (stride, stride, stride)
+
+
+class Conv2Plus1D(nn.Sequential):
+
+    def __init__(self,
+                 in_planes,
+                 out_planes,
+                 midplanes,
+                 stride=1,
+                 padding=1):
+        super(Conv2Plus1D, self).__init__(
+            nn.Conv3d(in_planes, midplanes, kernel_size=(1, 3, 3),
+                      stride=(1, stride, stride), padding=(0, padding, padding),
+                      bias=False),
+            nn.BatchNorm3d(midplanes),
+            nn.ReLU(inplace=True),
+            nn.Conv3d(midplanes, out_planes, kernel_size=(3, 1, 1),
+                      stride=(stride, 1, 1), padding=(padding, 0, 0),
+                      bias=False))
+
+    @staticmethod
+    def get_downsample_stride(stride):
+        return (stride, stride, stride)
+
+
+class Conv3DNoTemporal(nn.Conv3d):
+
+    def __init__(self,
+                 in_planes,
+                 out_planes,
+                 midplanes=None,
+                 stride=1,
+                 padding=1):
+
[…]

### 2019-11-30 src/models/resnet3d.py

@@ -1,292 +1,0 @@
-import torch
-import torch.nn as nn
-
-
-class Conv3DSimple(nn.Conv3d):
-    def __init__(self,
-                 in_planes,
-                 out_planes,
-                 midplanes=None,
-                 stride=1,
-                 padding=1):
-
-        super(Conv3DSimple, self).__init__(
-            in_channels=in_planes,
-            out_channels=out_planes,
-            kernel_size=(3, 3, 3),
-            stride=stride,
-            padding=padding,
-            bias=False)
-
-    @staticmethod
-    def get_downsample_stride(stride):
-        return (stride, stride, stride)
-
-
-class Conv2Plus1D(nn.Sequential):
-
-    def __init__(self,
-                 in_planes,
-                 out_planes,
-                 midplanes,
-                 stride=1,
-                 padding=1):
-        super(Conv2Plus1D, self).__init__(
-            nn.Conv3d(in_planes, midplanes, kernel_size=(1, 3, 3),
-                      stride=(1, stride, stride), padding=(0, padding, padding),
-                      bias=False),
-            nn.BatchNorm3d(midplanes),
-            nn.ReLU(inplace=True),
-            nn.Conv3d(midplanes, out_planes, kernel_size=(3, 1, 1),
-                      stride=(stride, 1, 1), padding=(padding, 0, 0),
-                      bias=False))
-
-    @staticmethod
-    def get_downsample_stride(stride):
-        return (stride, stride, stride)
-
-
-class Conv3DNoTemporal(nn.Conv3d):
-
-    def __init__(self,
-                 in_planes,
-                 out_planes,
-                 midplanes=None,
-                 stride=1,
-                 padding=1):
-
-        super(Conv3DNoTemporal, self).__init__(
-            in_channels=in_planes,
-            out_channels=out_planes,
-            kernel_size=(1, 3, 3),
-            stride=(1, stride, stride),
-            padding=(0, padding, padding),
-            bias=False)
-
-    @staticmethod
-    def get_downsample_stride(stride):
-        return (1, stride, stride)
-
-
-class BasicBlock(nn.Module):
-
-    expansion = 1
-
-    def __init__(self, inplanes, planes, conv_builder, stride=1, downsample=None):
-        midplanes = (inplanes * planes * 3 * 3 * 3) // (inplanes * 3 * 3 + 3 * planes)
-
[…]

### 2019-11-30 src/utils/plotting.py

@@ -13,71 +13,50 @@ def plot_accuracies(accuracies: list):
     plt.show()
 
 
-# from src.data.ml import Dataset
-# from src.utils.plotting import show_slices
-# d = Dataset()
-# t,l = d.train
-# s = t[0]
-# show_slices(s[15, :, :], s[:, 28, :], s[:, :, 32])
-#
-# avg_scan = s.mean(dim=0)
-# # plt.imshow(  tensor_image.permute(1, 2, 0)  )
-# # plt.imshow(avg_scan.view(1, 64, 64))
+def load_show_slices():
+    """
+    TODO
 
-"""
-s.shape
-from src.data.ml import Dataset
-from src.utils.plotting import show_slices
-d = Dataset(dimensions="2d",scale=5./6)
-t,l = d.train
-s = t[0]
-show_slices([s[15, :, :], s[:, 20, :], s[:, :, 24]])
-s.shape
-from src.data.ml import Dataset
-from src.utils.plotting import show_slices
-d = Dataset(dimensions="2d")
-t,l = d.train
-s = t[0]
-show_slices([s[:, 16, :], s[:, :, 16]])
-s.shape
-from matplotlib import pyplot as plt
-plt.imshow(s.T, cmap="gray")
+    def imshow(inp, title=None):
+        inp = inp.numpy().transpose((1, 2, 0))
+        # plt.figure(figsize=(10, 10))
+        plt.axis('off')
+        plt.imshow(inp)
+        if title is not None:
+            plt.title(title)
+        plt.pause(0.001)
 
-# SHOW THE AVERAGE
-plt.imshow(s[0,:,:].T, cmap="gray")
-30 * 44 * 40
+    def show_databatch(inputs, classes):
+        out = torchvision.utils.make_grid(inputs)
+        imshow(out, title=[class_names[x] for x in classes])
 
+    # Get a batch of training data
+    inputs, classes = next(iter(dataloaders[TRAIN]))
+    show_databatch(inputs, classes)
+    """
 
-def imshow(inp, title=None):
-    inp = inp.numpy().transpose((1, 2, 0))
-    # plt.figure(figsize=(10, 10))
-    plt.axis('off')
-    plt.imshow(inp)
-    if title is not None:
-        plt.title(title)
-    plt.pause(0.001)
+    from src.data.ml import Dataset
 
-def show_databatch(inputs, classes):
-    out = torchvision.utils.make_grid(inputs)
[…]

### 2019-11-30 src/models/conv2d.py
Removing circular import

@@ -9,7 +9,6 @@ import torchvision
 from torchvision.models import resnet
 
 from src.data.ml import Dataset
-from src.models import conv2d
 from src.utils import hyper, plotting
 
 

### 2019-11-30 src/args.py
Shortening up hyper on 2d cnn

@@ -7,7 +7,7 @@ parser.add_argument(
     "-m", "--model",
     nargs="?",
     choices=["svm", "mlp", "fc", "2d", "conv2d", "3d", "conv3d"],
-    default="conv3d",
+    default="3d",
     type=str,
 )
 parser.add_argument(

### 2019-11-30 src/models/conv2d.py

@@ -170,7 +170,7 @@ class SingleChannelResNet(nn.Module):
 
 
 def run(iargs):
-    dataset = Dataset(dimensions='2d')
+    dataset = Dataset(dimensions='2d', limit=6)
 
     hyper.train(
         model,
@@ -179,6 +179,6 @@ def run(iargs):
         dataset,
         hyper.Hyperparameters(
             epochs=iargs.epochs,
-            learning_rates=[1e-1, 1e-2, 1e-3],
+            learning_rates=[1e-2, 1e-3],
         )
     )

### 2019-11-30 src/__main__.py
Working on 3d cnn now

@@ -21,6 +21,8 @@ def run():
 
     elif iargs.model in ["conv2d", "2d"]:
         from src.models import conv2d
+        # stabilizes around 300 - 350 epochs
+        # might be different if we add in more participants
         conv2d.run(iargs)
 
     elif iargs.model in ["conv3d", "3d"]:

### 2019-11-30 src/models/conv2d.py

@@ -19,10 +19,6 @@ def optimizer(model, learning_rate: float = 1e-2):
         momentum=0.9,
         nesterov=True,
     )
-    # return optim.Adam(
-    #     model.parameters(),
-    #     lr=learning_rate,
-    # )
 
 
 def criterion():
@@ -48,29 +44,9 @@ def model(*args, **kwargs):
     return model
 
 
-class Net(nn.Module):
-    def __init__(self, idims):
-        super(Net, self).__init__()
-        self.conv1 = nn.Conv2d(1, 6, 5)
-        self.pool = nn.MaxPool2d(2, 2)
-        self.conv2 = nn.Conv2d(6, 16, 5)
-        # self.fc1 = nn.Linear(16 * 5 * 5, 120)
-        self.fc1 = nn.Linear(57344, 120)
-        self.fc2 = nn.Linear(120, 84)
-        self.fc3 = nn.Linear(84, 2)
-
-    def forward(self, x):
-        x = self.pool(F.relu(self.conv1(x)))
-        x = self.pool(F.relu(self.conv2(x)))
-        print("FOFO", x.shape)
-        # x = x.view(-1, 16 * 5 * 5)
-        x = x.view(-1, 57344)
-        x = F.relu(self.fc1(x))
-        x = F.relu(self.fc2(x))
-        x = self.fc3(x)
-        return x
-
-
+# -----------------------------------------
+# Sourced from torchvision.models.resnet
+# -----------------------------------------
 class SingleChannelResNet(nn.Module):
 
     def __init__(self, block, layers, num_classes=1000, zero_init_residual=False,
@@ -179,6 +155,7 @@ def run(iargs):
         dataset,
         hyper.Hyperparameters(
             epochs=iargs.epochs,
+            # 1e-3 better by a little, surprisingly?
             learning_rates=[1e-2, 1e-3],
         )
     )

### 2019-11-30 src/models/conv3d.py

@@ -9,6 +9,7 @@ import torch.optim as optim
 from torchvision.models import video
 
 from src.data.ml import Dataset
+from src.utils import hyper
 from src.utils.training import train
 
 
@@ -17,6 +18,7 @@ def optimizer(model, learning_rate: float = 1e-2):
         model.parameters(),
         lr=learning_rate,
         momentum=0.9,
+        nesterov=True,
     )
 
 
@@ -24,159 +26,8 @@ def criterion():
     return nn.CrossEntropyLoss()
 
 
-class SingleChannelStem(nn.Sequential):
-    """The default conv-batchnorm-relu stem
-    """
-    def __init__(self):
-        super(SingleChannelStem, self).__init__(
-            nn.Conv3d(1, 64, kernel_size=(3, 7, 7), stride=(1, 2, 2),
-                      padding=(1, 3, 3), bias=False),
-            nn.BatchNorm3d(64),
-            nn.ReLU(inplace=True))
-
-
-def resnet3d():
-    pretrained = False
-    progress = True
-
-    return video._video_resnet('r3d_18',
-         pretrained, progress,
-         block=video.BasicBlock,
-         conv_makers=[video.Conv3DSimple] * 4,
-         layers=[2, 2, 2, 2],
-         stem=SingleChannelStem)
-
-
-class Net(nn.Module):
-    def __init__(self, idim):
-        super(Net, self).__init__()
-
-        self.feature_layers = nn.Sequential(
-            nn.Conv3d(1, 64, kernel_size=3, padding=1),
-            nn.ReLU(),
-            # nn.MaxPool3d(kernel_size=(1, 2, 2), stride=(1, 2, 2)),
-            # nn.Conv3d(64, 128, kernel_size=3, padding=1),
-            # nn.ReLU(),
-            # nn.MaxPool3d(kernel_size=(2, 2, 2), stride=(2, 2, 2)),
-            # nn.Conv3d(128, 256, kernel_size=3, padding=1),
-            # nn.ReLU(),
-            # nn.Conv3d(256, 256, kernel_size=3, padding=1),
-            # nn.ReLU(),
-            # nn.MaxPool3d(kernel_size=(2, 2, 2), stride=(2, 2, 2)),
-            # nn.Conv3d(256, 512, kernel_size=3, padding=1),
-            # nn.ReLU(),
-            # nn.Conv3d(512, 512, kernel_size=3, padding=1),
-            # nn.ReLU(),
[…]

### 2019-11-30 paper/submission.tex
Ready to start graphing/writing up paper, 3d successful but not ostensibly more than 2d

@@
 % ----------------------------------------------------------------------
+% TODO
+% ----------------------------------------------------------------------
+\section{TODO}\label{sec:todo}
+
+NN project
+- 2d hyperparameters, let's use a LR scheduler? LR spikes all over the place still, why?
+
+- 3d CNN - typically meant for video classification using time as the third axis,
+here we are not using time, though that is a dimension we have access to,
+though doesn't make sense for this experiment/project to try to add in
+- https://pytorch.org/docs/stable/torchvision/models.html#video-classification
+
+- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
+- cut batch size to avoid scaling down images (TRIED NO BUENO?)
+
+be sure to talk about everything in paper
+
+normalization/mean image
+interpolation
+wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
+different styles of 1d (unfurl 3d, 2d) 2d (mean flattened, one slice) 3d (scale, orig)
+LR explorations
+sanity checks on brain image data, impossible for me to spot the differences
+
+batch_size and interpolation hard to juggle too?
+3d feels like "luck" plays a big part of getting training to catch/start
+seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
+
+images:
+LR / accuracy measures, of diff approaches
+hi res, mid res, interpolated down res brain scan examples,
+2d slice example
+total training time? num features comparison
+average slice image
+graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
+
+inclusion of more ppts/subjects means more training time (val/test are of same patients,
+hence more individual brain shapes means more general, but longer to learn)
+20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
+
+image segmentation attempt?
+saliency maps?
+filter views?
+
+torch save models - can eval that way
+host for downloads?
+
+future works
+better gpu / computer might be important, the 3d models are real fucking beasts to train, and their data is large as fuck
+model ensemble (free extra few percent perf)
+
+findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
+promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
+unless better hardware or serious effort is expended in getting it differentiable
+
+% ----------------------------------------------------------------------
 % END

### 2019-11-30 src/__main__.py

@@ -16,17 +16,20 @@ def run():
 
     elif iargs.model == "fc":
         from src.models import fc
-        # epochs recommended above 80 - 100
+        # stabilizes around 80 - 100 epochs @ 70%
         fc.run(iargs)
 
     elif iargs.model in ["conv2d", "2d"]:
         from src.models import conv2d
-        # stabilizes around 300 - 350 epochs
+        # stabilizes around 300 - 350 epochs @ 83%
         # might be different if we add in more participants
         conv2d.run(iargs)
 
     elif iargs.model in ["conv3d", "3d"]:
         from src.models import conv3d
+        # stabilizes around 200+ epochs @ 84%
+        # (might get to this point thanks to sgd/nesterov decay)
+        # very very slow to run (3+ hours)
         conv3d.run(iargs)
 
 

### 2019-11-30 src/data/ml.py

@@ -15,8 +15,15 @@ BATCH_SIZE = 64
 
 
 class Dataset:
-    def __init__(self, standardize=True, dimensions: str = "3d", scale=1.0, limit=3, splits=TVT_SPLITS):
+    def __init__(self,
+                 standardize=True,
+                 dimensions: str = "3d",
+                 scale=1.0,
+                 limit=3,
+                 splits=TVT_SPLITS,
+                 batch_size=BATCH_SIZE):
         self._d = dimensions
+        self._bs = batch_size
 
         X, y = self._get_experimental_data(limit=limit)
         dataset = self._transform(X, y, standardize=standardize, scale=scale)
@@ -200,7 +207,7 @@ class Dataset:
         return [
             DataLoader(
                 dataset,
-                batch_size=BATCH_SIZE,
+                batch_size=self._bs,
             ) for dataset in [
                 self._train,
                 self._val,

### 2019-11-30 src/models/conv3d.py

@@ -30,6 +30,9 @@ def model(*args, **kwargs):
     return r3d_18()
 
 
+# -----------------------------------------
+# Sourced from torchvision.models.video
+# -----------------------------------------
 class Conv3DSimple(nn.Conv3d):
     def __init__(self,
                  in_planes,
@@ -347,6 +350,7 @@ def r2plus1d_18(**kwargs):
 
 def run(iargs):
     dataset = Dataset(dimensions='3d', scale=5./6, limit=6)
+    # dataset = Dataset(dimensions='3d', batch_size=32, limit=10)
 
     hyper.train(
         model,
@@ -355,6 +359,7 @@ def run(iargs):
         dataset,
         hyper.Hyperparameters(
             epochs=iargs.epochs,
-            learning_rates=[1e-1, 1e-2, 1e-3, 1e-4],
+            # TODO can try near this, but this magnitude better than 1e-2 and 1e-4 by a lot
+            learning_rates=[1e-3],
         )
     )

### 2019-12-01 src/__init__.py
Adding save/load of model state, some more plotting details

@@ -6,3 +6,8 @@ SRC_DIR = Path(os.path.dirname(os.path.realpath(__file__)))
 PROJECT_DIR = SRC_DIR.parent
 
 DATASET_DIR = Path(os.path.join(PROJECT_DIR, "data", "ds000157"))
+
+OUTPUT_DIR = Path(os.path.join(PROJECT_DIR, "out"))
+
+PLOTS_DIR = Path(os.path.join(OUTPUT_DIR, "plots"))
+MODELS_DIR = Path(os.path.join(OUTPUT_DIR, "models"))

### 2019-12-01 src/args.py

@@ -30,5 +30,14 @@ parser.add_argument(
     action="store_true",
     default=False,
 )
+parser.add_argument(
+    "-s", "--save",
+    action="store_true",
+    default=False,
+)
+parser.add_argument(
+    "-l", "--load",
+    type=str,
+)
 
 iargs, _ = parser.parse_known_args()

### 2019-12-01 src/models/conv2d.py

@@ -9,7 +9,7 @@ import torchvision
 from torchvision.models import resnet
 
 from src.data.ml import Dataset
-from src.utils import hyper, plotting
+from src.utils import hyper, training, reload, plotting
 
 
 def optimizer(model, learning_rate: float = 1e-2):
@@ -146,16 +146,40 @@ class SingleChannelResNet(nn.Module):
 
 
 def run(iargs):
-    dataset = Dataset(dimensions='2d', limit=6)
+    dimensions = '2d'
+    subjects = 6
+    learning_rate = 1e-3
 
-    hyper.train(
-        model,
-        optimizer,
-        criterion,
-        dataset,
-        hyper.Hyperparameters(
-            epochs=iargs.epochs,
-            # 1e-3 better by a little, surprisingly?
-            learning_rates=[1e-2, 1e-3],
-        )
+    # name doesn't fully convey but helpful enough to recall most details
+    name = "conv_{}_s{}_lr{}_e{}".format(
+        dimensions,
+        subjects,
+        learning_rate,
+        iargs.epochs,
     )
+
+    dataset = Dataset(dimensions=dimensions, limit=subjects)
+
+    conv2d_model = model()
+
+    if iargs.load:
+        reload.load(conv2d_model, iargs.load)
+
+    losses, accuracies = training.train(
+        conv2d_model,
+        optimizer(conv2d_model, learning_rate=learning_rate),
+        criterion(),
+        dataset,
+        epochs=iargs.epochs,
+        print_frequency=iargs.print_freq,
+    )
+
+    if iargs.plot:
+        l_train, _, _ = dataset.get_loaders()
+        num_batches = len(l_train)
+
+        plotting.plot_loss(losses, num_batches, name)
+        plotting.plot_accuracies(accuracies, name)
+
+    if iargs.save:
+        reload.save(conv2d_model, name)

### 2019-12-01 src/utils/hyper.py

@@ -21,6 +21,7 @@ class Hyperparameters:
 
 def train(model_fn, optimizer_fn, criterion_fn, dataset: Dataset, h: Hyperparameters):
 
+    models, losses, accuracies = [], [], []
     for lr in h.learning_rates:
         print("\nLR: {}".format(lr))
 
@@ -28,10 +29,16 @@ def train(model_fn, optimizer_fn, criterion_fn, dataset: Dataset, h: Hyperparame
         optimizer = optimizer_fn(model, learning_rate=lr)
         criterion = criterion_fn()
 
-        vanilla_train(
+        l, a = vanilla_train(
             model,
             optimizer,
             criterion,
             dataset,
             epochs=h.epochs,
         )
+
+        models.append(model)
+        losses.append(l)
+        accuracies.append(a)
+
+    return models, losses, accuracies

### 2019-12-01 src/utils/plotting.py

@@ -1,22 +1,45 @@
 import matplotlib.pyplot as plt
 import numpy as np
+import os
 import seaborn as sns
 
-
-def plot_loss(losses: list):
-    ax = sns.lineplot(range(len(losses)), losses)
-    plt.show()
+from src import PLOTS_DIR
 
 
-def plot_accuracies(accuracies: list):
-    ax = sns.lineplot(range(len(accuracies)), accuracies)
-    plt.show()
+# TODO hardcode names begone
+def plot_loss(losses: list, num_batches: int, name: str):
+    # Use num_batches to smooth/avg?
+    ax = sns.lineplot(
+        range(len(losses)),
+        losses,
+    )
+    ax.set_title('Loss')
+    ax.set_ylabel('Cross Entropy Loss')
+    ax.set_xlabel('Epoch/Batch Iterations')
+
+    os.makedirs(PLOTS_DIR, exist_ok=True)
+    path = os.path.join(PLOTS_DIR, name + "_loss.png")
+    plt.savefig(path)
 
 
+# TODO other metrics, test/val
+def plot_accuracies(accuracies: list, name: str):
+    ax = sns.lineplot(
+        range(len(accuracies)),
+        accuracies,
+    )
+    ax.set_title('Validation Accuracy')
+    ax.set_ylabel('Accuracy')
+    ax.set_xlabel('Epoch')
+
+    os.makedirs(PLOTS_DIR, exist_ok=True)
+    path = os.path.join(PLOTS_DIR, name + "_accuracy.png")
+    plt.savefig(path)
+
+
+# TODO
 def load_show_slices():
     """
-    TODO
-
     def imshow(inp, title=None):
         inp = inp.numpy().transpose((1, 2, 0))
         # plt.figure(figsize=(10, 10))

### 2019-12-01 src/utils/reload.py

@@ -1,0 +1,27 @@
+import os
+
+import torch
+
+from src import MODELS_DIR
+
+
+PYTORCH_STATE_EXT = ".pt"
+
+
+def save(model, name):
+    os.makedirs(MODELS_DIR, exist_ok=True)
+
+    path = os.path.join(MODELS_DIR, name + PYTORCH_STATE_EXT)
+    print("Saving model state to {}".format(path))
+
+    torch.save(model.state_dict(), path)
+
+
+def load(model, name):
+    torch.load()
+
+    path = os.path.join(MODELS_DIR, name + PYTORCH_STATE_EXT)
+    print("Saving model state to {}".format(path))
+
+    model.load_state_dict(torch.load(path))
+    # model.eval

### 2019-12-01 src/utils/training.py

@@ -91,4 +91,5 @@ def train(
         losses.append(round_loss / num_batches)
         accuracies.append(accuracy)
 
+    print("\nDone training!\n")
     return losses, accuracies

### 2019-12-02 paper/submission.tex
Warm start results look good

@@
 \begin{abstract}
-   Abstract placeholder
+   Abstract placeholder (sub 300 words)
 \end{abstract}
@@
 \section{Problem Statement}\label{sec:problem-statement}
+
+(BACKGROUND / RELATED)
+
 The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
@@
 
+PUT BRAIN SCANS IN SUPPLEMENTARY?
+
+abstract > intro > background > related
+methods/approach > results > discussion > references
+6 pages? 7 maybe at most
+
+gradual warm-start improvements?
+variations on models? meh
+
+unsupervised segmentation - little hard
+
+surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
+but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
+bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential
+
+WORK ON GETTING PARAM NUMBERS AND
+DIFFERENT METRICS (AUC PLEASE, F1 PLEASE)
+2d is enough
+
+Submit your final submission through OpenReview. You will submit a single pdf file:
+A pdf file of your final report
+
+(OPTIONAL) supplementary can be concatenated at the end of the final report
+Report. The following is a suggested structure for the report:
+Title, Author(s)
+
+Abstract: It should not be more than 300 words
+
+Introduction: this section introduces your problem, and the overall plan for approaching your problem
+Background/Related Work: This section discusses relevant literature for your project
+Approach: This section details the framework of your project. Be specific, which means you might want to include equations, figures, plots, etc
+Experiment: This section begins with what kind of experiments you're doing, what kind of dataset(s) you're using, and what is the way you measure or evaluate your results.
+It then shows in details the results of your experiments.
+By details, we mean both quantitative evaluations (show numbers, figures, tables, etc) as well as qualitative results (show images, example results, etc).
+Conclusion: What have you learned? Suggest future ideas.
+References: This is absolutely necessary.
+Supplementary Material is not counted toward your 6-8 page limit. It is optional and is supposed to contain less important results/experiments/etc not critical for understanding the main report.
+
+try different architectures - more depth?
+see if we can break 85% to 90%, talk about potential data mislabeling frustration as a barrier to higher results possibly
+
+you are expected to show substantially more experiments and results with NN-related methods
+( :(  lame)
+
+~\cite{he2016deep} resnet
+~\cite{tran2018closer} 3d/time resnet
+~\cite{hara3dcnns} 3d cnn pt1?
+
 % ----------------------------------------------------------------------

### 2019-12-02 src/args.py

@@ -16,6 +16,11 @@ parser.add_argument(
     type=int,
 )
 parser.add_argument(
+    "--subjects",
+    default=6,
+    type=int,
+)
+parser.add_argument(
     "-p", "--plot",
     action="store_true",
     default=False,

### 2019-12-02 src/models/conv2d.py

@@ -147,18 +147,17 @@ class SingleChannelResNet(nn.Module):
 
 def run(iargs):
     dimensions = '2d'
-    subjects = 6
     learning_rate = 1e-3
 
     # name doesn't fully convey but helpful enough to recall most details
     name = "conv_{}_s{}_lr{}_e{}".format(
         dimensions,
-        subjects,
+        iargs.subjects,
         learning_rate,
         iargs.epochs,
     )
 
-    dataset = Dataset(dimensions=dimensions, limit=subjects)
+    dataset = Dataset(dimensions=dimensions, limit=iargs.subjects)
 
     conv2d_model = model()
 

### 2019-12-02 src/utils/reload.py

@@ -18,10 +18,8 @@ def save(model, name):
 
 
 def load(model, name):
-    torch.load()
-
     path = os.path.join(MODELS_DIR, name + PYTORCH_STATE_EXT)
-    print("Saving model state to {}".format(path))
+    print("Loading model state from {}".format(path))
 
     model.load_state_dict(torch.load(path))
     # model.eval

### 2019-12-02 src/utils/analytics.py
Adding saliency, some more visualizations

@@ -1,0 +1,5 @@
+"""Methods to help characterize or compare models & training"""
+
+
+def count_parameters(model):
+    return sum(p.numel() for p in model.parameters() if p.requires_grad)

### 2019-12-02 src/utils/plotting.py

@@ -8,6 +8,8 @@ from src import PLOTS_DIR
 
 # TODO hardcode names begone
 def plot_loss(losses: list, num_batches: int, name: str):
+    plt.figure()
+
     # Use num_batches to smooth/avg?
     ax = sns.lineplot(
         range(len(losses)),
@@ -24,6 +26,8 @@ def plot_loss(losses: list, num_batches: int, name: str):
 
 # TODO other metrics, test/val
 def plot_accuracies(accuracies: list, name: str):
+    plt.figure()
+
     ax = sns.lineplot(
         range(len(accuracies)),
         accuracies,
@@ -71,6 +75,8 @@ def load_show_slices():
 
 def show_slices(slices):
     """Function to display row of image slices"""
+    plt.figure()
+
     fig, axes = plt.subplots(1, len(slices))
     for i, slice in enumerate(slices):
         axes[i].imshow(slice.T, cmap="gray", origin="lower")
@@ -97,6 +103,8 @@ def load_show_experimental_scans():
     # pick a single scan to show
     single_scan = scans.slicer[:, :, :, 0]
 
+    _show_raw(single_scan)
+
     plotting.plot_stat_map(
         single_scan,
         # threshold=30,
@@ -110,6 +118,12 @@ def load_show_experimental_scans():
     plotting.plot_anat(single_scan)
 
 
+def _show_raw(scan):
+    d = scan.get_data()
+    halfway = [int(i / 2) for i in d.shape]
+    show_slices([d[halfway[0], :, :], d[:, halfway[1], :], d[:, :, halfway[2]]])
+
+
 def _show_variance(scans):
     d = scans.get_data()
     # Where in the image is there actually any information for classification?
@@ -121,3 +135,14 @@ def _show_variance(scans):
     halfway = [int(i / 2) for i in v.shape]
 
     show_slices([v[halfway[0], :, :], v[:, halfway[1], :], v[:, :, halfway[2]]])
+
+
+def _plot_acc_dump(path: str):
+    plt.figure()
+
+    z = np.fromfile(path)
+    ax = sns.lineplot(range(len(z)), z, color="orange")
+    ax.set_title("Conv2d Cold Start")
+    ax.set_ylabel("Validation Accuracy")
+    ax.set_xlabel("Epoch")
+    plt.show()

### 2019-12-02 src/utils/visualization/saliency.py

@@ -1,0 +1,61 @@
+from matplotlib import pyplot as plt
+
+import torch
+
+from src.data.ml import Dataset
+from src.models import conv2d
+from src.utils import reload
+
+
+def compute_saliency_maps(X, y, model):
+    """
+    Compute a class saliency map using the model for images X and labels y.
+
+    Input:
+    - X: Input images; Tensor of shape (N, 3, H, W)
+    - y: Labels for X; LongTensor of shape (N,)
+    - model: A pretrained CNN that will be used to compute the saliency map.
+
+    Returns:
+    - saliency: A Tensor of shape (N, H, W) giving the saliency maps for the input
+    images.
+    """
+    model.eval()
+    X.requires_grad_()
+
+    scores = model(X)
+    correct_scores = scores.gather(1, y.view(-1, 1)).squeeze()
+    correct_scores.backward(torch.ones(y.shape))
+
+    abs_grad = X.grad.abs()
+    max_values, _argmax_values = abs_grad.max(dim=1)  # dim 1 is channels
+    saliency = max_values
+
+    return saliency
+
+
+def show_saliency_maps():
+    d = Dataset(dimensions="2d", limit=1)
+    X, y = d.train
+
+    model = conv2d.model()
+    reload.load(model, "conv_2d_s6_lr0.001_e500")
+
+    saliency = compute_saliency_maps(X, y, model)
+
+    Xnp = X.detach().numpy()
+    snp = saliency.numpy()
+    # N = X.shape[0]
+    N = 6
+
+    for i in range(N):
+        plt.subplot(2, N, i + 1)
+        plt.imshow(Xnp[i].squeeze(), cmap="gray")
+        plt.axis('off')
+        # plt.title(class_names[y[i]])
+        plt.subplot(2, N, N + i + 1)
+        plt.imshow(snp[i].squeeze(), cmap="gray")
+        plt.axis('off')
+        plt.gcf().set_size_inches(12, 5)
+    plt.show()
+

### 2019-12-02 paper/submission.tex
Adding saliency, some more visualizations
Working on f1scores from trained model

@@
 
+11171266 trainable params in resnet18 2d
+33148482 trainable params in resnet18 3d (about 2.97x)
+
+data size tho -
+30x bigger (30 x 44 x 40) vs (44 x 40)
+
+Future works
+- could strive to refine 3d more, with more time and compute
+- unsupervised segmentation would be cool
+- find ways to not have to smoosh images for memory?
+- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too
+
+ACTUAL FINAL WORK
+- check out filter viz
+- see about unsupervised
+- try 3d alternative archs? (dont really need to)
+- "none of my alterations to the model architecture worked that well, most did nothing, a few broke things, could still poke around more though"
+- better assessment metrics (f1s, auc)
+
 ~\cite{he2016deep} resnet
@@
 ~\cite{hara3dcnns} 3d cnn pt1?
+
+              nonfood        food
+precision    0.893162    0.935323
+recall       0.941441    0.882629
+f1           0.916667    0.908213
+support    222.000000  213.000000
 

### 2019-12-02 src/models/conv2d.py

@@ -9,7 +9,7 @@ import torchvision
 from torchvision.models import resnet
 
 from src.data.ml import Dataset
-from src.utils import hyper, training, reload, plotting
+from src.utils import hyper, train, reload, plot
 
 
 def optimizer(model, learning_rate: float = 1e-2):
@@ -164,7 +164,7 @@ def run(iargs):
     if iargs.load:
         reload.load(conv2d_model, iargs.load)
 
-    losses, accuracies = training.train(
+    losses, accuracies = train.train(
         conv2d_model,
         optimizer(conv2d_model, learning_rate=learning_rate),
         criterion(),
@@ -177,8 +177,10 @@ def run(iargs):
         l_train, _, _ = dataset.get_loaders()
         num_batches = len(l_train)
 
-        plotting.plot_loss(losses, num_batches, name)
-        plotting.plot_accuracies(accuracies, name)
+        plot.plot_loss(losses, num_batches, name)
+        plot.plot_accuracies(accuracies, name)
 
     if iargs.save:
         reload.save(conv2d_model, name)
+
+    return conv2d_model

### 2019-12-02 src/utils/hyper.py

@@ -2,7 +2,7 @@
 Neural Network training and evaluation
 """
 from src.data.ml import Dataset
-from src.utils.training import train as vanilla_train
+from src.utils.train import train as vanilla_train
 
 
 LEARNING_RATES = [1e-1]

### 2019-12-02 src/utils/train.py

@@ -1,6 +1,7 @@
 """
 Neural Network training and evaluation
 """
+import pandas as pd
 from sklearn import metrics
 
 import torch
@@ -10,13 +11,14 @@ from src.env import device, dtype
 from src.data.ml import Dataset
 
 
-def evaluate(l, model):
+def evaluate_accuracy(l, model):
     num_correct = 0
     num_samples = 0
-    model.eval()  # set model to evaluation mode
+
+    model.eval()
     with torch.no_grad():
         for x, y in l:
-            x = x.to(device=device, dtype=dtype)  # move to device, e.g. GPU
+            x = x.to(device=device, dtype=dtype)
             y = y.to(device=device, dtype=torch.long)
             scores = model(x)
             _, preds = scores.max(1)
@@ -28,6 +30,28 @@ def evaluate(l, model):
     return acc
     # roc_auc = metrics.roc_auc_score
     # roc_auc(y, pred)
+
+
+def evaluate_f1(d: Dataset, model):
+    X, y = d.val
+
+    model.eval()  # set model to evaluation mode
+    with torch.no_grad():
+        x = X.to(device=device, dtype=dtype)
+        y = y.to(device=device, dtype=torch.long)
+        scores = model(x)
+        _, preds = scores.max(1)
+
+    y = y.cpu()
+    preds = preds.cpu()
+
+    results = metrics.precision_recall_fscore_support(y, preds)
+    df = pd.DataFrame(
+        data=results,
+        columns=['nonfood', 'food'],
+        index=['precision', 'recall', 'f1', 'support'])
+
+    return df
 
 
 def train(
@@ -86,10 +110,14 @@ def train(
             del x, y, scores
             torch.cuda.empty_cache()
 
-        accuracy = evaluate(l_validate, model)
+        accuracy = evaluate_accuracy(l_validate, model)
 
         losses.append(round_loss / num_batches)
         accuracies.append(accuracy)
+
+    p_r_f1_s = evaluate_f1(dataset, model)
+
+    print(p_r_f1_s)
 
     print("\nDone training!\n")
     return losses, accuracies

### 2019-12-03 paper/submission.tex
Working through paper, up to results looks pretty good

--- 2019-11-30 paper/submission.tex
@@
 
-\author{Evan Albright\\
+\author{Anonymized for Peer Review\\
 University of Massachusetts Amherst\\
 140 Governors Dr, Amherst, MA 01002\\
-{\tt\small ealbright@umass.edu}
+{\tt\small @umass.edu}
 }
@@
 
+% ----------------------------------------------------------------------
+% ABSTRACT
 %-------------------------------------------------------------------------
 \begin{abstract}
-   Abstract placeholder
+ Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
+ typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
+ and limit the ability to stitch together patterns across a third dimension.
+
+ Deep learning's ability to decipher and learn from such nearly impossible patterns in massive datasets,
+ imperceptible to preliminary observation makes it an excellent candidate to explore and teach us about
+ overlooked patterns of activity.
+
+ Here I showcase convolutional network's ability to correctly classify uncleaned activity data from a passive
+ image viewing experiment aimed at triggering activity in particular regions of the brain controlling reward and
+ restraint, using the data in both 2d and 3d to explore the merits of adapting video-processing techniques meant
+ to apply across time to a third spatial dimension and discuss adapting this to unsupervised segmentation to see
+ what the network can teach us from its findings.
+
 \end{abstract}
@@
 Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
-yet often only as a proof of concept.
+yet often only as a proof of concept, and limited to two dimensions.
 Others are working now to try and find universal architectures for most
@@
 
-My ambition is to create a highly accurate 3d convolutional network to first classify, then segment brain scans to
-help with my specific application of neuroimaging.
+Here we'll discuss creating highly accurate classifiers using both 2d and 3d convolutional networks
+to learn in the typical manner low level filters, here specifically created for brain scans of spatial activity, using
+no pretraining on unrelated datasets and no neuro-imaging feature analysis or reduction~\cite{shi2018feature},
+letting the model learn for itself what is important so that it can teach the observer instead.
 
-I may also attempt to explore feature analysis and reduction which has been shown to
-help~\cite{shi2018feature} with deep learning tasks, if only to manage the challenging
-dimensionality of their 3d inputs.
+Following this we'll discuss what this may mean for future works in the highly valuable ability to adapt the classifiers
+to the new, related task of unsupervised segmentation in 3 dimensions~\cite{shu2016unsupervised} to mirror what has been
+a fruitful field of 2d identification of tumors~\cite{akkus2017deep} for this specific application of telling brain
+activity patterns.
 
 % ----------------------------------------------------------------------
-% PROBLEM STATEMENT
+% BACKGROUND / RELATED
 % ----------------------------------------------------------------------
-\section{Problem Statement}\label{sec:problem-statement}
+\section{Background}\label{sec:background}
+
 The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
@@
 balanced against the hypothesized activation of restraints.
+
 In neuro-imaging, this task is known as passive image viewing.
@@
 and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
-This combination gives me high hopes of success in this specific application of learning to classify the viewed
-objects through brain activations of the subjects.
-
-I hope given the clarity in patterns in the brain scans, that I can develop a neural network to highlight what
-features and locations in the brain scans have been helpful in informing the decisions of classification, in the
-hope that such methods could be adapted to harder tasks in the future.
+This combination makes this an excellent selection for a proof of concept application of deep learning as the chances
+of success on a less subtle activation pattern, observable to trained researchers, are quite high.
+This specific application then can serve as a jumping off point for future works where the activity patterns are less
+distinguishable, and may be an interesting addition to the works in brain-pattern controlled prosthetics.
+This dataset is also ideal as it comes with a fairly usable ground-truth labeling mechanism, as I am not currently
+investigating the capacity for unsupervised differentiation.
 
@@
 OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}
-It consists of a survey detailing the relationship of each subject with food and diet taken after the experimental
-proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w brain scan,
-and time series of bold scans of about 370-375 scans, which encompasses the approximately 10 minute experiment
-in its entirety.
-The bold scans are quite sizeable given the number of them, however they are a relatively low resolution, consisting
-of 64 sagital by 64 coronal by 30 axial scans.
+
+The dataset follows the BIDS (Brain Imaging Data Structure) layout adopted by many neuroscience researchers.
+It consists of a survey detailing the relationship of each of the 30 subjects with food and diet taken after the experimental
+proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w Modality fMRI scan seen in figure~\ref{fig:ortho-t1w},
+and time series of 370-375 BOLD Modality fMRI scans seen in figure~\ref{fig:ortho-bold}, encompassing approximately 10 minutes
+capturing brain activity of the subjects during their exposure to the experimental triggers.
+
+The bold scans are quite sizeable as far as datasets go, given the number of each per subject,
+however they are a relatively low resolution, consisting
+of 64 sagital by 64 coronal by 30 axial scans, with only a single image channel.
+
+The activation is measured in a positive magnitude visible in figure~\ref{fig:ex_stat_map}, through a helper method in a neuro-imaging specialized
+image handling library ~\cite{brett_matthew_2019_3544468}.
+
+I had to manually infer pairings of the scans with the experimental phase for classification training given the events log associated
+with each subject and the advertised scan times from the paper.
+While the data does appear healthy, there may be some discrepancies of interpretation which would warrant future investigation if the
+performance on this particular experiment was the primary focus.
 
@@
 
-You may find this project repository here: \url{https://github.com/e-m-albright/CS682}
+You may find this project repository here: (personal github link removed for anonymity)
 Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
@@
 
-% ----------------------------------------------------------------------
-% TECHNICAL APPROACH
-% ----------------------------------------------------------------------
-\section{Technical Approach}\label{sec:technical-approach}
-I intend to explore the efficacy of 3d convolutional networks which have been shown to have a slight performance
-edge on 2d convolutions~\cite{payan2015predicting}.
-Given the nature of the input images being 3d brain scans this makes intuitive sense and could allow for much more
-interesting patterns of brain activity and structuring being recognized.
+\begin{figure}
+  \includegraphics[width=\linewidth]{images/orthoview_bold.png}
+  \caption{Bold Modality Low-Res 3d Experimental Scans of Subject Brain}
+  \label{fig:ortho-bold}
+\end{figure}
 
-As for the deeper architecture, I'm hoping to imitate some of the more successful networks like VGG or ResNet,
-however I wouldn't be surprised if they don't adapt well to this particular application, especially given how
-interesting and important it could be to keep the input data in a three dimensional format.
-
-It is my hope that I can also learn from other developments of object detection to see what regions in the
-brain become active for a particular classification, like Faster R-CNN
-
-\subsection{Novelty}\label{subsec:novelty}
-
-I have not seen any application of classification or object segmentation to neuroimaging in 3d
-on brain activity patterns.
-I have seen some works which I hope to leverage with similar aims like that of structural identification
-for tumors, in two dimensions~\cite{akkus2017deep}.
-However at the very least I hope to produce a very successful network on my chosen problem first and foremost rather
-than assume I am the only person to ever think of something as intuitive as bringing neural networks to bear
-on neuro-imaging tasks.
-
-I hope to find a methodology of first classifying correctly the brain scans, improve accuracy as much as possible,
-and adapt the approach into tools for identifying important regions of activity in the
-brain which are being utilized for classification.
-
-I am not chiefly focused on coping with the challenges of memory but it may become too important to overlook.
-Each dataset input is a $64 \times 64 \times 30$ slice of floats, for which we've got
-about $375 scans \times 30 subjects$ total.
-In my work on preliminary results I have already been confronted by the limitations of my current machine and have
-been trying working around the issue until I am ready to attempt to resolve it.
-This may be an unavoidable area to simultaneously try to innovate for the field of
-neuro-imaging analysis with deep learning
+\begin{figure}
+  \includegraphics[width=\linewidth]{images/example_stat_map.png}
+  \caption{Activity Measures from Bold Scans}
+  \label{fig:ex_stat_map}
+\end{figure}
 
 % ----------------------------------------------------------------------
-% INTERMEDIATE RESULTS
+% APPROACH
 % ----------------------------------------------------------------------
-\section{Intermediate Results}\label{sec:intermediate-results}
+\section{Approach}\label{sec:approach}
+
+Beginning with SKLearn~\cite{scikit-learn} data readiness was proven with moderately successful classification using
+a simple SVM and then the SKLearn Multi-Layer-Perceptron neural network, getting 70-80\% accuracy, sufficing the
+initial sanity check that deep learning would be ready to work on.
+
+% TODO these are relevant citations, right?
+Leveraging PyTorch~\cite{paszke2017automatic} first a framework for learning was likewise vetted using a simple fully
+connected shallow network, graduating to the state of the art convolutional networks, chiefly focussed
+on the ResNet~\cite{he2016deep} architecture styles, which have been developed for video classification~\cite{tran2018closer}
+utilizing 3 dimensional convolutions~\cite{hara3dcnns} much in the way 2 dimensional convolutions operate, which have
+been shown to hold a slight edge on 2 dimensions~\cite{payan2015predicting} in other applications.
+
+\subsection{Data Preparation}\label{subsec:data-preparation}
+
+First as the data was not prepared for classification, critical cleaning include subtracting the mean image to better
+highlight meaningful variance, and standardizing the data between 0 and 1.
+
+Given the nature of the input images being 3d brain scans it's intuitive that examining the brain in 3 dimensions
+allows for much more interesting patterns of brain activity and structuring being recognized.
+
+This comes with a substantial cost however, as the extra dimension explodes the memory demands of the input data, and
+the number of trainable parameters for the models.
+Many of the recommendations following will vary based on the particular hardware, but with a 4GB GPU I had to be quite
+careful about batch-sizes to not over-allocate memory.
+To combat this, first, a relatively conservative amount of blank space is cropped from the images, halving the total footprint.
+Following this, in order to use a batch-size of 64, I interpolate images to approximately $\frac{5}{6}$ their original dimension
+
+In addition, I typically do not use the full cohort of 30 subjects, as I even encounter some RAM restrictions with the
+full raw dataset consisting of
+$64 sagital \times 64 coronal \times 30 axial \times 32bit floats \times 375 scans \times 30 subjects$ or roughly 5.53 gigabytes
+
+In working with 2 dimensions, omitting the axial dimension for a flat cross section, you'll see the dataset is
+$\frac{1}{30}$ the size and hence, much more manageable.
+
+\subsection{Examining the Data}\label{subsec:examining-data}
+
+Two fully processed, correctly classified brain scans are presented in the figure~\ref{fig:comparison}
+which pose a daunting challenge to human labelers.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{images/nf_f_hard_comparison.png}
+  \caption{Nonfood-Stimuli \& Food-Stimuli Respectively from the same Subject}
+  \label{fig:comparison}
+\end{figure}
+
+\subsection{Model Selection}\label{subsec:model-selection}
+
+For the primary focus of comparing and improving classification accuracy using 2d and 3d architectures, ResNet models
+of the standard 18 layers to 100+ layers are examined, as the ResNet family is both highly competitive, and demonstrated
+in both 2 and 3 dimensions which will help comparisons in exploring the benefits of 3d brain scan usage.
+Video processing typically limits the layers to a more shallow level than that of image processing, as unfortunately as
+with the data, the model parameters explode out of control pretty fast.
+The following parameter counts are from standard ResNet architectures adapted to a single input channel.
+
+\begin{center}
+ \begin{tabular}{||c c c||}
+ \hline
+ - & ResNet-18 2d & ResNet-18 3d \\ [0.5ex]
+ \hline\hline
+ Counts & 11,171,266 & 33,148,482 \\ [1ex]
+ \hline
+ Ratio & 1.0 & 2.97 \\ [1ex]
+ \hline
+\end{tabular}
+\end{center}
+
+\subsection{Comparisons}\label{subsec:comparisons}
+
+Model performance is typically evaluated using accuracy, precision, recall, and AUC metrics, as well as total time to
+train.
+
+% ----------------------------------------------------------------------
+% RESULTS
+% ----------------------------------------------------------------------
+\section{Results}\label{sec:results}
 At the date of the milestone, I have shown that my experimental set-up is complete to begin
@@
 
-\subsection{Sanity Checks}\label{subsec:sanity-checks}
-Using the data, I have implemented a very simple, yet effective 80\% accurate SVM classifier to prove the data
-should be ready for use with my chosen deep learning methodologies.
+\subsection{Model Architecture}\label{subsec:model-arch}
 
-I'm now confident in the classifications of each of the scans (about 375 per subject, for 30 subjects)
+the project explored minimally the hyperparameters of the chosen model architectures
+chiefly was finding the right learning rate for the given normalized data to the model
+typically 1e-2 to 1e-4 worked, with 1e-3 being good, and decay being very very helpful
+model quite sensitive to adjustments
 
-\subsection{Deep Learning \& Future Works}\label{subsec:deep-learning}
-I have tried feeding the data into very basic neural networks with mixed success, and am in the active process of
-troubleshooting.
-Once I have matched performance with the SVM using an implementation of a reasonable neural network using PyTorch, I
-will begin the process of looking for refinements and innovations alluded to in previous sections, as well as
-transitioning the task to object detection if I can.
+- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
 
-% ----------------------------------------------------------------------
-% TODO
-% ----------------------------------------------------------------------
-\section{TODO}\label{sec:todo}
+\subsection{Scope of Number of Subjects}\label{subsec:num-subjects}
 
-NN project
-- 2d hyperparameters, let's use a LR scheduler? LR spikes all over the place still, why?
+inclusion of more ppts/subjects means more training time (val/test are of same patients,
+hence more individual brain shapes means more general, but longer to learn)
+20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
 
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/2d_cold.png}
+  \caption{Cold-Start 2D Training Progress}
+  \label{fig:2d_cold}
+\end{figure}
+
+\subsection{Warm Starts}\label{subsec:warm-starts}
+saving model parameter dict, loading and training again
+wowee good stuff, hope to use a LR scheduler in the future to try to control this better?
+surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
+but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
+bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential
+
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/2d_warm.png}
+  \caption{Warm-Start 2D Training Progress}
+  \label{fig:2d_warm}
+\end{figure}
+
+\subsection{3 Dimensions}\label{subsec:3-dimensions}
 - 3d CNN - typically meant for video classification using time as the third axis,
@@
 
-- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
-- cut batch size to avoid scaling down images (TRIED NO BUENO?)
+okay it works, and has potential, certainly given the data is richer to be the best, but wow it's frustratingly slow
+findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
+promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
+unless better hardware or serious effort is expended in getting it differentiable
 
-be sure to talk about everything in paper
-
-normalization/mean image
-interpolation
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/3d_cold.png}
+  \caption{Cold-Start 3D Training Progress}
+  \label{fig:3d_cold}
+\end{figure}
 wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
@@
 LR explorations
-sanity checks on brain image data, impossible for me to spot the differences
-
-batch_size and interpolation hard to juggle too?
-3d feels like "luck" plays a big part of getting training to catch/start
-seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
-
-images:
-LR / accuracy measures, of diff approaches
-hi res, mid res, interpolated down res brain scan examples,
-2d slice example
-total training time? num features comparison
-average slice image
-graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
-
-inclusion of more ppts/subjects means more training time (val/test are of same patients,
-hence more individual brain shapes means more general, but longer to learn)
-20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
-
-image segmentation attempt?
-saliency maps?
-filter views?
-
-torch save models - can eval that way
-host for downloads?
 
@@
 
-findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
-promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
-unless better hardware or serious effort is expended in getting it differentiable
+"none of my alterations to the model architecture worked that well, most did nothing, a few broke things, could still poke around more though"
+graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
+batch_size and interpolation hard to juggle too?
+3d feels like "luck" plays a big part of getting training to catch/start
+seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
+
+cut batch size to avoid scaling down images (TRIED NO BUENO?)
+
+\subsection{Performance}\label{subsec:performance}
+
+ % P-R-F1-S
+\begin{center}
+ \begin{tabular}{||c c c||}
+ \hline
+ - & Nonfood & Food \\ [0.5ex]
+ \hline\hline
+Precision & 0.949074 & 0.958904 \\
+ \hline
+Recall & 0.957944 & 0.950226 \\
+ \hline
+F1 & 0.953488 & 0.954545 \\
+ \hline
+Support & 214 & 221 \\ [1ex]
+ \hline
+\end{tabular}
+
+ % ROC-AUC / ACC
+\end{center}
+\begin{center}
+ \begin{tabular}{||c c||}
+ \hline
+ ROC-AUC & Accuracy \\ [0.5ex]
+ \hline\hline
+  0.9847 & 0.954 \\ [1ex]
+ \hline
+\end{tabular}
+\end{center}
+
+Can get a little higher but these are already encouraging numbers (96 seen) often the model starts leaning a little
+one way or the other and this was one of the most balanced runs I've seen
+
+I'm a little unclear about the reliability of my data labeling, so I'm happy with this result, I'd be stoked
+to reproduce on 3d, and trusting the model this much, open the door to segmentation
+
+% ----------------------------------------------------------------------
+% DISCUSSION
+% ----------------------------------------------------------------------
+\section{Discussion}\label{sec:discussion}
+
+\subsection{Learning}\label{subsec:learning}
+very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
+cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
+this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at
+
+\subsection{Future Works}\label{subsec:future-works}
+Unsupervised segmentation, especially in 3d would be cool
+Future works
+- could strive to refine 3d more, with more time and compute
+- unsupervised segmentation would be cool
+- find ways to not have to smoosh images for memory?
+- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too
+
+began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
+- check out filter viz
+- see about unsupervised
+- try 3d alternative archs? (dont really need to)
+
+might want to improve 3d before going on to segmentation,
+
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/saliency.png}
+  \caption{caption}
+  \label{fig:saliency}
+\end{figure}
 
@@
 
+% ----------------------------------------------------------------------
+% SUPPLEMENTARY
+% ----------------------------------------------------------------------
+\section{Supplementary Material}\label{sec:supplementary}
+
+Extra material for reference
+
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/orthoview_t1w.png}
+  \caption{caption}
+  \label{fig:ortho-t1w}
+\end{figure}
+
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
+  \caption{caption}
+  \label{fig:cortex-var}
+\end{figure}
+
 \end{document}

### 2019-12-03 src/args.py

@@ -3,6 +3,10 @@ import argparse
 
 parser = argparse.ArgumentParser()
 
+
+# -----------------------
+# Run mode selection
+# -----------------------
 parser.add_argument(
     "-m", "--model",
     nargs="?",
@@ -11,15 +15,41 @@ parser.add_argument(
     type=str,
 )
 parser.add_argument(
+    "-s", "--save",
+    action="store_true",
+    default=False,
+)
+parser.add_argument(
+    "-l", "--load",
+    type=str,
+)
+
+# -----------------------
+# Hyperparameters
+# -----------------------
+parser.add_argument(
     "-e", "--epochs",
     default=3,
     type=int,
 )
 parser.add_argument(
+    "--learning-rate",
+    default=None,
+    type=float,
+)
+
+# -----------------------
+# Data selection
+# -----------------------
+parser.add_argument(
     "--subjects",
     default=6,
     type=int,
 )
+
+# -----------------------
+# Operation selection
+# -----------------------
 parser.add_argument(
     "-p", "--plot",
     action="store_true",
@@ -35,14 +65,6 @@ parser.add_argument(
     action="store_true",
     default=False,
 )
-parser.add_argument(
-    "-s", "--save",
-    action="store_true",
-    default=False,
-)
-parser.add_argument(
-    "-l", "--load",
-    type=str,
-)
+
 
 iargs, _ = parser.parse_known_args()

### 2019-12-03 src/models/conv2d.py

@@ -46,6 +46,9 @@ def model(*args, **kwargs):
 
 # -----------------------------------------
 # Sourced from torchvision.models.resnet
+#
+# Code pulled in for convenience of customization
+# especially in altering the number of channels
 # -----------------------------------------
 class SingleChannelResNet(nn.Module):
 
@@ -147,7 +150,7 @@ class SingleChannelResNet(nn.Module):
 
 def run(iargs):
     dimensions = '2d'
-    learning_rate = 1e-3
+    learning_rate = 1e-3 if iargs.learning_rate is None else iargs.learning_rate
 
     # name doesn't fully convey but helpful enough to recall most details
     name = "conv_{}_s{}_lr{}_e{}".format(

### 2019-12-03 src/models/conv3d.py

@@ -10,7 +10,7 @@ from torchvision.models import video
 
 from src.data.ml import Dataset
 from src.utils import hyper
-from src.utils.training import train
+from src.utils import train
 
 
 def optimizer(model, learning_rate: float = 1e-2):
@@ -32,6 +32,9 @@ def model(*args, **kwargs):
 
 # -----------------------------------------
 # Sourced from torchvision.models.video
+#
+# Code pulled in for convenience of customization
+# especially in altering the number of channels
 # -----------------------------------------
 class Conv3DSimple(nn.Conv3d):
     def __init__(self,
@@ -338,7 +341,7 @@ def r2plus1d_18(**kwargs):
 #     # model = conv3d.Net(dataset.dimensions)
 #     model = r3d_18()
 #
-#     train(
+#     train.train(
 #         model,
 #         optimizer(model, learning_rate=1e-2),
 #         criterion(),

### 2019-12-03 src/models/fc.py

@@ -11,8 +11,7 @@ import torch.optim as optim
 
 from src.data.ml import Dataset
 from src.defs.layers import flatten, Flatten
-from src.utils import plotting
-from src.utils.training import train
+from src.utils import train, plot
 
 
 # def model(idim, hdim: int = 300):
@@ -82,7 +81,7 @@ def run(iargs):
 
     model = Net(dataset.dimensions)
 
-    losses, accuracies = train(
+    losses, accuracies = train.train(
         model,
         optimizer(model, learning_rate=1e-1),
         criterion(),
@@ -92,5 +91,5 @@ def run(iargs):
     )
 
     if iargs.plot:
-        plotting.plot_loss(losses)
-        plotting.plot_accuracies(accuracies)
+        plot.plot_loss(losses)
+        plot.plot_accuracies(accuracies)

### 2019-12-03 src/utils/train.py

@@ -12,6 +12,7 @@ from src.data.ml import Dataset
 
 
 def evaluate_accuracy(l, model):
+    """Intermediate training monitor metric"""
     num_correct = 0
     num_samples = 0
 
@@ -28,11 +29,10 @@ def evaluate_accuracy(l, model):
         print("{:d} / {:d} correct (%{:.2f})".format(num_correct, num_samples, 100 * acc))
 
     return acc
-    # roc_auc = metrics.roc_auc_score
-    # roc_auc(y, pred)
 
 
-def evaluate_f1(d: Dataset, model):
+def evaluate_final(d: Dataset, model):
+    """Final evaluation using the val/test looking at performance of model more in depth"""
     X, y = d.val
 
     model.eval()  # set model to evaluation mode
@@ -40,10 +40,18 @@ def evaluate_f1(d: Dataset, model):
         x = X.to(device=device, dtype=dtype)
         y = y.to(device=device, dtype=torch.long)
         scores = model(x)
+
+        # For AUC
+        probs = F.softmax(scores)
+        # probs = probs.gather(1, y.view(-1, 1)).squeeze()
+        probs = probs[:, 1]  # roc_auc works oddly
+
+        # For F1
         _, preds = scores.max(1)
 
     y = y.cpu()
     preds = preds.cpu()
+    probs = probs.cpu()
 
     results = metrics.precision_recall_fscore_support(y, preds)
     df = pd.DataFrame(
@@ -51,7 +59,9 @@ def evaluate_f1(d: Dataset, model):
         columns=['nonfood', 'food'],
         index=['precision', 'recall', 'f1', 'support'])
 
-    return df
+    auc = metrics.roc_auc_score(y, probs)
+
+    return df, auc
 
 
 def train(
@@ -115,9 +125,10 @@ def train(
         losses.append(round_loss / num_batches)
         accuracies.append(accuracy)
 
-    p_r_f1_s = evaluate_f1(dataset, model)
+    f1, auc = evaluate_final(dataset, model)
 
-    print(p_r_f1_s)
+    print(f1)
+    print("Final AUC: {:.4f}".format(auc))
 
     print("\nDone training!\n")
     return losses, accuracies

### 2019-12-03 src/utils/visualize/classes.py

@@ -1,0 +1,118 @@
+"""
+TODO
+    incomplete, trying to visualize model activations on different class types to see the abstract idea of what
+    the filters key off of.
+
+    Base code from HW3
+
+    Might find help here for the classes
+    https://towardsdatascience.com/how-to-visualize-convolutional-features-in-40-lines-of-code-70b7d87b0030
+
+"""
+
+import random
+import matplotlib.pyplot as plt
+from scipy.ndimage.filters import gaussian_filter1d
+
+import torch
+
+
+def rescale(x):
+    low, high = x.min(), x.max()
+    x_rescaled = (x - low) / (high - low)
+    return x_rescaled
+
+
+def blur_image(X, sigma=1):
+    X_np = X.cpu().clone().numpy()
+    X_np = gaussian_filter1d(X_np, sigma, axis=2)
+    X_np = gaussian_filter1d(X_np, sigma, axis=3)
+    X.copy_(torch.Tensor(X_np).type_as(X))
+    return X
+
+
+def jitter(X, ox, oy):
+    """
+    Helper function to randomly jitter an image.
+
+    Inputs
+    - X: PyTorch Tensor of shape (N, C, H, W)
+    - ox, oy: Integers giving number of pixels to jitter along W and H axes
+
+    Returns: A new PyTorch Tensor of shape (N, C, H, W)
+    """
+    if ox != 0:
+        left = X[:, :, :, :-ox]
+        right = X[:, :, :, -ox:]
+        X = torch.cat([right, left], dim=3)
+    if oy != 0:
+        top = X[:, :, :-oy]
+        bottom = X[:, :, -oy:]
+        X = torch.cat([bottom, top], dim=2)
+    return X
+
+
+def create_class_visualization(target_y, model, dtype, **kwargs):
+    """
+    Generate an image to maximize the score of target_y under a pretrained model.
+
+    Inputs:
+    - target_y: Integer in the range [0, 1000) giving the index of the class
+    - model: A pretrained CNN that will be used to generate the image
+    - dtype: Torch datatype to use for computations
+
+    Keyword arguments:
+    - l2_reg: Strength of L2 regularization on the image
+    - learning_rate: How big of a step to take
+    - num_iterations: How many iterations to use
+    - blur_every: How often to blur the image as an implicit regularizer
+    - max_jitter: How much to gjitter the image as an implicit regularizer
[…]

### 2019-12-03 src/utils/visualize/filters.py

@@ -1,0 +1,6 @@
+"""
+TODO
+    have a go at filter visualizations
+    https://towardsdatascience.com/visualizing-convolution-neural-networks-using-pytorch-3dfa8443e74e
+
+"""

### 2019-12-03 src/utils/visualize/saliency.py

@@ -1,3 +1,9 @@
+"""
+TODO
+    Saliency isn't looking that great, 1 channel makes it hard to see much I suppose
+    Base code from HW3, might want to tinker more to get more sensible saliency maps
+"""
+
 from matplotlib import pyplot as plt
 
 import torch

### 2019-12-03 paper/sections/approach.tex
Breaking out sections to make working on paper easier

--- 2019-12-03 paper/submission.tex
@@
-\documentclass[10pt,twocolumn,letterpaper]{article}
-
-\usepackage{cvpr}
-%\usepackage{times}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-% Include other packages here, before hyperref.
-
-% If you comment hyperref and then uncomment it, you should delete
-% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
-% run, let it finish, and you should be clear).
-\usepackage[breaklinks=true,bookmarks=false]{hyperref}
-
-\cvprfinalcopy % *** Uncomment this line for the final submission
-
-\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
-\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
-
-% Pages are numbered in submission mode, and unnumbered in camera-ready
-%\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{1}
-\begin{document}
-
-%-------------------------------------------------------------------------
-\title{Deep Learning for 3D Brain Activity Scans}
-
-\author{Anonymized for Peer Review\\
-University of Massachusetts Amherst\\
-140 Governors Dr, Amherst, MA 01002\\
-{\tt\small @umass.edu}
-}
-
-\maketitle
-%\thispagestyle{empty}
-
-% ----------------------------------------------------------------------
-% ABSTRACT
-%-------------------------------------------------------------------------
-\begin{abstract}
- Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
- typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
- and limit the ability to stitch together patterns across a third dimension.
-
- Deep learning's ability to decipher and learn from such nearly impossible patterns in massive datasets,
- imperceptible to preliminary observation makes it an excellent candidate to explore and teach us about
- overlooked patterns of activity.
-
- Here I showcase convolutional network's ability to correctly classify uncleaned activity data from a passive
- image viewing experiment aimed at triggering activity in particular regions of the brain controlling reward and
- restraint, using the data in both 2d and 3d to explore the merits of adapting video-processing techniques meant
- to apply across time to a third spatial dimension and discuss adapting this to unsupervised segmentation to see
- what the network can teach us from its findings.
-
-\end{abstract}
-
-% ----------------------------------------------------------------------
-% INTRODUCTION
-% ----------------------------------------------------------------------
-\section{Introduction}\label{sec:introduction}
-Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
-for mental health research and disease diagnosis.
-High dimensional, massive datasets, with intertwining patterns of activity and structure in three or four
-dimensions, are a lot to manage by hand.
-
-Classification of activity and structures should come quite naturally with deep learning practices.
-Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
-yet often only as a proof of concept, and limited to two dimensions.
-Others are working now to try and find universal architectures for most
-neuro-imaging tasks~\cite{henschel2019fastsurfer}, though it seems this is still far from a solved
-application with satisfactory performances across the board.
-
-Of more interest to myself, deep learning networks in their proven capacity to segment images based on what entities
-they identify in classification could surely be adopted to highlight what patterns of activity have informed their
-own decision making and perhaps clue researchers in to overlooked details.
-
-Here we'll discuss creating highly accurate classifiers using both 2d and 3d convolutional networks
-to learn in the typical manner low level filters, here specifically created for brain scans of spatial activity, using
-no pretraining on unrelated datasets and no neuro-imaging feature analysis or reduction~\cite{shi2018feature},
-letting the model learn for itself what is important so that it can teach the observer instead.
-
-Following this we'll discuss what this may mean for future works in the highly valuable ability to adapt the classifiers
-to the new, related task of unsupervised segmentation in 3 dimensions~\cite{shu2016unsupervised} to mirror what has been
-a fruitful field of 2d identification of tumors~\cite{akkus2017deep} for this specific application of telling brain
-activity patterns.
-
-% ----------------------------------------------------------------------
-% BACKGROUND / RELATED
-% ----------------------------------------------------------------------
-\section{Background}\label{sec:background}
-
-The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
-Netherlands circa 2013 who were presented with various grids of images featuring
-either neutral entities, (e.g.\ office chairs, windows, jackets) or foods~\cite{smeets2013allured}.
-The foods were specifically types of calorically dense, tempting foods, which the individuals viewing would likely
-have formed restraint with given their success in maintaining healthy body mass indices.
-The experiment was meant to highlight brain structures roles in anticipated rewards from consuming delicious foods,
-balanced against the hypothesized activation of restraints.
-
-In neuro-imaging, this task is known as passive image viewing.
-
-Given the quite important role of food and willpower in our lives, the activation patterns are likely to be powerful
-and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
-This combination makes this an excellent selection for a proof of concept application of deep learning as the chances
-of success on a less subtle activation pattern, observable to trained researchers, are quite high.
-This specific application then can serve as a jumping off point for future works where the activity patterns are less
-distinguishable, and may be an interesting addition to the works in brain-pattern controlled prosthetics.
-This dataset is also ideal as it comes with a fairly usable ground-truth labeling mechanism, as I am not currently
-investigating the capacity for unsupervised differentiation.
-
-The dataset can be found from the excellent
-OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}
-
-The dataset follows the BIDS (Brain Imaging Data Structure) layout adopted by many neuroscience researchers.
-It consists of a survey detailing the relationship of each of the 30 subjects with food and diet taken after the experimental
-proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w Modality fMRI scan seen in figure~\ref{fig:ortho-t1w},
-and time series of 370-375 BOLD Modality fMRI scans seen in figure~\ref{fig:ortho-bold}, encompassing approximately 10 minutes
-capturing brain activity of the subjects during their exposure to the experimental triggers.
-
-The bold scans are quite sizeable as far as datasets go, given the number of each per subject,
-however they are a relatively low resolution, consisting
-of 64 sagital by 64 coronal by 30 axial scans, with only a single image channel.
-
-The activation is measured in a positive magnitude visible in figure~\ref{fig:ex_stat_map}, through a helper method in a neuro-imaging specialized
-image handling library ~\cite{brett_matthew_2019_3544468}.
-
-I had to manually infer pairings of the scans with the experimental phase for classification training given the events log associated
-with each subject and the advertised scan times from the paper.
-While the data does appear healthy, there may be some discrepancies of interpretation which would warrant future investigation if the
-performance on this particular experiment was the primary focus.
-
-You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.
-
-You may find this project repository here: (personal github link removed for anonymity)
-Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
-abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{images/orthoview_bold.png}
-  \caption{Bold Modality Low-Res 3d Experimental Scans of Subject Brain}
-  \label{fig:ortho-bold}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{images/example_stat_map.png}
-  \caption{Activity Measures from Bold Scans}
-  \label{fig:ex_stat_map}
-\end{figure}
-
-% ----------------------------------------------------------------------
-% APPROACH
-% ----------------------------------------------------------------------
 \section{Approach}\label{sec:approach}
@@
 on the ResNet~\cite{he2016deep} architecture styles, which have been developed for video classification~\cite{tran2018closer}
-utilizing 3 dimensional convolutions~\cite{hara3dcnns} much in the way 2 dimensional convolutions operate, which have
+utilizing 3 dimensional convolutions much in the way 2 dimensional convolutions operate, which have
 been shown to hold a slight edge on 2 dimensions~\cite{payan2015predicting} in other applications.
@@
 full raw dataset consisting of
-$64 sagital \times 64 coronal \times 30 axial \times 32bit floats \times 375 scans \times 30 subjects$ or roughly 5.53 gigabytes
+$64 sagital \times 64 coronal \times 30 axial \times 1 channel \times 32bit floats \times 375 scans \times 30 subjects$ or roughly 5.53 gigabytes
 
@@
 Video processing typically limits the layers to a more shallow level than that of image processing, as unfortunately as
-with the data, the model parameters explode out of control pretty fast.
+with the data, the model parameters explode out of control pretty fast, though some have seen minor improvements by
+going deeper~\cite{hara3dcnns}.
 The following parameter counts are from standard ResNet architectures adapted to a single input channel.
@@
 train.
-
-% ----------------------------------------------------------------------
-% RESULTS
-% ----------------------------------------------------------------------
-\section{Results}\label{sec:results}
-At the date of the milestone, I have shown that my experimental set-up is complete to begin
-iterating towards the novel contributions aimed for by this project.
-
-I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
-network for classification on my chosen dataset incorporating my own contributions to the process.
-
-I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
-excited nevertheless to try.
-
-\subsection{Data}\label{subsec:data}
-I have become familiar with the chosen dataset, and created a machine learning friendly format
-based off of the data.
-
-I have inferred classification labels from the experimental events timings and the estimated time
-of each complete scan in the dataset.
-
-\subsection{Model Architecture}\label{subsec:model-arch}
-
-the project explored minimally the hyperparameters of the chosen model architectures
-chiefly was finding the right learning rate for the given normalized data to the model
-typically 1e-2 to 1e-4 worked, with 1e-3 being good, and decay being very very helpful
-model quite sensitive to adjustments
-
-- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
-
-\subsection{Scope of Number of Subjects}\label{subsec:num-subjects}
-
-inclusion of more ppts/subjects means more training time (val/test are of same patients,
-hence more individual brain shapes means more general, but longer to learn)
-20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/2d_cold.png}
-  \caption{Cold-Start 2D Training Progress}
-  \label{fig:2d_cold}
-\end{figure}
-
-\subsection{Warm Starts}\label{subsec:warm-starts}
-saving model parameter dict, loading and training again
-wowee good stuff, hope to use a LR scheduler in the future to try to control this better?
-surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
-but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
-bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/2d_warm.png}
-  \caption{Warm-Start 2D Training Progress}
-  \label{fig:2d_warm}
-\end{figure}
-
-\subsection{3 Dimensions}\label{subsec:3-dimensions}
-- 3d CNN - typically meant for video classification using time as the third axis,
-here we are not using time, though that is a dimension we have access to,
-though doesn't make sense for this experiment/project to try to add in
-- https://pytorch.org/docs/stable/torchvision/models.html#video-classification
-
-okay it works, and has potential, certainly given the data is richer to be the best, but wow it's frustratingly slow
-findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
-promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
-unless better hardware or serious effort is expended in getting it differentiable
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/3d_cold.png}
-  \caption{Cold-Start 3D Training Progress}
-  \label{fig:3d_cold}
-\end{figure}
-wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
-different styles of 1d (unfurl 3d, 2d) 2d (mean flattened, one slice) 3d (scale, orig)
-LR explorations
-
-future works
-better gpu / computer might be important, the 3d models are real fucking beasts to train, and their data is large as fuck
-model ensemble (free extra few percent perf)
-
-"none of my alterations to the model architecture worked that well, most did nothing, a few broke things, could still poke around more though"
-graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
-batch_size and interpolation hard to juggle too?
-3d feels like "luck" plays a big part of getting training to catch/start
-seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
-
-cut batch size to avoid scaling down images (TRIED NO BUENO?)
-
-\subsection{Performance}\label{subsec:performance}
-
- % P-R-F1-S
-\begin{center}
- \begin{tabular}{||c c c||}
- \hline
- - & Nonfood & Food \\ [0.5ex]
- \hline\hline
-Precision & 0.949074 & 0.958904 \\
- \hline
-Recall & 0.957944 & 0.950226 \\
- \hline
-F1 & 0.953488 & 0.954545 \\
- \hline
-Support & 214 & 221 \\ [1ex]
- \hline
-\end{tabular}
-
- % ROC-AUC / ACC
-\end{center}
-\begin{center}
- \begin{tabular}{||c c||}
- \hline
- ROC-AUC & Accuracy \\ [0.5ex]
- \hline\hline
-  0.9847 & 0.954 \\ [1ex]
- \hline
-\end{tabular}
-\end{center}
-
-Can get a little higher but these are already encouraging numbers (96 seen) often the model starts leaning a little
-one way or the other and this was one of the most balanced runs I've seen
-
-I'm a little unclear about the reliability of my data labeling, so I'm happy with this result, I'd be stoked
-to reproduce on 3d, and trusting the model this much, open the door to segmentation
-
-% ----------------------------------------------------------------------
-% DISCUSSION
-% ----------------------------------------------------------------------
-\section{Discussion}\label{sec:discussion}
-
-\subsection{Learning}\label{subsec:learning}
-very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
-cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
-this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at
-
-\subsection{Future Works}\label{subsec:future-works}
-Unsupervised segmentation, especially in 3d would be cool
-Future works
-- could strive to refine 3d more, with more time and compute
-- unsupervised segmentation would be cool
-- find ways to not have to smoosh images for memory?
-- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too
-
-began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
-- check out filter viz
-- see about unsupervised
-- try 3d alternative archs? (dont really need to)
-
-might want to improve 3d before going on to segmentation,
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/saliency.png}
-  \caption{caption}
-  \label{fig:saliency}
-\end{figure}
-
-% ----------------------------------------------------------------------
-% END
-% ----------------------------------------------------------------------
-{\small
-\bibliographystyle{ieee}
-\bibliography{project}
-}
-
-% ----------------------------------------------------------------------
-% SUPPLEMENTARY
-% ----------------------------------------------------------------------
-\section{Supplementary Material}\label{sec:supplementary}
-
-Extra material for reference
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/orthoview_t1w.png}
-  \caption{caption}
-  \label{fig:ortho-t1w}
-\end{figure}
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
-  \caption{caption}
-  \label{fig:cortex-var}
-\end{figure}
-
-\end{document}

### 2019-12-03 paper/sections/background.tex
Breaking out sections to make working on paper easier

--- 2019-12-03 paper/submission.tex
@@
-\documentclass[10pt,twocolumn,letterpaper]{article}
+\usepackage{graphicx}
 
-\usepackage{cvpr}
-%\usepackage{times}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-% Include other packages here, before hyperref.
-
-% If you comment hyperref and then uncomment it, you should delete
-% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
-% run, let it finish, and you should be clear).
-\usepackage[breaklinks=true,bookmarks=false]{hyperref}
-
-\cvprfinalcopy % *** Uncomment this line for the final submission
-
-\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
-\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
-
-% Pages are numbered in submission mode, and unnumbered in camera-ready
-%\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{1}
-\begin{document}
-
-%-------------------------------------------------------------------------
-\title{Deep Learning for 3D Brain Activity Scans}
-
-\author{Anonymized for Peer Review\\
-University of Massachusetts Amherst\\
-140 Governors Dr, Amherst, MA 01002\\
-{\tt\small @umass.edu}
-}
-
-\maketitle
-%\thispagestyle{empty}
-
-% ----------------------------------------------------------------------
-% ABSTRACT
-%-------------------------------------------------------------------------
-\begin{abstract}
- Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
- typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
- and limit the ability to stitch together patterns across a third dimension.
-
- Deep learning's ability to decipher and learn from such nearly impossible patterns in massive datasets,
- imperceptible to preliminary observation makes it an excellent candidate to explore and teach us about
- overlooked patterns of activity.
-
- Here I showcase convolutional network's ability to correctly classify uncleaned activity data from a passive
- image viewing experiment aimed at triggering activity in particular regions of the brain controlling reward and
- restraint, using the data in both 2d and 3d to explore the merits of adapting video-processing techniques meant
- to apply across time to a third spatial dimension and discuss adapting this to unsupervised segmentation to see
- what the network can teach us from its findings.
-
-\end{abstract}
-
-% ----------------------------------------------------------------------
-% INTRODUCTION
-% ----------------------------------------------------------------------
-\section{Introduction}\label{sec:introduction}
-Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
-for mental health research and disease diagnosis.
-High dimensional, massive datasets, with intertwining patterns of activity and structure in three or four
-dimensions, are a lot to manage by hand.
-
-Classification of activity and structures should come quite naturally with deep learning practices.
-Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
-yet often only as a proof of concept, and limited to two dimensions.
-Others are working now to try and find universal architectures for most
-neuro-imaging tasks~\cite{henschel2019fastsurfer}, though it seems this is still far from a solved
-application with satisfactory performances across the board.
-
-Of more interest to myself, deep learning networks in their proven capacity to segment images based on what entities
-they identify in classification could surely be adopted to highlight what patterns of activity have informed their
-own decision making and perhaps clue researchers in to overlooked details.
-
-Here we'll discuss creating highly accurate classifiers using both 2d and 3d convolutional networks
-to learn in the typical manner low level filters, here specifically created for brain scans of spatial activity, using
-no pretraining on unrelated datasets and no neuro-imaging feature analysis or reduction~\cite{shi2018feature},
-letting the model learn for itself what is important so that it can teach the observer instead.
-
-Following this we'll discuss what this may mean for future works in the highly valuable ability to adapt the classifiers
-to the new, related task of unsupervised segmentation in 3 dimensions~\cite{shu2016unsupervised} to mirror what has been
-a fruitful field of 2d identification of tumors~\cite{akkus2017deep} for this specific application of telling brain
-activity patterns.
-
-% ----------------------------------------------------------------------
-% BACKGROUND / RELATED
-% ----------------------------------------------------------------------
 \section{Background}\label{sec:background}
@@
 investigating the capacity for unsupervised differentiation.
+
+\subsection{Data Format}\label{subsec:data-format}
 
@@
 
-You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.
-
-You may find this project repository here: (personal github link removed for anonymity)
-Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
-abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/orthoview_t1w.png}
+  \caption{T1w Modality High-Res 3d Scan of Subject Brain}
+  \label{fig:ortho-t1w}
+\end{figure}
 
@@
 
-% ----------------------------------------------------------------------
-% APPROACH
-% ----------------------------------------------------------------------
-\section{Approach}\label{sec:approach}
+\subsection{Review}\label{subsec:review}
 
-Beginning with SKLearn~\cite{scikit-learn} data readiness was proven with moderately successful classification using
-a simple SVM and then the SKLearn Multi-Layer-Perceptron neural network, getting 70-80\% accuracy, sufficing the
-initial sanity check that deep learning would be ready to work on.
+You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.
 
-% TODO these are relevant citations, right?
-Leveraging PyTorch~\cite{paszke2017automatic} first a framework for learning was likewise vetted using a simple fully
-connected shallow network, graduating to the state of the art convolutional networks, chiefly focussed
-on the ResNet~\cite{he2016deep} architecture styles, which have been developed for video classification~\cite{tran2018closer}
-utilizing 3 dimensional convolutions~\cite{hara3dcnns} much in the way 2 dimensional convolutions operate, which have
-been shown to hold a slight edge on 2 dimensions~\cite{payan2015predicting} in other applications.
-
-\subsection{Data Preparation}\label{subsec:data-preparation}
-
-First as the data was not prepared for classification, critical cleaning include subtracting the mean image to better
-highlight meaningful variance, and standardizing the data between 0 and 1.
-
-Given the nature of the input images being 3d brain scans it's intuitive that examining the brain in 3 dimensions
-allows for much more interesting patterns of brain activity and structuring being recognized.
-
-This comes with a substantial cost however, as the extra dimension explodes the memory demands of the input data, and
-the number of trainable parameters for the models.
-Many of the recommendations following will vary based on the particular hardware, but with a 4GB GPU I had to be quite
-careful about batch-sizes to not over-allocate memory.
-To combat this, first, a relatively conservative amount of blank space is cropped from the images, halving the total footprint.
-Following this, in order to use a batch-size of 64, I interpolate images to approximately $\frac{5}{6}$ their original dimension
-
-In addition, I typically do not use the full cohort of 30 subjects, as I even encounter some RAM restrictions with the
-full raw dataset consisting of
-$64 sagital \times 64 coronal \times 30 axial \times 32bit floats \times 375 scans \times 30 subjects$ or roughly 5.53 gigabytes
-
-In working with 2 dimensions, omitting the axial dimension for a flat cross section, you'll see the dataset is
-$\frac{1}{30}$ the size and hence, much more manageable.
-
-\subsection{Examining the Data}\label{subsec:examining-data}
-
-Two fully processed, correctly classified brain scans are presented in the figure~\ref{fig:comparison}
-which pose a daunting challenge to human labelers.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{images/nf_f_hard_comparison.png}
-  \caption{Nonfood-Stimuli \& Food-Stimuli Respectively from the same Subject}
-  \label{fig:comparison}
-\end{figure}
-
-\subsection{Model Selection}\label{subsec:model-selection}
-
-For the primary focus of comparing and improving classification accuracy using 2d and 3d architectures, ResNet models
-of the standard 18 layers to 100+ layers are examined, as the ResNet family is both highly competitive, and demonstrated
-in both 2 and 3 dimensions which will help comparisons in exploring the benefits of 3d brain scan usage.
-Video processing typically limits the layers to a more shallow level than that of image processing, as unfortunately as
-with the data, the model parameters explode out of control pretty fast.
-The following parameter counts are from standard ResNet architectures adapted to a single input channel.
-
-\begin{center}
- \begin{tabular}{||c c c||}
- \hline
- - & ResNet-18 2d & ResNet-18 3d \\ [0.5ex]
- \hline\hline
- Counts & 11,171,266 & 33,148,482 \\ [1ex]
- \hline
- Ratio & 1.0 & 2.97 \\ [1ex]
- \hline
-\end{tabular}
-\end{center}
-
-\subsection{Comparisons}\label{subsec:comparisons}
-
-Model performance is typically evaluated using accuracy, precision, recall, and AUC metrics, as well as total time to
-train.
-
-% ----------------------------------------------------------------------
-% RESULTS
-% ----------------------------------------------------------------------
-\section{Results}\label{sec:results}
-At the date of the milestone, I have shown that my experimental set-up is complete to begin
-iterating towards the novel contributions aimed for by this project.
-
-I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
-network for classification on my chosen dataset incorporating my own contributions to the process.
-
-I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
-excited nevertheless to try.
-
-\subsection{Data}\label{subsec:data}
-I have become familiar with the chosen dataset, and created a machine learning friendly format
-based off of the data.
-
-I have inferred classification labels from the experimental events timings and the estimated time
-of each complete scan in the dataset.
-
-\subsection{Model Architecture}\label{subsec:model-arch}
-
-the project explored minimally the hyperparameters of the chosen model architectures
-chiefly was finding the right learning rate for the given normalized data to the model
-typically 1e-2 to 1e-4 worked, with 1e-3 being good, and decay being very very helpful
-model quite sensitive to adjustments
-
-- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
-
-\subsection{Scope of Number of Subjects}\label{subsec:num-subjects}
-
-inclusion of more ppts/subjects means more training time (val/test are of same patients,
-hence more individual brain shapes means more general, but longer to learn)
-20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/2d_cold.png}
-  \caption{Cold-Start 2D Training Progress}
-  \label{fig:2d_cold}
-\end{figure}
-
-\subsection{Warm Starts}\label{subsec:warm-starts}
-saving model parameter dict, loading and training again
-wowee good stuff, hope to use a LR scheduler in the future to try to control this better?
-surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
-but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
-bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/2d_warm.png}
-  \caption{Warm-Start 2D Training Progress}
-  \label{fig:2d_warm}
-\end{figure}
-
-\subsection{3 Dimensions}\label{subsec:3-dimensions}
-- 3d CNN - typically meant for video classification using time as the third axis,
-here we are not using time, though that is a dimension we have access to,
-though doesn't make sense for this experiment/project to try to add in
-- https://pytorch.org/docs/stable/torchvision/models.html#video-classification
-
-okay it works, and has potential, certainly given the data is richer to be the best, but wow it's frustratingly slow
-findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
-promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
-unless better hardware or serious effort is expended in getting it differentiable
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/3d_cold.png}
-  \caption{Cold-Start 3D Training Progress}
-  \label{fig:3d_cold}
-\end{figure}
-wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
-different styles of 1d (unfurl 3d, 2d) 2d (mean flattened, one slice) 3d (scale, orig)
-LR explorations
-
-future works
-better gpu / computer might be important, the 3d models are real fucking beasts to train, and their data is large as fuck
-model ensemble (free extra few percent perf)
-
-"none of my alterations to the model architecture worked that well, most did nothing, a few broke things, could still poke around more though"
-graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
-batch_size and interpolation hard to juggle too?
-3d feels like "luck" plays a big part of getting training to catch/start
-seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
-
-cut batch size to avoid scaling down images (TRIED NO BUENO?)
-
-\subsection{Performance}\label{subsec:performance}
-
- % P-R-F1-S
-\begin{center}
- \begin{tabular}{||c c c||}
- \hline
- - & Nonfood & Food \\ [0.5ex]
- \hline\hline
-Precision & 0.949074 & 0.958904 \\
- \hline
-Recall & 0.957944 & 0.950226 \\
- \hline
-F1 & 0.953488 & 0.954545 \\
- \hline
-Support & 214 & 221 \\ [1ex]
- \hline
-\end{tabular}
-
- % ROC-AUC / ACC
-\end{center}
-\begin{center}
- \begin{tabular}{||c c||}
- \hline
- ROC-AUC & Accuracy \\ [0.5ex]
- \hline\hline
-  0.9847 & 0.954 \\ [1ex]
- \hline
-\end{tabular}
-\end{center}
-
-Can get a little higher but these are already encouraging numbers (96 seen) often the model starts leaning a little
-one way or the other and this was one of the most balanced runs I've seen
-
-I'm a little unclear about the reliability of my data labeling, so I'm happy with this result, I'd be stoked
-to reproduce on 3d, and trusting the model this much, open the door to segmentation
-
-% ----------------------------------------------------------------------
-% DISCUSSION
-% ----------------------------------------------------------------------
-\section{Discussion}\label{sec:discussion}
-
-\subsection{Learning}\label{subsec:learning}
-very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
-cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
-this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at
-
-\subsection{Future Works}\label{subsec:future-works}
-Unsupervised segmentation, especially in 3d would be cool
-Future works
-- could strive to refine 3d more, with more time and compute
-- unsupervised segmentation would be cool
-- find ways to not have to smoosh images for memory?
-- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too
-
-began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
-- check out filter viz
-- see about unsupervised
-- try 3d alternative archs? (dont really need to)
-
-might want to improve 3d before going on to segmentation,
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/saliency.png}
-  \caption{caption}
-  \label{fig:saliency}
-\end{figure}
-
-% ----------------------------------------------------------------------
-% END
-% ----------------------------------------------------------------------
-{\small
-\bibliographystyle{ieee}
-\bibliography{project}
-}
-
-% ----------------------------------------------------------------------
-% SUPPLEMENTARY
-% ----------------------------------------------------------------------
-\section{Supplementary Material}\label{sec:supplementary}
-
-Extra material for reference
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/orthoview_t1w.png}
-  \caption{caption}
-  \label{fig:ortho-t1w}
-\end{figure}
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
-  \caption{caption}
-  \label{fig:cortex-var}
-\end{figure}
-
-\end{document}
+You may find this project repository here: (personal github link removed for anonymity)
+Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
+abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.

### 2019-12-03 paper/sections/discussion.tex
Breaking out sections to make working on paper easier

--- 2019-12-02 paper/submission.tex
@@
-\documentclass[10pt,twocolumn,letterpaper]{article}
+\usepackage{graphicx}
 
-\usepackage{cvpr}
-%\usepackage{times}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
+\section{Discussion}\label{sec:discussion}
 
-% Include other packages here, before hyperref.
+\subsection{Learning}\label{subsec:learning}
+very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
+cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
+this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at
 
-% If you comment hyperref and then uncomment it, you should delete
-% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
-% run, let it finish, and you should be clear).
-\usepackage[breaklinks=true,bookmarks=false]{hyperref}
+\subsection{Future Works}\label{subsec:future-works}
+Unsupervised segmentation, especially in 3d would be cool
+Future works
+- could strive to refine 3d more, with more time and compute
+- unsupervised segmentation would be cool
+- find ways to not have to smoosh images for memory?
+- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too
 
-\cvprfinalcopy % *** Uncomment this line for the final submission
+began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
+- check out filter viz
+- see about unsupervised
+- try 3d alternative archs? (dont really need to)
 
-\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
-\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
+might want to improve 3d before going on to segmentation,
 
-% Pages are numbered in submission mode, and unnumbered in camera-ready
-%\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{1}
-\begin{document}
-
-%-------------------------------------------------------------------------
-\title{Deep Learning for 3D Brain Activity Scans}
-
-\author{Evan Albright\\
-University of Massachusetts Amherst\\
-140 Governors Dr, Amherst, MA 01002\\
-{\tt\small ealbright@umass.edu}
-}
-
-\maketitle
-%\thispagestyle{empty}
-
-%-------------------------------------------------------------------------
-\begin{abstract}
-   Abstract placeholder (sub 300 words)
-\end{abstract}
-
-% ----------------------------------------------------------------------
-% INTRODUCTION
-% ----------------------------------------------------------------------
-\section{Introduction}\label{sec:introduction}
-Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
-for mental health research and disease diagnosis.
-High dimensional, massive datasets, with intertwining patterns of activity and structure in three or four
-dimensions, are a lot to manage by hand.
-
-Classification of activity and structures should come quite naturally with deep learning practices.
-Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
-yet often only as a proof of concept.
-Others are working now to try and find universal architectures for most
-neuro-imaging tasks~\cite{henschel2019fastsurfer}, though it seems this is still far from a solved
-application with satisfactory performances across the board.
-
-Of more interest to myself, deep learning networks in their proven capacity to segment images based on what entities
-they identify in classification could surely be adopted to highlight what patterns of activity have informed their
-own decision making and perhaps clue researchers in to overlooked details.
-
-My ambition is to create a highly accurate 3d convolutional network to first classify, then segment brain scans to
-help with my specific application of neuroimaging.
-
-I may also attempt to explore feature analysis and reduction which has been shown to
-help~\cite{shi2018feature} with deep learning tasks, if only to manage the challenging
-dimensionality of their 3d inputs.
-
-% ----------------------------------------------------------------------
-% PROBLEM STATEMENT
-% ----------------------------------------------------------------------
-\section{Problem Statement}\label{sec:problem-statement}
-
-(BACKGROUND / RELATED)
-
-The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
-Netherlands circa 2013 who were presented with various grids of images featuring
-either neutral entities, (e.g.\ office chairs, windows, jackets) or foods~\cite{smeets2013allured}.
-The foods were specifically types of calorically dense, tempting foods, which the individuals viewing would likely
-have formed restraint with given their success in maintaining healthy body mass indices.
-The experiment was meant to highlight brain structures roles in anticipated rewards from consuming delicious foods,
-balanced against the hypothesized activation of restraints.
-In neuro-imaging, this task is known as passive image viewing.
-
-Given the quite important role of food and willpower in our lives, the activation patterns are likely to be powerful
-and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
-This combination gives me high hopes of success in this specific application of learning to classify the viewed
-objects through brain activations of the subjects.
-
-I hope given the clarity in patterns in the brain scans, that I can develop a neural network to highlight what
-features and locations in the brain scans have been helpful in informing the decisions of classification, in the
-hope that such methods could be adapted to harder tasks in the future.
-
-The dataset can be found from the excellent
-OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}
-It consists of a survey detailing the relationship of each subject with food and diet taken after the experimental
-proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w brain scan,
-and time series of bold scans of about 370-375 scans, which encompasses the approximately 10 minute experiment
-in its entirety.
-The bold scans are quite sizeable given the number of them, however they are a relatively low resolution, consisting
-of 64 sagital by 64 coronal by 30 axial scans.
-
-You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.
-
-You may find this project repository here: \url{https://github.com/e-m-albright/CS682}
-Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
-abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.
-
-% ----------------------------------------------------------------------
-% TECHNICAL APPROACH
-% ----------------------------------------------------------------------
-\section{Technical Approach}\label{sec:technical-approach}
-I intend to explore the efficacy of 3d convolutional networks which have been shown to have a slight performance
-edge on 2d convolutions~\cite{payan2015predicting}.
-Given the nature of the input images being 3d brain scans this makes intuitive sense and could allow for much more
-interesting patterns of brain activity and structuring being recognized.
-
-As for the deeper architecture, I'm hoping to imitate some of the more successful networks like VGG or ResNet,
-however I wouldn't be surprised if they don't adapt well to this particular application, especially given how
-interesting and important it could be to keep the input data in a three dimensional format.
-
-It is my hope that I can also learn from other developments of object detection to see what regions in the
-brain become active for a particular classification, like Faster R-CNN
-
-\subsection{Novelty}\label{subsec:novelty}
-
-I have not seen any application of classification or object segmentation to neuroimaging in 3d
-on brain activity patterns.
-I have seen some works which I hope to leverage with similar aims like that of structural identification
-for tumors, in two dimensions~\cite{akkus2017deep}.
-However at the very least I hope to produce a very successful network on my chosen problem first and foremost rather
-than assume I am the only person to ever think of something as intuitive as bringing neural networks to bear
-on neuro-imaging tasks.
-
-I hope to find a methodology of first classifying correctly the brain scans, improve accuracy as much as possible,
-and adapt the approach into tools for identifying important regions of activity in the
-brain which are being utilized for classification.
-
-I am not chiefly focused on coping with the challenges of memory but it may become too important to overlook.
-Each dataset input is a $64 \times 64 \times 30$ slice of floats, for which we've got
-about $375 scans \times 30 subjects$ total.
-In my work on preliminary results I have already been confronted by the limitations of my current machine and have
-been trying working around the issue until I am ready to attempt to resolve it.
-This may be an unavoidable area to simultaneously try to innovate for the field of
-neuro-imaging analysis with deep learning
-
-% ----------------------------------------------------------------------
-% INTERMEDIATE RESULTS
-% ----------------------------------------------------------------------
-\section{Intermediate Results}\label{sec:intermediate-results}
-At the date of the milestone, I have shown that my experimental set-up is complete to begin
-iterating towards the novel contributions aimed for by this project.
-
-I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
-network for classification on my chosen dataset incorporating my own contributions to the process.
-
-I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
-excited nevertheless to try.
-
-\subsection{Data}\label{subsec:data}
-I have become familiar with the chosen dataset, and created a machine learning friendly format
-based off of the data.
-
-I have inferred classification labels from the experimental events timings and the estimated time
-of each complete scan in the dataset.
-
-\subsection{Sanity Checks}\label{subsec:sanity-checks}
-Using the data, I have implemented a very simple, yet effective 80\% accurate SVM classifier to prove the data
-should be ready for use with my chosen deep learning methodologies.
-
-I'm now confident in the classifications of each of the scans (about 375 per subject, for 30 subjects)
-
-\subsection{Deep Learning \& Future Works}\label{subsec:deep-learning}
-I have tried feeding the data into very basic neural networks with mixed success, and am in the active process of
-troubleshooting.
-Once I have matched performance with the SVM using an implementation of a reasonable neural network using PyTorch, I
-will begin the process of looking for refinements and innovations alluded to in previous sections, as well as
-transitioning the task to object detection if I can.
-
-% ----------------------------------------------------------------------
-% TODO
-% ----------------------------------------------------------------------
-\section{TODO}\label{sec:todo}
-
-NN project
-- 2d hyperparameters, let's use a LR scheduler? LR spikes all over the place still, why?
-
-- 3d CNN - typically meant for video classification using time as the third axis,
-here we are not using time, though that is a dimension we have access to,
-though doesn't make sense for this experiment/project to try to add in
-- https://pytorch.org/docs/stable/torchvision/models.html#video-classification
-
-- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
-- cut batch size to avoid scaling down images (TRIED NO BUENO?)
-
-be sure to talk about everything in paper
-
-normalization/mean image
-interpolation
-wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
-different styles of 1d (unfurl 3d, 2d) 2d (mean flattened, one slice) 3d (scale, orig)
-LR explorations
-sanity checks on brain image data, impossible for me to spot the differences
-
-batch_size and interpolation hard to juggle too?
-3d feels like "luck" plays a big part of getting training to catch/start
-seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
-
-images:
-LR / accuracy measures, of diff approaches
-hi res, mid res, interpolated down res brain scan examples,
-2d slice example
-total training time? num features comparison
-average slice image
-graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
-
-inclusion of more ppts/subjects means more training time (val/test are of same patients,
-hence more individual brain shapes means more general, but longer to learn)
-20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
-
-image segmentation attempt?
-saliency maps?
-filter views?
-
-torch save models - can eval that way
-host for downloads?
-
-future works
-better gpu / computer might be important, the 3d models are real fucking beasts to train, and their data is large as fuck
-model ensemble (free extra few percent perf)
-
-findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
-promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
-unless better hardware or serious effort is expended in getting it differentiable
-
-PUT BRAIN SCANS IN SUPPLEMENTARY?
-
-abstract > intro > background > related
-methods/approach > results > discussion > references
-6 pages? 7 maybe at most
-
-gradual warm-start improvements?
-variations on models? meh
-
-unsupervised segmentation - little hard
-
-surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
-but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
-bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential
-
-WORK ON GETTING PARAM NUMBERS AND
-DIFFERENT METRICS (AUC PLEASE, F1 PLEASE)
-2d is enough
-
-Submit your final submission through OpenReview. You will submit a single pdf file:
-A pdf file of your final report
-
-(OPTIONAL) supplementary can be concatenated at the end of the final report
-Report. The following is a suggested structure for the report:
-Title, Author(s)
-
-Abstract: It should not be more than 300 words
-
-Introduction: this section introduces your problem, and the overall plan for approaching your problem
-Background/Related Work: This section discusses relevant literature for your project
-Approach: This section details the framework of your project. Be specific, which means you might want to include equations, figures, plots, etc
-Experiment: This section begins with what kind of experiments you're doing, what kind of dataset(s) you're using, and what is the way you measure or evaluate your results.
-It then shows in details the results of your experiments.
-By details, we mean both quantitative evaluations (show numbers, figures, tables, etc) as well as qualitative results (show images, example results, etc).
-Conclusion: What have you learned? Suggest future ideas.
-References: This is absolutely necessary.
-Supplementary Material is not counted toward your 6-8 page limit. It is optional and is supposed to contain less important results/experiments/etc not critical for understanding the main report.
-
-try different architectures - more depth?
-see if we can break 85% to 90%, talk about potential data mislabeling frustration as a barrier to higher results possibly
-
-you are expected to show substantially more experiments and results with NN-related methods
-( :(  lame)
-
-~\cite{he2016deep} resnet
-~\cite{tran2018closer} 3d/time resnet
-~\cite{hara3dcnns} 3d cnn pt1?
-
-% ----------------------------------------------------------------------
-% END
-% ----------------------------------------------------------------------
-{\small
-\bibliographystyle{ieee}
-\bibliography{project}
-}
-
-\end{document}
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/saliency.png}
+  \caption{caption}
+  \label{fig:saliency}
+\end{figure}

### 2019-12-03 paper/sections/intro.tex
Breaking out sections to make working on paper easier

--- 2019-12-02 paper/submission.tex
@@
-\documentclass[10pt,twocolumn,letterpaper]{article}
+\section{Introduction}\label{sec:introduction}
 
-\usepackage{cvpr}
-%\usepackage{times}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-% Include other packages here, before hyperref.
-
-% If you comment hyperref and then uncomment it, you should delete
-% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
-% run, let it finish, and you should be clear).
-\usepackage[breaklinks=true,bookmarks=false]{hyperref}
-
-\cvprfinalcopy % *** Uncomment this line for the final submission
-
-\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
-\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
-
-% Pages are numbered in submission mode, and unnumbered in camera-ready
-%\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{1}
-\begin{document}
-
-%-------------------------------------------------------------------------
-\title{Deep Learning for 3D Brain Activity Scans}
-
-\author{Evan Albright\\
-University of Massachusetts Amherst\\
-140 Governors Dr, Amherst, MA 01002\\
-{\tt\small ealbright@umass.edu}
-}
-
-\maketitle
-%\thispagestyle{empty}
-
-%-------------------------------------------------------------------------
-\begin{abstract}
-   Abstract placeholder (sub 300 words)
-\end{abstract}
-
-% ----------------------------------------------------------------------
-% INTRODUCTION
-% ----------------------------------------------------------------------
-\section{Introduction}\label{sec:introduction}
 Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
@@
 
-My ambition is to create a highly accurate 3d convolutional network to first classify, then segment brain scans to
-help with my specific application of neuroimaging.
+Here we'll discuss creating highly accurate classifiers using both 2d and 3d convolutional networks
+to learn in the typical manner low level filters, here specifically created for brain scans of spatial activity, using
+no pretraining on unrelated datasets and no neuro-imaging feature analysis or reduction~\cite{shi2018feature},
+letting the model learn for itself what is important so that it can teach the observer instead.
 
-I may also attempt to explore feature analysis and reduction which has been shown to
-help~\cite{shi2018feature} with deep learning tasks, if only to manage the challenging
-dimensionality of their 3d inputs.
-
-% ----------------------------------------------------------------------
-% PROBLEM STATEMENT
-% ----------------------------------------------------------------------
-\section{Problem Statement}\label{sec:problem-statement}
-
-(BACKGROUND / RELATED)
-
-The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
-Netherlands circa 2013 who were presented with various grids of images featuring
-either neutral entities, (e.g.\ office chairs, windows, jackets) or foods~\cite{smeets2013allured}.
-The foods were specifically types of calorically dense, tempting foods, which the individuals viewing would likely
-have formed restraint with given their success in maintaining healthy body mass indices.
-The experiment was meant to highlight brain structures roles in anticipated rewards from consuming delicious foods,
-balanced against the hypothesized activation of restraints.
-In neuro-imaging, this task is known as passive image viewing.
-
-Given the quite important role of food and willpower in our lives, the activation patterns are likely to be powerful
-and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
-This combination gives me high hopes of success in this specific application of learning to classify the viewed
-objects through brain activations of the subjects.
-
-I hope given the clarity in patterns in the brain scans, that I can develop a neural network to highlight what
-features and locations in the brain scans have been helpful in informing the decisions of classification, in the
-hope that such methods could be adapted to harder tasks in the future.
-
-The dataset can be found from the excellent
-OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}
-It consists of a survey detailing the relationship of each subject with food and diet taken after the experimental
-proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w brain scan,
-and time series of bold scans of about 370-375 scans, which encompasses the approximately 10 minute experiment
-in its entirety.
-The bold scans are quite sizeable given the number of them, however they are a relatively low resolution, consisting
-of 64 sagital by 64 coronal by 30 axial scans.
-
-You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.
-
-You may find this project repository here: \url{https://github.com/e-m-albright/CS682}
-Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
-abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.
-
-% ----------------------------------------------------------------------
-% TECHNICAL APPROACH
-% ----------------------------------------------------------------------
-\section{Technical Approach}\label{sec:technical-approach}
-I intend to explore the efficacy of 3d convolutional networks which have been shown to have a slight performance
-edge on 2d convolutions~\cite{payan2015predicting}.
-Given the nature of the input images being 3d brain scans this makes intuitive sense and could allow for much more
-interesting patterns of brain activity and structuring being recognized.
-
-As for the deeper architecture, I'm hoping to imitate some of the more successful networks like VGG or ResNet,
-however I wouldn't be surprised if they don't adapt well to this particular application, especially given how
-interesting and important it could be to keep the input data in a three dimensional format.
-
-It is my hope that I can also learn from other developments of object detection to see what regions in the
-brain become active for a particular classification, like Faster R-CNN
-
-\subsection{Novelty}\label{subsec:novelty}
-
-I have not seen any application of classification or object segmentation to neuroimaging in 3d
-on brain activity patterns.
-I have seen some works which I hope to leverage with similar aims like that of structural identification
-for tumors, in two dimensions~\cite{akkus2017deep}.
-However at the very least I hope to produce a very successful network on my chosen problem first and foremost rather
-than assume I am the only person to ever think of something as intuitive as bringing neural networks to bear
-on neuro-imaging tasks.
-
-I hope to find a methodology of first classifying correctly the brain scans, improve accuracy as much as possible,
-and adapt the approach into tools for identifying important regions of activity in the
-brain which are being utilized for classification.
-
-I am not chiefly focused on coping with the challenges of memory but it may become too important to overlook.
-Each dataset input is a $64 \times 64 \times 30$ slice of floats, for which we've got
-about $375 scans \times 30 subjects$ total.
-In my work on preliminary results I have already been confronted by the limitations of my current machine and have
-been trying working around the issue until I am ready to attempt to resolve it.
-This may be an unavoidable area to simultaneously try to innovate for the field of
-neuro-imaging analysis with deep learning
-
-% ----------------------------------------------------------------------
-% INTERMEDIATE RESULTS
-% ----------------------------------------------------------------------
-\section{Intermediate Results}\label{sec:intermediate-results}
-At the date of the milestone, I have shown that my experimental set-up is complete to begin
-iterating towards the novel contributions aimed for by this project.
-
-I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
-network for classification on my chosen dataset incorporating my own contributions to the process.
-
-I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
-excited nevertheless to try.
-
-\subsection{Data}\label{subsec:data}
-I have become familiar with the chosen dataset, and created a machine learning friendly format
-based off of the data.
-
-I have inferred classification labels from the experimental events timings and the estimated time
-of each complete scan in the dataset.
-
-\subsection{Sanity Checks}\label{subsec:sanity-checks}
-Using the data, I have implemented a very simple, yet effective 80\% accurate SVM classifier to prove the data
-should be ready for use with my chosen deep learning methodologies.
-
-I'm now confident in the classifications of each of the scans (about 375 per subject, for 30 subjects)
-
-\subsection{Deep Learning \& Future Works}\label{subsec:deep-learning}
-I have tried feeding the data into very basic neural networks with mixed success, and am in the active process of
-troubleshooting.
-Once I have matched performance with the SVM using an implementation of a reasonable neural network using PyTorch, I
-will begin the process of looking for refinements and innovations alluded to in previous sections, as well as
-transitioning the task to object detection if I can.
-
-% ----------------------------------------------------------------------
-% TODO
-% ----------------------------------------------------------------------
-\section{TODO}\label{sec:todo}
-
-NN project
-- 2d hyperparameters, let's use a LR scheduler? LR spikes all over the place still, why?
-
-- 3d CNN - typically meant for video classification using time as the third axis,
-here we are not using time, though that is a dimension we have access to,
-though doesn't make sense for this experiment/project to try to add in
-- https://pytorch.org/docs/stable/torchvision/models.html#video-classification
-
-- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
-- cut batch size to avoid scaling down images (TRIED NO BUENO?)
-
-be sure to talk about everything in paper
-
-normalization/mean image
-interpolation
-wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
-different styles of 1d (unfurl 3d, 2d) 2d (mean flattened, one slice) 3d (scale, orig)
-LR explorations
-sanity checks on brain image data, impossible for me to spot the differences
-
-batch_size and interpolation hard to juggle too?
-3d feels like "luck" plays a big part of getting training to catch/start
-seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
-
-images:
-LR / accuracy measures, of diff approaches
-hi res, mid res, interpolated down res brain scan examples,
-2d slice example
-total training time? num features comparison
-average slice image
-graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
-
-inclusion of more ppts/subjects means more training time (val/test are of same patients,
-hence more individual brain shapes means more general, but longer to learn)
-20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
-
-image segmentation attempt?
-saliency maps?
-filter views?
-
-torch save models - can eval that way
-host for downloads?
-
-future works
-better gpu / computer might be important, the 3d models are real fucking beasts to train, and their data is large as fuck
-model ensemble (free extra few percent perf)
-
-findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
-promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
-unless better hardware or serious effort is expended in getting it differentiable
-
-PUT BRAIN SCANS IN SUPPLEMENTARY?
-
-abstract > intro > background > related
-methods/approach > results > discussion > references
-6 pages? 7 maybe at most
-
-gradual warm-start improvements?
-variations on models? meh
-
-unsupervised segmentation - little hard
-
-surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
-but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
-bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential
-
-WORK ON GETTING PARAM NUMBERS AND
-DIFFERENT METRICS (AUC PLEASE, F1 PLEASE)
-2d is enough
-
-Submit your final submission through OpenReview. You will submit a single pdf file:
-A pdf file of your final report
-
-(OPTIONAL) supplementary can be concatenated at the end of the final report
-Report. The following is a suggested structure for the report:
-Title, Author(s)
-
-Abstract: It should not be more than 300 words
-
-Introduction: this section introduces your problem, and the overall plan for approaching your problem
-Background/Related Work: This section discusses relevant literature for your project
-Approach: This section details the framework of your project. Be specific, which means you might want to include equations, figures, plots, etc
-Experiment: This section begins with what kind of experiments you're doing, what kind of dataset(s) you're using, and what is the way you measure or evaluate your results.
-It then shows in details the results of your experiments.
-By details, we mean both quantitative evaluations (show numbers, figures, tables, etc) as well as qualitative results (show images, example results, etc).
-Conclusion: What have you learned? Suggest future ideas.
-References: This is absolutely necessary.
-Supplementary Material is not counted toward your 6-8 page limit. It is optional and is supposed to contain less important results/experiments/etc not critical for understanding the main report.
-
-try different architectures - more depth?
-see if we can break 85% to 90%, talk about potential data mislabeling frustration as a barrier to higher results possibly
-
-you are expected to show substantially more experiments and results with NN-related methods
-( :(  lame)
-
-~\cite{he2016deep} resnet
-~\cite{tran2018closer} 3d/time resnet
-~\cite{hara3dcnns} 3d cnn pt1?
-
-% ----------------------------------------------------------------------
-% END
-% ----------------------------------------------------------------------
-{\small
-\bibliographystyle{ieee}
-\bibliography{project}
-}
-
-\end{document}
+Following this we'll discuss what this may mean for future works in the highly valuable ability to adapt the classifiers
+to the new, related task of unsupervised segmentation in 3 dimensions~\cite{shu2016unsupervised} to mirror what has been
+a fruitful field of 2d identification of tumors~\cite{akkus2017deep} for this specific application of telling brain
+activity patterns.

### 2019-12-03 paper/sections/results.tex
Breaking out sections to make working on paper easier

--- 2019-12-03 paper/submission.tex
@@
-\documentclass[10pt,twocolumn,letterpaper]{article}
+\section{Results}\label{sec:results}
 
-\usepackage{cvpr}
-%\usepackage{times}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
+With a complete experimental set-up, benchmark performance was established using the simple models ranging from the
+SVM to the basic fully-connected network, at around 70-80\%.
+Both 2 dimensional and 3 dimensional convolutional networks were able to surpass the benchmarks,
+attaining 95\%+ and 90\%+ accuracy respectively.
 
-% Include other packages here, before hyperref.
-
-% If you comment hyperref and then uncomment it, you should delete
-% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
-% run, let it finish, and you should be clear).
-\usepackage[breaklinks=true,bookmarks=false]{hyperref}
-
-\cvprfinalcopy % *** Uncomment this line for the final submission
-
-\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
-\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
-
-% Pages are numbered in submission mode, and unnumbered in camera-ready
-%\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{1}
-\begin{document}
-
-%-------------------------------------------------------------------------
-\title{Deep Learning for 3D Brain Activity Scans}
-
-\author{Anonymized for Peer Review\\
-University of Massachusetts Amherst\\
-140 Governors Dr, Amherst, MA 01002\\
-{\tt\small @umass.edu}
-}
-
-\maketitle
-%\thispagestyle{empty}
-
-% ----------------------------------------------------------------------
-% ABSTRACT
-%-------------------------------------------------------------------------
-\begin{abstract}
- Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
- typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
- and limit the ability to stitch together patterns across a third dimension.
-
- Deep learning's ability to decipher and learn from such nearly impossible patterns in massive datasets,
- imperceptible to preliminary observation makes it an excellent candidate to explore and teach us about
- overlooked patterns of activity.
-
- Here I showcase convolutional network's ability to correctly classify uncleaned activity data from a passive
- image viewing experiment aimed at triggering activity in particular regions of the brain controlling reward and
- restraint, using the data in both 2d and 3d to explore the merits of adapting video-processing techniques meant
- to apply across time to a third spatial dimension and discuss adapting this to unsupervised segmentation to see
- what the network can teach us from its findings.
-
-\end{abstract}
-
-% ----------------------------------------------------------------------
-% INTRODUCTION
-% ----------------------------------------------------------------------
-\section{Introduction}\label{sec:introduction}
-Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
-for mental health research and disease diagnosis.
-High dimensional, massive datasets, with intertwining patterns of activity and structure in three or four
-dimensions, are a lot to manage by hand.
-
-Classification of activity and structures should come quite naturally with deep learning practices.
-Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
-yet often only as a proof of concept, and limited to two dimensions.
-Others are working now to try and find universal architectures for most
-neuro-imaging tasks~\cite{henschel2019fastsurfer}, though it seems this is still far from a solved
-application with satisfactory performances across the board.
-
-Of more interest to myself, deep learning networks in their proven capacity to segment images based on what entities
-they identify in classification could surely be adopted to highlight what patterns of activity have informed their
-own decision making and perhaps clue researchers in to overlooked details.
-
-Here we'll discuss creating highly accurate classifiers using both 2d and 3d convolutional networks
-to learn in the typical manner low level filters, here specifically created for brain scans of spatial activity, using
-no pretraining on unrelated datasets and no neuro-imaging feature analysis or reduction~\cite{shi2018feature},
-letting the model learn for itself what is important so that it can teach the observer instead.
-
-Following this we'll discuss what this may mean for future works in the highly valuable ability to adapt the classifiers
-to the new, related task of unsupervised segmentation in 3 dimensions~\cite{shu2016unsupervised} to mirror what has been
-a fruitful field of 2d identification of tumors~\cite{akkus2017deep} for this specific application of telling brain
-activity patterns.
-
-% ----------------------------------------------------------------------
-% BACKGROUND / RELATED
-% ----------------------------------------------------------------------
-\section{Background}\label{sec:background}
-
-The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
-Netherlands circa 2013 who were presented with various grids of images featuring
-either neutral entities, (e.g.\ office chairs, windows, jackets) or foods~\cite{smeets2013allured}.
-The foods were specifically types of calorically dense, tempting foods, which the individuals viewing would likely
-have formed restraint with given their success in maintaining healthy body mass indices.
-The experiment was meant to highlight brain structures roles in anticipated rewards from consuming delicious foods,
-balanced against the hypothesized activation of restraints.
-
-In neuro-imaging, this task is known as passive image viewing.
-
-Given the quite important role of food and willpower in our lives, the activation patterns are likely to be powerful
-and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
-This combination makes this an excellent selection for a proof of concept application of deep learning as the chances
-of success on a less subtle activation pattern, observable to trained researchers, are quite high.
-This specific application then can serve as a jumping off point for future works where the activity patterns are less
-distinguishable, and may be an interesting addition to the works in brain-pattern controlled prosthetics.
-This dataset is also ideal as it comes with a fairly usable ground-truth labeling mechanism, as I am not currently
-investigating the capacity for unsupervised differentiation.
-
-The dataset can be found from the excellent
-OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}
-
-The dataset follows the BIDS (Brain Imaging Data Structure) layout adopted by many neuroscience researchers.
-It consists of a survey detailing the relationship of each of the 30 subjects with food and diet taken after the experimental
-proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w Modality fMRI scan seen in figure~\ref{fig:ortho-t1w},
-and time series of 370-375 BOLD Modality fMRI scans seen in figure~\ref{fig:ortho-bold}, encompassing approximately 10 minutes
-capturing brain activity of the subjects during their exposure to the experimental triggers.
-
-The bold scans are quite sizeable as far as datasets go, given the number of each per subject,
-however they are a relatively low resolution, consisting
-of 64 sagital by 64 coronal by 30 axial scans, with only a single image channel.
-
-The activation is measured in a positive magnitude visible in figure~\ref{fig:ex_stat_map}, through a helper method in a neuro-imaging specialized
-image handling library ~\cite{brett_matthew_2019_3544468}.
-
-I had to manually infer pairings of the scans with the experimental phase for classification training given the events log associated
-with each subject and the advertised scan times from the paper.
-While the data does appear healthy, there may be some discrepancies of interpretation which would warrant future investigation if the
-performance on this particular experiment was the primary focus.
-
-You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.
-
-You may find this project repository here: (personal github link removed for anonymity)
-Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
-abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{images/orthoview_bold.png}
-  \caption{Bold Modality Low-Res 3d Experimental Scans of Subject Brain}
-  \label{fig:ortho-bold}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{images/example_stat_map.png}
-  \caption{Activity Measures from Bold Scans}
-  \label{fig:ex_stat_map}
-\end{figure}
-
-% ----------------------------------------------------------------------
-% APPROACH
-% ----------------------------------------------------------------------
-\section{Approach}\label{sec:approach}
-
-Beginning with SKLearn~\cite{scikit-learn} data readiness was proven with moderately successful classification using
-a simple SVM and then the SKLearn Multi-Layer-Perceptron neural network, getting 70-80\% accuracy, sufficing the
-initial sanity check that deep learning would be ready to work on.
-
-% TODO these are relevant citations, right?
-Leveraging PyTorch~\cite{paszke2017automatic} first a framework for learning was likewise vetted using a simple fully
-connected shallow network, graduating to the state of the art convolutional networks, chiefly focussed
-on the ResNet~\cite{he2016deep} architecture styles, which have been developed for video classification~\cite{tran2018closer}
-utilizing 3 dimensional convolutions~\cite{hara3dcnns} much in the way 2 dimensional convolutions operate, which have
-been shown to hold a slight edge on 2 dimensions~\cite{payan2015predicting} in other applications.
-
-\subsection{Data Preparation}\label{subsec:data-preparation}
-
-First as the data was not prepared for classification, critical cleaning include subtracting the mean image to better
-highlight meaningful variance, and standardizing the data between 0 and 1.
-
-Given the nature of the input images being 3d brain scans it's intuitive that examining the brain in 3 dimensions
-allows for much more interesting patterns of brain activity and structuring being recognized.
-
-This comes with a substantial cost however, as the extra dimension explodes the memory demands of the input data, and
-the number of trainable parameters for the models.
-Many of the recommendations following will vary based on the particular hardware, but with a 4GB GPU I had to be quite
-careful about batch-sizes to not over-allocate memory.
-To combat this, first, a relatively conservative amount of blank space is cropped from the images, halving the total footprint.
-Following this, in order to use a batch-size of 64, I interpolate images to approximately $\frac{5}{6}$ their original dimension
-
-In addition, I typically do not use the full cohort of 30 subjects, as I even encounter some RAM restrictions with the
-full raw dataset consisting of
-$64 sagital \times 64 coronal \times 30 axial \times 32bit floats \times 375 scans \times 30 subjects$ or roughly 5.53 gigabytes
-
-In working with 2 dimensions, omitting the axial dimension for a flat cross section, you'll see the dataset is
-$\frac{1}{30}$ the size and hence, much more manageable.
-
-\subsection{Examining the Data}\label{subsec:examining-data}
-
-Two fully processed, correctly classified brain scans are presented in the figure~\ref{fig:comparison}
-which pose a daunting challenge to human labelers.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{images/nf_f_hard_comparison.png}
-  \caption{Nonfood-Stimuli \& Food-Stimuli Respectively from the same Subject}
-  \label{fig:comparison}
-\end{figure}
-
-\subsection{Model Selection}\label{subsec:model-selection}
-
-For the primary focus of comparing and improving classification accuracy using 2d and 3d architectures, ResNet models
-of the standard 18 layers to 100+ layers are examined, as the ResNet family is both highly competitive, and demonstrated
-in both 2 and 3 dimensions which will help comparisons in exploring the benefits of 3d brain scan usage.
-Video processing typically limits the layers to a more shallow level than that of image processing, as unfortunately as
-with the data, the model parameters explode out of control pretty fast.
-The following parameter counts are from standard ResNet architectures adapted to a single input channel.
-
-\begin{center}
- \begin{tabular}{||c c c||}
- \hline
- - & ResNet-18 2d & ResNet-18 3d \\ [0.5ex]
- \hline\hline
- Counts & 11,171,266 & 33,148,482 \\ [1ex]
- \hline
- Ratio & 1.0 & 2.97 \\ [1ex]
- \hline
-\end{tabular}
-\end{center}
-
-\subsection{Comparisons}\label{subsec:comparisons}
-
-Model performance is typically evaluated using accuracy, precision, recall, and AUC metrics, as well as total time to
-train.
-
-% ----------------------------------------------------------------------
-% RESULTS
-% ----------------------------------------------------------------------
-\section{Results}\label{sec:results}
-At the date of the milestone, I have shown that my experimental set-up is complete to begin
-iterating towards the novel contributions aimed for by this project.
-
-I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
-network for classification on my chosen dataset incorporating my own contributions to the process.
-
-I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
-excited nevertheless to try.
-
-\subsection{Data}\label{subsec:data}
-I have become familiar with the chosen dataset, and created a machine learning friendly format
-based off of the data.
-
-I have inferred classification labels from the experimental events timings and the estimated time
-of each complete scan in the dataset.
+The stretch goal of region highlighting using unsupervised segmentation proved infeasible for this project timeframe, given
+the challenges overcome with the 3d convolutional model architecture.
 
@@
 chiefly was finding the right learning rate for the given normalized data to the model
-typically 1e-2 to 1e-4 worked, with 1e-3 being good, and decay being very very helpful
+typically 1e-2 to 1e-4 worked, with 1e-3 being best, and decay being very very helpful
 model quite sensitive to adjustments
@@
 
-\subsection{Performance}\label{subsec:performance}
+\subsection{Performance Measures}\label{subsec:performance}
 
@@
 to reproduce on 3d, and trusting the model this much, open the door to segmentation
-
-% ----------------------------------------------------------------------
-% DISCUSSION
-% ----------------------------------------------------------------------
-\section{Discussion}\label{sec:discussion}
-
-\subsection{Learning}\label{subsec:learning}
-very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
-cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
-this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at
-
-\subsection{Future Works}\label{subsec:future-works}
-Unsupervised segmentation, especially in 3d would be cool
-Future works
-- could strive to refine 3d more, with more time and compute
-- unsupervised segmentation would be cool
-- find ways to not have to smoosh images for memory?
-- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too
-
-began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
-- check out filter viz
-- see about unsupervised
-- try 3d alternative archs? (dont really need to)
-
-might want to improve 3d before going on to segmentation,
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/saliency.png}
-  \caption{caption}
-  \label{fig:saliency}
-\end{figure}
-
-% ----------------------------------------------------------------------
-% END
-% ----------------------------------------------------------------------
-{\small
-\bibliographystyle{ieee}
-\bibliography{project}
-}
-
-% ----------------------------------------------------------------------
-% SUPPLEMENTARY
-% ----------------------------------------------------------------------
-\section{Supplementary Material}\label{sec:supplementary}
-
-Extra material for reference
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/orthoview_t1w.png}
-  \caption{caption}
-  \label{fig:ortho-t1w}
-\end{figure}
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
-  \caption{caption}
-  \label{fig:cortex-var}
-\end{figure}
-
-\end{document}

### 2019-12-03 paper/submission.tex
Breaking out sections to make working on paper easier

--- 2019-12-02 paper/submission.tex
@@
 
+% ------------------------------------------------------------------------
+% ABSTRACT
 %-------------------------------------------------------------------------
 \begin{abstract}
-   Abstract placeholder (sub 300 words)
+ Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
+ typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
+ and limit the ability to stitch together patterns across a third dimension.
+
+ Deep learning's ability to decipher and learn from such nearly impossible patterns in massive datasets,
+ imperceptible to preliminary observation makes it an excellent candidate to explore and teach us about
+ overlooked patterns of activity.
+
+ Here I showcase convolutional network's ability to correctly classify uncleaned activity data from a passive
+ image viewing experiment aimed at triggering activity in particular regions of the brain controlling reward and
+ restraint, using the data in both 2d and 3d to explore the merits of adapting video-processing techniques meant
+ to apply across time to a third spatial dimension and discuss adapting this to unsupervised segmentation to see
+ what the network can teach us from its findings.
+
 \end{abstract}
 
-% ----------------------------------------------------------------------
-% INTRODUCTION
-% ----------------------------------------------------------------------
-\section{Introduction}\label{sec:introduction}
-Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
-for mental health research and disease diagnosis.
-High dimensional, massive datasets, with intertwining patterns of activity and structure in three or four
-dimensions, are a lot to manage by hand.
-
-Classification of activity and structures should come quite naturally with deep learning practices.
-Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
-yet often only as a proof of concept.
-Others are working now to try and find universal architectures for most
-neuro-imaging tasks~\cite{henschel2019fastsurfer}, though it seems this is still far from a solved
-application with satisfactory performances across the board.
-
-Of more interest to myself, deep learning networks in their proven capacity to segment images based on what entities
-they identify in classification could surely be adopted to highlight what patterns of activity have informed their
-own decision making and perhaps clue researchers in to overlooked details.
-
-My ambition is to create a highly accurate 3d convolutional network to first classify, then segment brain scans to
-help with my specific application of neuroimaging.
-
-I may also attempt to explore feature analysis and reduction which has been shown to
-help~\cite{shi2018feature} with deep learning tasks, if only to manage the challenging
-dimensionality of their 3d inputs.
-
-% ----------------------------------------------------------------------
-% PROBLEM STATEMENT
-% ----------------------------------------------------------------------
-\section{Problem Statement}\label{sec:problem-statement}
-
-(BACKGROUND / RELATED)
-
-The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
-Netherlands circa 2013 who were presented with various grids of images featuring
-either neutral entities, (e.g.\ office chairs, windows, jackets) or foods~\cite{smeets2013allured}.
-The foods were specifically types of calorically dense, tempting foods, which the individuals viewing would likely
-have formed restraint with given their success in maintaining healthy body mass indices.
-The experiment was meant to highlight brain structures roles in anticipated rewards from consuming delicious foods,
-balanced against the hypothesized activation of restraints.
-In neuro-imaging, this task is known as passive image viewing.
-
-Given the quite important role of food and willpower in our lives, the activation patterns are likely to be powerful
-and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
-This combination gives me high hopes of success in this specific application of learning to classify the viewed
-objects through brain activations of the subjects.
-
-I hope given the clarity in patterns in the brain scans, that I can develop a neural network to highlight what
-features and locations in the brain scans have been helpful in informing the decisions of classification, in the
-hope that such methods could be adapted to harder tasks in the future.
-
-The dataset can be found from the excellent
-OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}
-It consists of a survey detailing the relationship of each subject with food and diet taken after the experimental
-proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w brain scan,
-and time series of bold scans of about 370-375 scans, which encompasses the approximately 10 minute experiment
-in its entirety.
-The bold scans are quite sizeable given the number of them, however they are a relatively low resolution, consisting
-of 64 sagital by 64 coronal by 30 axial scans.
-
-You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.
-
-You may find this project repository here: \url{https://github.com/e-m-albright/CS682}
-Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
-abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.
-
-% ----------------------------------------------------------------------
-% TECHNICAL APPROACH
-% ----------------------------------------------------------------------
-\section{Technical Approach}\label{sec:technical-approach}
-I intend to explore the efficacy of 3d convolutional networks which have been shown to have a slight performance
-edge on 2d convolutions~\cite{payan2015predicting}.
-Given the nature of the input images being 3d brain scans this makes intuitive sense and could allow for much more
-interesting patterns of brain activity and structuring being recognized.
-
-As for the deeper architecture, I'm hoping to imitate some of the more successful networks like VGG or ResNet,
-however I wouldn't be surprised if they don't adapt well to this particular application, especially given how
-interesting and important it could be to keep the input data in a three dimensional format.
-
-It is my hope that I can also learn from other developments of object detection to see what regions in the
-brain become active for a particular classification, like Faster R-CNN
-
-\subsection{Novelty}\label{subsec:novelty}
-
-I have not seen any application of classification or object segmentation to neuroimaging in 3d
-on brain activity patterns.
-I have seen some works which I hope to leverage with similar aims like that of structural identification
-for tumors, in two dimensions~\cite{akkus2017deep}.
-However at the very least I hope to produce a very successful network on my chosen problem first and foremost rather
-than assume I am the only person to ever think of something as intuitive as bringing neural networks to bear
-on neuro-imaging tasks.
-
-I hope to find a methodology of first classifying correctly the brain scans, improve accuracy as much as possible,
-and adapt the approach into tools for identifying important regions of activity in the
-brain which are being utilized for classification.
-
-I am not chiefly focused on coping with the challenges of memory but it may become too important to overlook.
-Each dataset input is a $64 \times 64 \times 30$ slice of floats, for which we've got
-about $375 scans \times 30 subjects$ total.
-In my work on preliminary results I have already been confronted by the limitations of my current machine and have
-been trying working around the issue until I am ready to attempt to resolve it.
-This may be an unavoidable area to simultaneously try to innovate for the field of
-neuro-imaging analysis with deep learning
-
-% ----------------------------------------------------------------------
-% INTERMEDIATE RESULTS
-% ----------------------------------------------------------------------
-\section{Intermediate Results}\label{sec:intermediate-results}
-At the date of the milestone, I have shown that my experimental set-up is complete to begin
-iterating towards the novel contributions aimed for by this project.
-
-I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
-network for classification on my chosen dataset incorporating my own contributions to the process.
-
-I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
-excited nevertheless to try.
-
-\subsection{Data}\label{subsec:data}
-I have become familiar with the chosen dataset, and created a machine learning friendly format
-based off of the data.
-
-I have inferred classification labels from the experimental events timings and the estimated time
-of each complete scan in the dataset.
-
-\subsection{Sanity Checks}\label{subsec:sanity-checks}
-Using the data, I have implemented a very simple, yet effective 80\% accurate SVM classifier to prove the data
-should be ready for use with my chosen deep learning methodologies.
-
-I'm now confident in the classifications of each of the scans (about 375 per subject, for 30 subjects)
-
-\subsection{Deep Learning \& Future Works}\label{subsec:deep-learning}
-I have tried feeding the data into very basic neural networks with mixed success, and am in the active process of
-troubleshooting.
-Once I have matched performance with the SVM using an implementation of a reasonable neural network using PyTorch, I
-will begin the process of looking for refinements and innovations alluded to in previous sections, as well as
-transitioning the task to object detection if I can.
-
-% ----------------------------------------------------------------------
-% TODO
-% ----------------------------------------------------------------------
-\section{TODO}\label{sec:todo}
-
-NN project
-- 2d hyperparameters, let's use a LR scheduler? LR spikes all over the place still, why?
-
-- 3d CNN - typically meant for video classification using time as the third axis,
-here we are not using time, though that is a dimension we have access to,
-though doesn't make sense for this experiment/project to try to add in
-- https://pytorch.org/docs/stable/torchvision/models.html#video-classification
-
-- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
-- cut batch size to avoid scaling down images (TRIED NO BUENO?)
-
-be sure to talk about everything in paper
-
-normalization/mean image
-interpolation
-wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
-different styles of 1d (unfurl 3d, 2d) 2d (mean flattened, one slice) 3d (scale, orig)
-LR explorations
-sanity checks on brain image data, impossible for me to spot the differences
-
-batch_size and interpolation hard to juggle too?
-3d feels like "luck" plays a big part of getting training to catch/start
-seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
-
-images:
-LR / accuracy measures, of diff approaches
-hi res, mid res, interpolated down res brain scan examples,
-2d slice example
-total training time? num features comparison
-average slice image
-graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
-
-inclusion of more ppts/subjects means more training time (val/test are of same patients,
-hence more individual brain shapes means more general, but longer to learn)
-20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
-
-image segmentation attempt?
-saliency maps?
-filter views?
-
-torch save models - can eval that way
-host for downloads?
-
-future works
-better gpu / computer might be important, the 3d models are real fucking beasts to train, and their data is large as fuck
-model ensemble (free extra few percent perf)
-
-findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
-promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
-unless better hardware or serious effort is expended in getting it differentiable
-
-PUT BRAIN SCANS IN SUPPLEMENTARY?
-
-abstract > intro > background > related
-methods/approach > results > discussion > references
-6 pages? 7 maybe at most
-
-gradual warm-start improvements?
-variations on models? meh
-
-unsupervised segmentation - little hard
-
-surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
-but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
-bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential
-
-WORK ON GETTING PARAM NUMBERS AND
-DIFFERENT METRICS (AUC PLEASE, F1 PLEASE)
-2d is enough
-
-Submit your final submission through OpenReview. You will submit a single pdf file:
-A pdf file of your final report
-
-(OPTIONAL) supplementary can be concatenated at the end of the final report
-Report. The following is a suggested structure for the report:
-Title, Author(s)
-
-Abstract: It should not be more than 300 words
-
-Introduction: this section introduces your problem, and the overall plan for approaching your problem
-Background/Related Work: This section discusses relevant literature for your project
-Approach: This section details the framework of your project. Be specific, which means you might want to include equations, figures, plots, etc
-Experiment: This section begins with what kind of experiments you're doing, what kind of dataset(s) you're using, and what is the way you measure or evaluate your results.
-It then shows in details the results of your experiments.
-By details, we mean both quantitative evaluations (show numbers, figures, tables, etc) as well as qualitative results (show images, example results, etc).
-Conclusion: What have you learned? Suggest future ideas.
-References: This is absolutely necessary.
-Supplementary Material is not counted toward your 6-8 page limit. It is optional and is supposed to contain less important results/experiments/etc not critical for understanding the main report.
-
-try different architectures - more depth?
-see if we can break 85% to 90%, talk about potential data mislabeling frustration as a barrier to higher results possibly
-
-you are expected to show substantially more experiments and results with NN-related methods
-( :(  lame)
-
-~\cite{he2016deep} resnet
-~\cite{tran2018closer} 3d/time resnet
-~\cite{hara3dcnns} 3d cnn pt1?
+% ------------------------------------------------------------------------
+% BODY
+%-------------------------------------------------------------------------
+\input{sections/intro.tex}
+\input{sections/background.tex}
+\input{sections/approach.tex}
+\input{sections/results}
+\input{sections/discussion.tex}
 
@@
 
+% ----------------------------------------------------------------------
+% SUPPLEMENTARY
+% ----------------------------------------------------------------------
+\section{Supplementary Material}\label{sec:supplementary}
+
+Extra material for reference
+
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
+  \caption{caption}
+  \label{fig:cortex-var}
+\end{figure}
+
 \end{document}

### 2019-12-03 paper/sections/discussion.tex
Just need discussion now

@@
-\usepackage{graphicx}
-
 \section{Discussion}\label{sec:discussion}
@@
 might want to improve 3d before going on to segmentation,
+- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
+wowee good stuff, hope to use a LR scheduler in the future to try to control this better?
 
+ I'd be stoked
+to reproduce on 3d, and trusting the model this much, open the door to segmentation
+
+model ensemble (free extra few percent perf)
  \begin{figure}

### 2019-12-03 paper/sections/results.tex
Just need discussion now

@@
 Both 2 dimensional and 3 dimensional convolutional networks were able to surpass the benchmarks,
-attaining 95\%+ and 90\%+ accuracy respectively.
+attaining 95\%+ accuracy.
 
@@
 chiefly was finding the right learning rate for the given normalized data to the model
-typically 1e-2 to 1e-4 worked, with 1e-3 being best, and decay being very very helpful
-model quite sensitive to adjustments
+typically 1e-2 to 1e-4 worked, with learning rates on the order of 1e-3 being best.
+Stochastic Gradient Descent with nesterov momentum on Cross Entropy Loss performed very well.
+Other optimizers like Adam and loss metrics like NLL Loss explored were either on par or worse and thus not
+further utilized.
 
-- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types
+None of my attempted customizations to the ResNet architecture netted any improvements.
 
-\subsection{Scope of Number of Subjects}\label{subsec:num-subjects}
+\subsection{Subject individuality}\label{subsec:subjects-individ}
 
-inclusion of more ppts/subjects means more training time (val/test are of same patients,
-hence more individual brain shapes means more general, but longer to learn)
-20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts
+Interestingly, performance of models on subjects not included in the training run was subpar.
+For example, a model trained to 95\% accuracy on 6 subjects alone would fair closer to 60\% on 6 new subjects.
+I speculate this is not overly worrisome, and a consequence of using such a small handful of subjects.
+Better generalizability seems likely given an investment in much larger machinery or methodology capable of handling
+more subjects, or better normalization practices.
+One exploration of the data revealed by far most of the variance comes at the perimeter of the cortex suggesting slightly
+varied sizes of brains not overlapping perfectly.
+Convolutional networks should be able to tackle this dilemma with no problem but perhaps the small dataset makes
+spatial and orientation noise more consequential.
+
+\subsection{Training Progress}\label{subsec:training-prog}
+
+Generally, training catches within about 20 epochs, learning to differentiate better than chance on 2
+dimensional data~\ref{fig:2d_cold}, taking noticeably longer for 3 dimensional data~\ref{fig:3d_cold} to catch up,
+as there's presumably more nuance to learn.
+The learning rate may seem fairly aggressive~\ref{fig:cold_500_loss.png}
+and takes one step back for every two steps forward, but overall produces
+more competitive results than a gentler learning rate.
 
@@
 
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/3d_cold.png}
+  \caption{Cold-Start 3D Training Progress}
+  \label{fig:3d_cold}
+\end{figure}
+
+ \begin{figure}
+  \includegraphics[width=\linewidth]{images/cold_500_loss.png}
+  \caption{Cold-Start Training Progress Loss}
+  \label{fig:cold_500_loss.png}
+\end{figure}
+
 \subsection{Warm Starts}\label{subsec:warm-starts}
-saving model parameter dict, loading and training again
-wowee good stuff, hope to use a LR scheduler in the future to try to control this better?
-surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
-but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
-bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential
+
+Saving the model parameters and training again on a new subset of subjects or even resuming learning
+on the same division worked wonderfully.
+
+Getting the model to a stable, low 80\% performance leaves us in an excellent place to begin aggressively learning again.
+
+With this method performances of over 95\%~\ref{fig:2d_warm} are common.
+
+As mentioned before however, higher performances are typically seen on the same subject, though it doesn't take
+long for the model to adapt to the new group of subjects for training, but of course it sacrifices performance on the
+previous grouping to do so.
 
@@
 \subsection{3 Dimensions}\label{subsec:3-dimensions}
-- 3d CNN - typically meant for video classification using time as the third axis,
-here we are not using time, though that is a dimension we have access to,
-though doesn't make sense for this experiment/project to try to add in
-- https://pytorch.org/docs/stable/torchvision/models.html#video-classification
 
-okay it works, and has potential, certainly given the data is richer to be the best, but wow it's frustratingly slow
-findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
-promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
-unless better hardware or serious effort is expended in getting it differentiable
+One of the biggest challenges of this endeavor has been coping with the computational demands that the third
+dimension adds.
+Video data is notoriously large, and now I've seen 3d CNN's can be as well.
+Our application here substitutes time for a third spatial axis, and does seem to be a richer dataset to draw from for
+classification.
+Training on a larger, more complex dataset does feel more challenging as well, with training taking longer to catch on
+to something productive.
 
- \begin{figure}
-  \includegraphics[width=\linewidth]{images/3d_cold.png}
-  \caption{Cold-Start 3D Training Progress}
-  \label{fig:3d_cold}
-\end{figure}
-wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
-different styles of 1d (unfurl 3d, 2d) 2d (mean flattened, one slice) 3d (scale, orig)
-LR explorations
-
-future works
-better gpu / computer might be important, the 3d models are real fucking beasts to train, and their data is large as fuck
-model ensemble (free extra few percent perf)
-
-"none of my alterations to the model architecture worked that well, most did nothing, a few broke things, could still poke around more though"
-graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
-batch_size and interpolation hard to juggle too?
-3d feels like "luck" plays a big part of getting training to catch/start
-seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)
-
-cut batch size to avoid scaling down images (TRIED NO BUENO?)
+Unfortunately, I've found iterating with 3 dimensional models to be much harder than 2d which makes refinement also a
+bit of a challenge.
+Where training may take 15 minutes with a 2d CNN, it might be 5 hours or more with a 3d CNN, and the limitations on the
+quality and abundance of the dataset size kick in much faster than for 2d.
+The dataset is already a fairly conservative low-grade rendering of the brain and aggresive interpolation cuts into
+performance directly.
+Additionally, lowering the batch size is very detrimental to performance, with at least 64 images being critical to
+learning generalizable patterns at each step in an epoch.
+However, what I have seen in working with the 3d CNN's is that they are very performant, achieving on par or
+3 to 4\% higher classification accuracy compared to their 2d counterparts, and responding encouragingly as well to warm-starts,
+to which I'm very pleased.
 
 \subsection{Performance Measures}\label{subsec:performance}
+
+Here I examine one highly performant 2d convolutional network in more detail.
+Thankfully it is a well balanced model, with seemingly high intuitive understanding of the differentiable characteristics
+of brain activity, as it's highly accurate, and highly confident in it's decisions.
+
+While I have been able to tweak the model to exceed this accuracy, this is in my opinion the most balanced
+run observed as typically the model begins to favor performance on one class over the other despite near
+equal representation.
 
@@
 \end{center}
-
-Can get a little higher but these are already encouraging numbers (96 seen) often the model starts leaning a little
-one way or the other and this was one of the most balanced runs I've seen
-
-I'm a little unclear about the reliability of my data labeling, so I'm happy with this result, I'd be stoked
-to reproduce on 3d, and trusting the model this much, open the door to segmentation

### 2019-12-03 paper/sections/approach.tex
Final paper edits

@@
 Beginning with SKLearn~\cite{scikit-learn} data readiness was proven with moderately successful classification using
-a simple SVM and then the SKLearn Multi-Layer-Perceptron neural network, getting 70-80\% accuracy, sufficing the
+a simple SVM and then the SKLearn Multi-Layer-Perceptron neural network, getting 70-80\% accuracy, well above the
+random chance accuracy of 50\%, sufficing the
 initial sanity check that deep learning would be ready to work on.

### 2019-12-03 paper/sections/discussion.tex
Final paper edits

--- 2019-12-03 paper/sections/discussion.tex
@@
 \subsection{Learning}\label{subsec:learning}
-very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
-cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
-this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at
+
+Using real world brain scans of a very low resolution, we've been able to achieve excellent classification using both
+2 dimensional and 3 dimensional convolutional networks, with somewhat promising generalizability to previously unseen subjects.
+The models have been very able to differentiate when a subject was scanned viewing a tempting food image against when the image
+presented is neutral, where I personally can see absolutely no clear difference.
+
+Unsurprisingly deeper architectures also tend to improve performance marginally,
+as they can cope with increasingly subtle changes presumably, in both the performance of the 2d and 3d CNN's
+
+While we may have empirically found improved performance using 3 dimensional CNN's over all alternatives,
+the hefty cost of computation has proven difficult to cope with and limited the ability to explore segmentation farther.
 
 \subsection{Future Works}\label{subsec:future-works}
-Unsupervised segmentation, especially in 3d would be cool
-Future works
-- could strive to refine 3d more, with more time and compute
-- unsupervised segmentation would be cool
-- find ways to not have to smoosh images for memory?
-- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too
 
-began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
-- check out filter viz
-- see about unsupervised
-- try 3d alternative archs? (dont really need to)
+While classification feels well proven and a desirable performance was achieved,
+the natural next step is to try an highlight what the network has learned to recognize
+as indicative of a particular stimuli, turning the blackbox computation into a powerful teacher.
+My attempts made to the point of this submission to adapt the classifiers to unsupervised segmentation
+have not yet been fully successful.
 
-might want to improve 3d before going on to segmentation,
+\subsubsection{Refinements}\label{refinements}
+
+Refinements to the classifier are always a worthwhile endeavor, though this particular classification feels well
+handled by the chosen architectures at relatively shallow levels.
+
+Given the promise of warm-starts progressing learning, I'd be very interested in exploring different
+learning rate schedulers to see what combination of minute and aggressive changes lands the model in the most performant
+optimization.
+
+Lastly as there's typically a free boost from using an ensemble model, I'd be interested in exploring what combination
+of models yields a more robust cohesive model which generalizes very well to previously unseen subject's brain scans.
+
+\subsubsection{Visualizations}\label{fw-viz}
+
+The models are also naturally capable of producing visualizations of their convolutional filters, class representations
+and creating saliency maps.
+
+In figure~\ref{fig:saliency} we see one example of Saliency maps in action.
+Interestingly enough the model does not key strongly off of discrete brain structures as I had assumed it would.
+I speculate that the data being so downgraded provides difficulty in precisely conveying brain activity and the models
+are instead forced to learn off of other subtle activities across the brain, which are of course not fully silent during
+the passive image viewing task.
 
@@
   \includegraphics[width=\linewidth]{images/saliency.png}
-  \caption{caption}
+  \caption{Three Nonfood \& Three Food Brain Scan Slices with their respective Saliency Maps}
   \label{fig:saliency}

### 2019-12-03 paper/sections/results.tex
Final paper edits

@@
 
-the project explored minimally the hyperparameters of the chosen model architectures
-chiefly was finding the right learning rate for the given normalized data to the model
-typically 1e-2 to 1e-4 worked, with learning rates on the order of 1e-3 being best.
+I explored the available hyperparameters of the chosen model architectures to a minor degree, focusing on
+establishing optimal learning rates with the chosen optimizer and model, with the given normalized data.
+
+Typically 1e-2 to 1e-4 worked, with learning rates on the order of 1e-3 being best for both ResNet style convolutional
+networks.
+
 Stochastic Gradient Descent with nesterov momentum on Cross Entropy Loss performed very well.
@@
 
-None of my attempted customizations to the ResNet architecture netted any improvements.
+None of my attempted customizations to the ResNet architecture netted any improvements, including expanding filter sizes
+or adding normalization layers.
 

### 2019-12-03 paper/submission.tex
Final paper edits

@@
 
-\author{Anonymized for Peer Review\\
+\author{Anonymous1\\
 University of Massachusetts Amherst\\
@@
 \begin{abstract}
- Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
+ Analyzing the brain's activity patterns starts with identifying known signals of well established structures,
  typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
@@
   \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
-  \caption{caption}
+  \caption{Region of Variance Across the Dataset of Brain Scans}
   \label{fig:cortex-var}
